---
format: 
  pdf:
    geometry: "margin=2cm"
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"
editor: source
toc: false
embed-resources: true 
execute: 
  echo: false
  message: false
  warning: false
  fig-width: 12
  fig-height: 8
params:
  year: 2025
  country_code: 'NU'
  data_provided: 'Yes'
  aceman: 'aceman'
  additional_sections: ''
---

```{r}
# Load necessary libraries
library(DBI)
library(odbc)
library(tidyverse)
library(flextable)
library(scatterpie)
library(mregions2)
library(countrycode)
source("utils.R")

sp_colours <- setNames(object = c("seagreen3","red3","black","cyan1","steelblue","orange",
                                  "pink","wheat","purple3","grey40","dodgerblue4","purple4",
                                  "#66CCFF","salmon4","magenta3","gold", "gray"),
                         nm = c("Albacore","Bigeye tuna","Black marlin","Blue marlin","Blue shark",
                                "Hammerhead sharks nei","Mako sharks","Oceanic whitetip shark",
                                "Pacific bluefin tuna","Silky shark","Skipjack tuna",
                                "Striped marlin","Swordfish","Thresher sharks nei",
                                "Whale shark","Yellowfin tuna", "Other"))
# adjust path as needed
flag_lookup <- read_csv("../ofp-dm-part1-report-template/config/flag_names.csv") |>
  select(country_code, country_name, country_short) |>
  mutate(country_name = na_if(country_name, "NULL"),
         country_short = na_if(country_short, "NULL")) |>
  # use short name when available, fall back to long
  mutate(name_pretty = coalesce(str_to_title(tolower(str_trim(country_short))), str_to_title(country_name)))

flag_labels <- flag_lookup |>
  mutate(flag_code_lower = tolower(country_code)) |>
  select(flag_code_lower, name_pretty) |>
  deframe()  # from tibble: turns 2-col df into a named vector

# Helper: get a country's 200NM EEZ as an sf object, keyed by ISO-2 code
get_country_eez <- function(iso2_code) {
  iso3 <- countrycode(iso2_code, "iso2c", "iso3c", warn = FALSE)
  
  if (is.na(iso3)) {
    stop("Could not resolve ISO-2 code '", iso2_code, "' to ISO-3.")
  }
  
  eez <- mrp_get(
    "eez",
    cql_filter = sprintf("iso_ter1 = '%s' AND pol_type = '200NM'", iso3)
  )
  
  if (nrow(eez) == 0) {
    warning("No EEZ found for '", iso2_code, "' (ISO-3: ", iso3, ").")
  }
  
  eez
}

# Helper: build a gridded species-composition catch summary for one EEZ
grid_catch <- function(data, eez_code, cell_size = 1,
                       species = c("skj", "yft", "bet")) {
  
  species_cols <- paste0(species, "_mt")
  half <- cell_size / 2
  
  data |>
    filter(eez == eez_code) |>
    mutate(
      lat_bin = floor(latd / cell_size) * cell_size + half,
      lon_bin = floor(lond / cell_size) * cell_size + half
    ) |>
    group_by(lon_bin, lat_bin) |>
    summarise(across(all_of(species_cols), \(x) sum(x, na.rm = TRUE)),
              .groups = "drop") |>
    rename_with(\(x) str_remove(x, "_mt$"), all_of(species_cols)) |>
    mutate(total = rowSums(across(all_of(species)))) |>
    filter(total > 0)
}

# Detect context: standalone render vs. knit_child from parent
# - Standalone: params$country_code exists (set via execute_params)
# - Child:      country_cd is already defined by the parent
if (!exists("country_cd") || length(country_cd) == 0 || !nzchar(country_cd)) {
  country_cd <- tolower(params$country_code)
}

if (!exists("report_year") || length(report_year) == 0) {
  report_year <- params$year
}

if (!exists("this_yr_folder") || length(this_yr_folder) == 0 || !nzchar(this_yr_folder)) {
  this_yr_folder <- paste0("./data/report_", as.character(report_year), "_", country_cd, "/") |>
    tolower()
}

# Sanity check — fail loudly if anything is still missing
stopifnot(
  length(country_cd)     == 1, nzchar(country_cd),
  length(report_year)    == 1,
  length(this_yr_folder) == 1, nzchar(this_yr_folder)
)

# Define years to use for tables and figures
yrs_long <- (report_year - 4):report_year
yr_range <- paste(min(yrs_long), max(yrs_long), sep = "-")
'%nin%' <- function(x, y) !x %in% y

country_name <- flag_labels[[country_cd]]
```

```{r data-prep}
# Load all raw data - from t2
data_3615 <- read_and_clean(this_yr_folder, country_cd, r_code = 3615) |>
      mutate(across(where(is.character), tolower))

data_3527 <- read_and_clean(this_yr_folder, country_cd, r_code = 3527)|>
      mutate(across(where(is.character), tolower))

data_3614 <- read_and_clean(this_yr_folder, country_cd, r_code = 3614)|>
      mutate(across(where(is.character), tolower))

ika_trips_file = paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")

if (file.exists(ika_trips_file)){
  data_ika_raw <- tryCatch(
    read.csv(paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")),
    error = function(e) NULL
  )
  
  if (exists("data_ika_raw")){
    if ("LandingSite" %in% colnames(data_ika_raw) && nrow(data_ika_raw) > 0){
        data_ika_raw <- data_ika_raw|>
          rename(landing_site_name = LandingSite) |>
          group_by(Country, landing_site_name, Year, Month) |>  
          summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop") |>
          filter(Year == report_year)
    }
    
  }
  
  if (is.null(data_ika_raw) || nrow(data_ika_raw) == 0 || is.null(data_ika_raw)) rm(data_ika_raw)
  }

if (!exists("data_ika_raw") && nrow(data_3615) == 0) {
  cat("No Ikasavea or Tails trip data found for this country/year. Document generation stopped.\n")
  knitr::knit_exit()
}

if (exists("data_ika_raw")){
  
   data_ika_trips <- data_ika_raw |>
    mutate(across(where(is.character), tolower)) |>
    janitor::clean_names() |>
    select(landing_site_name, month,
           number_trips)
  
  df_ika_wide <- data_ika_trips %>%
    complete(landing_site_name, month = 1:12, fill = list(number_trips = 0)) %>%
    mutate(month = month.abb[month]) %>%
    pivot_wider(names_from = month, values_from = number_trips) %>%
    select(landing_site_name, all_of(month.abb))
  
  data_ika_catch_kg <- data_ika_raw |>
    mutate(across(where(is.character), tolower)) |>
    janitor::clean_names() |>
    select(landing_site_name, month, skj = skj_kg, yft = yft_kg, bet = bet_kg)#, alb = alb_kg)
  
  data_ika_catch_kg <- data_ika_catch_kg |>
    pivot_longer(cols = c("skj", "bet", "yft"), names_to = "sp_code", values_to = "sp_kg")
  
  }

both_exist <- exists('data_ika_trips') && nrow(data_ika_trips) > 0 && exists('data_3615') && 
              is.data.frame(get('data_3615')) && nrow(data_3615) > 0

local_knowl_trips_file <-  str_c(this_yr_folder,  'additional_files/est_trips.csv')

```

\newpage

# Abstract

Niue is not a flag state and does not host transshipments in its ports or EEZ, and is therefore exempt from compiling the information required in Annual Report Part 1 under WCPFC Conservation and Management Measures (CMMs). Yet, this report compiles relevant information on Niue's domestic artisanal fishery, including raised and unraised tuna catch estimates.

# Background

Niue has not licensed any foreign fishing vessels for the past six years, resulting in no offshore catch to report for `r params$year`. Domestic artisanal fishers continue to supply restaurants, families, and local markets, though at an inconsistent rate. Reduced annual catches have at times made it difficult to maintain reliable supply for food security. Shark depredation has also contributed to the decline, significantly impacting fishers' gear and morale.

As tourism grows, marine activities have diversified accordingly. Niue Fisheries is working with the Niue Ocean Wide (NOW) project to establish management arrangements for Niue's newest and largest marine protected area. The MPA will cover significant inshore and offshore habitats and species, as well as fishing and non-fishing activities. Catch-and-release sport fisheries may be permitted within the area under licence.

\newpage

# Flag state reporting

Although Niue is not a flag state, this section presents annual catch estimates derived from artisanal fishery data. To improve the estimate of annual production, catch landings recorded in *TAILS* are raised using estimated fishing activity --- the number of boats and canoes estimated to fish daily at each landing site. \@tbl-raised-ace presents the raised and unraised estimates per species for 2025, and \@tbl-raised-ace-perc shows the percentage contribution of each species.

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: Monthly-effort
#| tbl-cap: !expr if(nrow(data_3615) > 0) glue::glue("Number of trips per month and landing site reported in Tails") else ""

# Check if data exists
if (!(nrow(data_3615) == 0)) {
  
  names(data_3615) <- gsub(x = names(data_3615), pattern = "_trips$", replacement = "")  
  
  data_3615 <- data_3615 |>
    rename_with(~ month.abb[match(., tolower(month.abb))], 
              .cols = any_of(tolower(month.abb))) |> 
    select(-entity)
  
  flextable(data_3615) |>  
  set_header_labels(landing_site_name = "Landing Site") |>
  width(j=1, width=1.2) |> 
  fontsize(size = 8, part = "all") |>
  fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: Monthly-effort-ika
#| tbl-cap: !expr if(exists('data_ika_trips')) glue::glue("Number of trips per month and landing site reported in Ikasavea") else ""
if (exists('data_ika_trips')){
  flextable(df_ika_wide) |> 
  set_header_labels(landing_site_name = "Landing Site") |>

  width(j=1, width=1.2) |> 
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
  }
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr both_exist

if (both_exist) {
  cat("### Differences in Landing Sites: Ikasavea + Tails\n\n")
  cat("Artisanal data is collected using Tails and in Ikasavea platforms. The summary table below of all landing sites common to T2 and Ikasavea as well the ones unique to each of these sources (if any).")
}
```

```{r}
#| echo: false
#| include: false
#| eval: !expr both_exist
#| results: 'asis'
#| label: diff_t2_ika

if (both_exist) {

  ls_t2  <- tolower(trimws(unique(data_3615$landing_site_name)))
  ls_ika <- tolower(trimws(unique(df_ika_wide$landing_site_name)))
  
  in_both    <- intersect(ls_t2, ls_ika)
  only_t2    <- setdiff(ls_t2, ls_ika)
  only_ika   <- setdiff(ls_ika, ls_t2)
  
  if (length(only_t2) == 0 && length(only_ika) == 0) {
    cat("✅ All landing site names match perfectly between both sources.\n")
  } else {
    
    # Summary table
    summary_df <- data.frame(
      ` ` = c("Matched in both", "Only in T2 (data_3615)", "Only in Ikasavea"),
      `Count` = c(length(in_both), length(only_t2), length(only_ika)),
      check.names = FALSE
    )
    
    # cat("### Landing site name comparison\n\n")
      flextable(summary_df) |>  
        set_header_labels(landing_site_name = "Landing Site") |>
         width(j=1, width=1.2) |> 
        fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)

    
    # Table of mismatches
    if (length(only_t2) > 0 || length(only_ika) > 0) {
      max_len <- max(length(only_t2), length(only_ika))
      mismatch_df <- data.frame(
        `Only in T2 (data_3615)` = c(only_t2,  rep("", max_len - length(only_t2))),
        `Only in Ikasavea`       = c(only_ika, rep("", max_len - length(only_ika))),
        `In both` = c(in_both, rep("", max_len - length(in_both))),
        check.names = FALSE
      )
      # cat("\n### Unmatched landing site names\n\n")
      flextable(mismatch_df) |>  
        set_header_labels(landing_site_name = "Landing Site") |>
         width(j=1, width=1.2) |> 
        fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
    }
  }
}

```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr both_exist

if (both_exist) {
  cat("### Trips: Ikasavea + Tails\n\n")
  cat("The final number of trips combining what was reported on these two platforms is presented in the table below.")
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: tails_ika_summary

# ika and tails combined
if (both_exist) {
  prep <- function(df) {
    df %>%
      mutate(landing_site_name = tolower(trimws(landing_site_name))) %>%
      pivot_longer(-landing_site_name, names_to = "month", values_to = "trips") |>
      mutate(month = tolower(gsub("_trips", "", month)))
  }
  
  combined_t2_ika <- bind_rows(prep(data_3615), prep(df_ika_wide)) %>%
    group_by(landing_site_name, month) %>%
    summarise(trips = sum(trips), .groups = "drop") %>%
    mutate(month = factor(month, levels = tolower(month.abb))) %>%
    pivot_wider(names_from = month, values_from = trips) %>%
    arrange(landing_site_name) |>
    select(landing_site_name, tolower(month.abb))
  
  ft <- flextable(combined_t2_ika) |>  

    width(j=1, width=1.2) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
  
  # flextable::htmltools_value(ft)
  ft

} else {
  combined_t2_ika <- data.frame()
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr nrow(data_3615) > 0 && file.exists(local_knowl_trips_file)

if (nrow(data_3615) > 0 && file.exists(local_knowl_trips_file)) {
  cat("### Trips: Local Knowledge\n\n")
  cat("Member-estimated **trips** per month at each landing site are presented in the table below. These estimates are used to compare with what is reported in the ereportng tools to produce a 'best' estimate of effort.")
}
```

```{r trips-est-mem}
#| echo: false
#| include: false
#| results: 'asis'
#| label: mem-trips
#| tbl-cap: !expr if(nrow(data_3615) > 0) glue::glue("Member reported estimated trips per month per landing site") else ""

if (file.exists(local_knowl_trips_file)){
  local_knowl_trips = read.csv(local_knowl_trips_file)
  flextable(local_knowl_trips)  |>  
    fontsize(size = 8, part = "all") |> 
    width(j=1, width=2) |>
        fit_to_width(max_width = 6.5)

}

```

<!-- ### Best estimate of fishing trips -->

<!-- Estimate monthly trips are based on the **max** *value either reported by the member or reported in the ereporting tools Tails and Ikasavea*. In case ereporting data is collected through Tails and Ikasavea tools the combination of trips in both sources will be used as your best estimate monthly trips (tab above). -->

```{r prep-ereporting_trips, echo = F}

# check logic to be applied for ereporting: tails, ikasavea or tails + ikasavea?

if (nrow(combined_t2_ika)>0){ # it means tails + ikasavea
  
  ereporting_trips <- combined_t2_ika |>
    mutate(landing_site_name = str_squish(landing_site_name)) |>
    data.frame()
  ereporting_trips_tag <- c("Trips: Ikasavea + Tails")
  
  }else if (!exists("data_ika_trips") && nrow(data_3615) > 0){ # so it means you only have tails
    ereporting_trips <- data_3615 |>
      mutate(landing_site_name = str_squish(landing_site_name)) |>
      data.frame()
    ereporting_trips_tag <- c("Trips: Tails")
      
  }else if(nrow(data_3615) == 0 && nrow(data_ika_trips) > 0){ # it means you only have ikasavea 
    ereporting_trips <- df_ika_wide |>
      data.frame()
    ereporting_trips_tag <- c("Trips: Ikasavea")
  }else{
    ereporting_trips <- data.frame()
    ereporting_trips_tag <- c("")
  }
```

```{r max.trips}
#| echo: false
#| include: false
#| results: 'asis'
#| label: est-trips-monthly
#| tbl-cap: !expr if(nrow(ereporting_trips) > 0) glue::glue("Best estimate of trips per month at each landing site") else ""
 
if (nrow(ereporting_trips) > 0){
  
  if (file.exists(local_knowl_trips_file)){
    
    # check if columns are correct, if not, flag the issue
    if (!all(c("landing_site_name", "est.trips") %in% colnames(local_knowl_trips))) {
      cat("Your estimated trip file has to have the following columns: 'landing_site_name', 'est.trips'")
      best_estimate_trips <- ereporting_trips
        
    }else{
      
      # if all columns are in the df, groom local_knowl_trips and produce best_estimate_trips
      local_knowl_trips <- local_knowl_trips |>
        mutate(across(where(is.character), tolower)) |>
        mutate(landing_site_name = str_squish(landing_site_name))

      best_estimate_trips = ereporting_trips |> 
        full_join(local_knowl_trips |> 
                    select(landing_site_name, est.trips))
      
      best_estimate_trips[is.na(best_estimate_trips)] <- 0
      best_estimate_trips <- data.frame(best_estimate_trips)
      for( i in 2:13){
        best_estimate_trips[ ,i] = ifelse( best_estimate_trips[ ,i] > best_estimate_trips[ ,14],  
                                           best_estimate_trips[ ,i] , best_estimate_trips[ ,14] )
        }

      # Get rid of sites that only have NAs
      best_estimate_trips$num_na_cols <- rowSums(best_estimate_trips[, 2:13] )

      best_estimate_trips <- best_estimate_trips |>  filter(num_na_cols>0) |>
        select(-c(est.trips, num_na_cols))
    }
    }
  else{
    cat(paste0("There is no member-estimated trips available, so the best estimate fishing is the table above: <b>", ereporting_trips_tag,
             "</b>"))    
    best_estimate_trips <- ereporting_trips
  }
  }

flextable(best_estimate_trips)  |>    
  set_header_labels(landing_site_name = "Landing Site") |>

  fontsize(size = 8, part = "all") |>
  fit_to_width(max_width = 6.5)

```

<!-- ## Catch Estimates -->

```{r prep-catch-data}
# ── Month number → abbreviated name (keeps columns in calendar order) ─────────
month_labels <- c("Jan","Feb","Mar","Apr","May","Jun",
                  "Jul","Aug","Sep","Oct","Nov","Dec")

# ── Helper: pivot one subset to wide format ───────────────────────────────────
make_wide <- function(data) {
  data %>%
    # Sum sp_kg for each landing_site × month combination
    group_by(landing_site_name, month) %>%
    summarise(sp_kg = sum(sp_kg, na.rm = TRUE), .groups = "drop") %>%
    # Convert month numbers to abbreviated names
    mutate(month = month_labels[month]) %>%
    # Ensure all 12 months are present before pivoting
    complete(
      landing_site_name,
      month = month_labels,
      fill  = list(sp_kg = 0)
    ) %>%
    # Pivot to wide
    pivot_wider(
      names_from  = month,
      values_from = sp_kg,
      values_fill = 0          # missing combos → 0
    ) %>%
    # Reorder columns so months appear Jan → Dec (only those present in data)
    select(landing_site_name, any_of(month_labels))
}

# helper to combine ika and tails
combine_wide <- function(df1, df2) {
  if (is.null(df1) && is.null(df2)) return(NULL)
  
  month_cols <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
  
  drop_zero_rows <- function(df) {
    if (is.null(df)) return(NULL)
    df |> filter(if_any(all_of(month_cols), ~ . != 0))
  }
  
  df1 <- drop_zero_rows(df1)
  df2 <- drop_zero_rows(df2)
  
  if (is.null(df1) || nrow(df1) == 0) return(df2)
  if (is.null(df2) || nrow(df2) == 0) return(df1)
  
  full_join(df1, df2, by = "landing_site_name", suffix = c("_a", "_b")) |>
    mutate(across(all_of(paste0(month_cols, "_a")), 
                  ~ coalesce(., 0L))) |>
    mutate(across(all_of(paste0(month_cols, "_b")), 
                  ~ coalesce(., 0L))) |>
    mutate(Jan = Jan_a + Jan_b, Feb = Feb_a + Feb_b,
           Mar = Mar_a + Mar_b, Apr = Apr_a + Apr_b,
           May = May_a + May_b, Jun = Jun_a + Jun_b,
           Jul = Jul_a + Jul_b, Aug = Aug_a + Aug_b,
           Sep = Sep_a + Sep_b, Oct = Oct_a + Oct_b,
           Nov = Nov_a + Nov_b, Dec = Dec_a + Dec_b) |>
    select(landing_site_name, all_of(month_cols))
}

# Check if tails data exists - if yes, reshape and create a df per sp
if (nrow(data_3614) > 0) {
  df_by_species <- data_3614 |>
    mutate(landing_site_name = str_squish(landing_site_name)) |>
    group_by(sp_code) %>%
    group_map(~ make_wide(.x), .keep = FALSE) %>%
    setNames(sort(unique(data_3614$sp_code))) 
  
  df_bet <- df_by_species[["bet"]]
  df_skj <- df_by_species[["skj"]]
  df_yft <- df_by_species[["yft"]]
}

# Check if ika data exists - if yeas, reshape and create a df per sp
if(exists('data_ika_catch_kg') && is.data.frame(get('data_ika_catch_kg'))){
  
  df_by_species_ika <- data_ika_catch_kg %>%
    mutate(landing_site_name = str_squish(landing_site_name)) |>

    group_by(sp_code) %>%
    group_map(~ make_wide(.x), .keep = FALSE) %>%
    setNames(sort(unique(data_ika_catch_kg$sp_code))) 
  
  df_bet_ika <- df_by_species_ika[["bet"]]
  df_skj_ika <- df_by_species_ika[["skj"]]
  df_yft_ika <- df_by_species_ika[["yft"]]
  
  
  # If required (if ika and tails exist), combine catch values from both sources
  if (nrow(data_3614) > 0){
    df_bet_total <- combine_wide(df_bet, df_bet_ika)
    df_skj_total <- combine_wide(df_skj, df_skj_ika)
    df_yft_total <- combine_wide(df_yft, df_yft_ika)
  } else{
    # Just use the ika source if t2 data doesn't exist
    df_bet_total <- df_bet_ika
    df_skj_total <- df_skj_ika
    df_yft_total <- df_yft_ika
    
  }
} else {
  # Just use catch from tails if ika data doesn't exist
  df_bet_total <- df_bet
  df_skj_total <- df_skj
  df_yft_total <- df_yft
}
```

```{r prep-unr-data}
# Calculate total effort for the year by landing site using the ereporting trips
tot_trips_ereport = data.frame(landing_site_name = ereporting_trips$landing_site_name,
                    tot_trips = rowSums(ereporting_trips[,-1],na.rm=T)) |>
                    mutate(landing_site_name = str_squish(landing_site_name))
# Reshape trips
best_estimate_trips_long <- best_estimate_trips |>
  janitor::clean_names() |>
  mutate(landing_site_name = str_squish(landing_site_name)) |>
  pivot_longer(cols = 2:13, names_to = "months", values_to = "tot_trips_best")

ereporting_trips_long <- ereporting_trips |>
  janitor::clean_names() |>
  mutate(landing_site_name = str_squish(landing_site_name)) |>
  pivot_longer(cols = 2:13, names_to = "months", values_to = "tot_trips_spc")

# Check if data exists
species = c("skj", "bet", "yft")

unr_list <- list()
raised_list <- list()
for (sp in species){
  df_name <- paste0("df_", sp, "_total")
  if (exists(df_name) && is.data.frame(get(df_name)) && nrow(get(df_name)) > 0) {
    df <- get(df_name)
    # Now calculate monthly average catch per trip
    avg_sp = data.frame(landing_site_name = df$landing_site_name,
                    tot_catch = rowSums(df[,-1],na.rm=T)) |>
                  left_join(tot_trips_ereport) |>
      mutate(cpd = round(tot_catch/tot_trips,0))
    
    unr_sp <- df  |> 
      left_join(avg_sp |> 
                  select(landing_site_name, 
                         cpd))
    
    # Add result to unraised list
    unr_list[[sp]] <- unr_sp # 
    
    # Calculate raised
    unr_sp_long <- unr_sp |>
      janitor::clean_names() |>
      mutate(landing_site_name = str_squish(landing_site_name)) |>
      pivot_longer(cols = 2:13, names_to = "months", values_to = "unr_catch")
  
    raised_df <- unr_sp_long |>
      left_join(ereporting_trips_long, by = c("landing_site_name", "months")) |>
      left_join(best_estimate_trips_long, by = c("landing_site_name", "months")) |>
      mutate(raised = round(case_when(
        is.na(tot_trips_spc) | tot_trips_spc == 0 ~ cpd * tot_trips_best,  # fallback: no sampling data
        .default = (unr_catch / tot_trips_spc) * tot_trips_best            # main: monthly CPUE × best trips
      ), 0))
            # mutate(raised = round(case_when(unr_catch == 0 & tot_trips_spc == 0 ~ cpd * tot_trips_best,
      #                           .default = (unr_catch/tot_trips_spc) * tot_trips_best), 0))

    raised_df_wide <- raised_df |>
      select(landing_site_name, months, raised) |>
      pivot_wider(id_cols = 1, names_from = "months", values_from = "raised")
    
    # Add result to raised list
    raised_list[[sp]] <- raised_df_wide
    
  }
}
```

<!-- **Unraised** tuna catches reported in ereporting tools are illustrated in the tables below, in kilograms (kgs). -->

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: data_skj
#| tbl-cap: !expr if ("skj" %in% names(unr_list)) glue::glue("Unraised skipjack catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("skj" %in% names(unr_list)) { 
    flextable(unr_list$skj|> 
                mutate(across(where(is.numeric), as.integer))) |>
    set_header_labels(landing_site_name = "Landing Site") |>
    colformat_num(j = 2:13, digits = 0) |>  
    width(j=1, width=1.2) |> 
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 17, unit = "cm")
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: df_yft_total
#| tbl-cap: !expr if ("yft" %in% names(unr_list)) glue::glue("Unraised yellowfin catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("yft" %in% names(unr_list)) { 
    flextable(unr_list$yft|> 
                mutate(across(where(is.numeric), as.integer))) |>
    set_header_labels(landing_site_name = "Landing Site") |>

    width(j=1, width=1.2) |> 
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 17, unit = "cm")
}

```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: df_bet_total
#| tbl-cap: !expr if ("bet" %in% names(unr_list)) glue::glue("Unraised bigeye catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("bet" %in% names(unr_list)) { 
    flextable(unr_list$bet|> 
                mutate(across(where(is.numeric), as.integer))) |>  
    set_header_labels(landing_site_name = "Landing Site") |>
    colformat_num(j = 2:13, digits = 0) |>  

    width(j=1, width=1.2) |> 
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 17, unit = "cm")
    }

```

<!-- **Raised** catches are produced using the average catch per trip, raised by the best estimate of trips, to produce a representative estimate of total artisanal tuna catches. It should be **noted** that if there are sites with no catches reported, these are removed from the totals, even if effort was indicated by the member. This should be revisited - and an error message is noted in this document when that occurs. -->

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: raise-skj
#| tbl-cap: !expr if ("skj" %in% names(raised_list)) glue::glue("Raised skipjack catch (kgs) reported in ereporting tools.") else ""

# Remove rows where no catch was reported - as no average exists...

if ("skj" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$skj 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_skj = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |>
    set_header_labels(landing_site_name = "Landing Site") |>

    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0

if (exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0) {
  cat("Missing skipjack catch with effort are reported below.\n\n")
}
```

```{r missing_skj, include=FALSE}
# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0){
  flextable(missing_unraised_skj |> select(-no_catch)) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

  colformat_num(j = 2:13, digits = 0) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: raise-yft
#| tbl-cap: !expr if ("yft" %in% names(raised_list)) glue::glue("Raised yellowfin catch (kgs) reported in ereporting tools.") else ""
# Remove rows where no catch was reported - as no average exists...

if ("yft" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$yft 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_yft = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |>
    set_header_labels(landing_site_name = "Landing Site") |>

    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0

if (exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0) {
  cat("Missing yellowfin catch with effort are reported below.\n\n")
}
```

```{r missing_yft, include=FALSE}

# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0){
  flextable(missing_unraised_yft |> select(-no_catch)) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

  colformat_num(j = 2:13, digits = 0) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: raise-bet
#| tbl-cap: !expr if ("bet" %in% names(raised_list)) glue::glue("Raised bigeye catch (kgs) reported in ereporting tools.") else ""

# Remove rows where no catch was reported - as no average exists...

if ("bet" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$bet 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_bet = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0

if (exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0) {
  cat("Missing bigeye catch with effort are reported below.\n\n")
}
```

```{r missing_bet, include=FALSE}
# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0){
  flextable(missing_unraised_bet|> select(-no_catch)) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

  colformat_num(j = 2:13, digits = 0) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

<!-- ## Final Estimates -->

<!-- Here we take the raised catch estimates for the three key tuna species and sum them across all landing sites and months to produce an annual estimate in metric tonnes (mt). -->

```{r final.ace}
#| label: tbl-raised-ace
#| tbl-cap: Raised and unraised annual artisanal catch estimates in mt

ace <- data.frame()

for (sp in species) {
  
  if (sp %in% names(unr_list)) {
    
    unraised <- unr_list[[sp]] |>
      data.frame() |>
      ungroup() |>
      select(-c(1, 14)) |>
      summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) |>
      rowSums()
    
    raised <- raised_list[[sp]] |>
      data.frame() |>
      ungroup() |>
      select(-1) |>
      summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) |>
      rowSums()
    
  } else {
    unraised <- 0
    raised   <- 0
  }
  
  ace <- bind_rows(ace, data.frame(
    sp       = toupper(sp),
    unraised = round(unraised / 1000, 2),
    raised   = round(raised / 1000, 2)
  ))
}

# Produce final output
ace_final = ace |>
  gather(key='Method', value='ACE', -1) |>
  spread(key = sp, value=ACE)

flextable(ace_final) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)

```

```{r perc.sp}
#| label: tbl-raised-ace-perc
#| include: true
#| tbl-cap: Percentage of the artisanal tuna ACE associated with each of the key species

ace_final$tot = rowSums(ace_final[,-1])
ace_final <- ace_final |>  mutate(SKJ = round(SKJ/tot*100,2),
                      YFT = round(YFT/tot*100,2),
                      BET = round(BET/tot*100,2)) |>
  mutate(across(BET:YFT, ~replace_na(.,0))) |>
  select(-tot)

flextable(ace_final) |>
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)

```

# Socio-economic factors

Whale watching is on the rise, with locals increasingly taking tourists out on charters during the June--September whale migration season.

# Disposal of catch

Sharks that are caught in Niue's waters must be discarded as per Niue's legislation and the National Pelagic Management plan.

# Onshore developments

The processing plant remains inactive, with no immediate plans for reactivation. Most current onshore development relates to access infrastructure, in particular the wharf upgrade. The upgrade aims to provide easier and safer offloading for fishing vessels, given the exposed nature of Niue's only port, an issue identified during the last operational review as a key area for improvement. A boat and canoe launching site on the southern side of the island has also been used to offload catch when the main port was too rough for normal operations.

# Future prospects of the fishery

The Pelagic Development and Management Plan provides the framework for guiding current and future prospects in the fishery, supported by long- and short-term partnerships. One key long-term partnership is with the Niue Ocean Wide (NOW) project, which is establishing management arrangements for Niue's newest and largest marine protected area. The MPA will cover significant inshore and offshore habitats and species, along with both fishing and non-fishing activities. Work with NOW is already underway to develop surveillance tools for monitoring parts of the MPA. Negotiations regarding Niue's High Seas allocation are also ongoing.

# Status of tuna fishery data collection systems

The process of data collection is enshrined in the Licensing terms and conditions of foreign vessels fishing in Niue waters. Reporting and data collection is mandatory with authorised Officers rights to set the terms, though these are based on data requirements of SPC and WCPFC. Penalties are imposed on denying or refusal of access to data. For Artisanal data, a data officer has been identified and appointed to recommence collecting of data 4 days per week. In some cases including Saturday to cover Fishing competitions held by various Communities. In previous years fishermen were relied upon to record their own data. Along with the offshore data, this data is entered and stored in the TUFMAN database developed by SPC.

# Research activities covering target and non-target species

Research on pelagic species is subject to funding availability and the priorities set out in the National Pelagic Management and Development Plan. In 2010, a wahoo tagging programme was conducted using satellite tags and the ARGOS tracking system to study wahoo movements. In 2019, tuna and shark tagging was carried out in collaboration with the NOW project. In 2016, Niue hosted Pristine Seas (National Geographic) to research pelagic fish assemblages using mid-water baited remote underwater video stations (BRUVS); at the same time, the Pacific Community conducted finfish and invertebrate surveys with Niue Fisheries staff. Further research is expected subject to funding availability.

\newpage

# Addendum

```{r loadaddendum}
#| output: asis
#| echo: false

cat(knitr::knit_child("template_addendum.qmd", quiet = TRUE))
```
