---
format: 
  pdf:
    geometry: "margin=2cm"
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"
    pdf-engine: xelatex
    latex-max-runs: 2
    keep-tex: true
    include-in-header:
      text: |
        \PassOptionsToPackage{quiet}{fontspec}
editor: source
toc: false
embed-resources: true 
execute: 
  echo: false
  message: false
  warning: false
  fig-width: 12
  fig-height: 8
params:
  year: 2025
  country_code: 'TK'
  data_provided: 'Yes'
  aceman: 'aceman'
  additional_sections: ''
---

```{r}
# Load necessary libraries
library(DBI)
library(odbc)
library(tidyverse)
library(flextable)
library(scatterpie)
library(mregions2)
library(countrycode)
source("utils.R")

sp_colours <- setNames(object = c("seagreen3","red3","black","cyan1","steelblue","orange",
                                  "pink","wheat","purple3","grey40","dodgerblue4","purple4",
                                  "#66CCFF","salmon4","magenta3","gold", "gray"),
                         nm = c("Albacore","Bigeye tuna","Black marlin","Blue marlin","Blue shark",
                                "Hammerhead sharks nei","Mako sharks","Oceanic whitetip shark",
                                "Pacific bluefin tuna","Silky shark","Skipjack tuna",
                                "Striped marlin","Swordfish","Thresher sharks nei",
                                "Whale shark","Yellowfin tuna", "Other"))
# adjust path as needed
flag_lookup <- read_csv("../ofp-dm-part1-report-template/config/flag_names.csv") |>
  select(country_code, country_name, country_short) |>
  mutate(country_name = na_if(country_name, "NULL"),
         country_short = na_if(country_short, "NULL")) |>
  # use short name when available, fall back to long
  mutate(name_pretty = coalesce(str_to_title(tolower(str_trim(country_short))), str_to_title(country_name)))

flag_labels <- flag_lookup |>
  mutate(flag_code_lower = tolower(country_code)) |>
  select(flag_code_lower, name_pretty) |>
  deframe()  # from tibble: turns 2-col df into a named vector

# Helper: get a country's 200NM EEZ as an sf object, keyed by ISO-2 code
get_country_eez <- function(iso2_code) {
  iso3 <- countrycode(iso2_code, "iso2c", "iso3c", warn = FALSE)
  
  if (is.na(iso3)) {
    stop("Could not resolve ISO-2 code '", iso2_code, "' to ISO-3.")
  }
  
  eez <- mrp_get(
    "eez",
    cql_filter = sprintf("iso_ter1 = '%s' AND pol_type = '200NM'", iso3)
  )
  
  if (nrow(eez) == 0) {
    warning("No EEZ found for '", iso2_code, "' (ISO-3: ", iso3, ").")
  }
  
  eez
}

# Helper: build a gridded species-composition catch summary for one EEZ
grid_catch <- function(data, eez_code, cell_size = 1,
                       species = c("skj", "yft", "bet")) {
  
  species_cols <- paste0(species, "_mt")
  half <- cell_size / 2
  
  data |>
    filter(eez == eez_code) |>
    mutate(
      lat_bin = floor(latd / cell_size) * cell_size + half,
      lon_bin = floor(lond / cell_size) * cell_size + half
    ) |>
    group_by(lon_bin, lat_bin) |>
    summarise(across(all_of(species_cols), \(x) sum(x, na.rm = TRUE)),
              .groups = "drop") |>
    rename_with(\(x) str_remove(x, "_mt$"), all_of(species_cols)) |>
    mutate(total = rowSums(across(all_of(species)))) |>
    filter(total > 0)
}

# Detect context: standalone render vs. knit_child from parent
# - Standalone: params$country_code exists (set via execute_params)
# - Child:      country_cd is already defined by the parent
if (!exists("country_cd") || length(country_cd) == 0 || !nzchar(country_cd)) {
  country_cd <- tolower(params$country_code)
}

if (!exists("report_year") || length(report_year) == 0) {
  report_year <- params$year
}

if (!exists("this_yr_folder") || length(this_yr_folder) == 0 || !nzchar(this_yr_folder)) {
  this_yr_folder <- paste0("./data/report_", as.character(report_year), "_", country_cd, "/") |>
    tolower()
}

# Sanity check — fail loudly if anything is still missing
stopifnot(
  length(country_cd)     == 1, nzchar(country_cd),
  length(report_year)    == 1,
  length(this_yr_folder) == 1, nzchar(this_yr_folder)
)

# Define years to use for tables and figures
yrs_long <- (report_year - 4):report_year
yr_range <- paste(min(yrs_long), max(yrs_long), sep = "-")
'%nin%' <- function(x, y) !x %in% y

country_name <- flag_labels[[country_cd]]
```

```{r data-prep}
# Load all raw data - from t2
data_3615 <- read_and_clean(this_yr_folder, country_cd, r_code = 3615) |>
      mutate(across(where(is.character), tolower))

data_3527 <- read_and_clean(this_yr_folder, country_cd, r_code = 3527)|>
      mutate(across(where(is.character), tolower))

data_3614 <- read_and_clean(this_yr_folder, country_cd, r_code = 3614)|>
      mutate(across(where(is.character), tolower))

data_2900 <- read_and_clean(this_yr_folder, country_cd, r_code = 2900)|>
      mutate(across(where(is.character), tolower)) |>
  filter(!(flag == country_cd))  # PS catch EEZ per country

data_2894<- read_and_clean(this_yr_folder, country_cd, r_code = 2894)|>
      mutate(across(where(is.character), tolower))|>
      filter(!(flag == country_cd))# LL catch EEZ (I think…)

data_2904<- read_and_clean(this_yr_folder, country_cd, r_code = 2904)|>
      mutate(across(where(is.character), tolower)) |>
      filter(!(flag == country_cd))  # PS catch spatial EEZ

data_3375<- read_and_clean(this_yr_folder, country_cd, r_code = 3375)|>
      mutate(across(where(is.character), tolower))|>
      filter(!(flag == country_cd)) # LL catch spatial EEZ

# FIX FOR TO
if(nrow(data_3375) > 0 && country_cd == "to"){
  # there are some issues with WS vessels positions - they are wrongly assigning catches in TO EEZ
  data_3375 <- data_3375 |> 
    filter(!(flag == "ws"))
}


if(nrow(data_2904) > 0){
  data_2904 <- data_2904 |>
    rename(year = t2col0) |>
    filter(year == report_year) 
}

if (nrow(data_2900) > 0){
  data_2900 <- data_2900 |>
    mutate(year = as.numeric(substr(t2col1, 1, 4))) |>
    filter(year >= report_year - 4) |>
    filter(year <= report_year) |>
    select(-fish_days_aws_only, - t2col1)
}

if (nrow(data_2894) > 0){
  data_2894 <- data_2894 |>
    mutate(year = as.numeric(substr(t2col4, 1, 4))) |>
    filter(year >= report_year - 4) |>
    filter(year <= report_year) |>
    select(flag, year, vessels = vessel_name, trips, days, days, alb_mt, bet_mt, yft_mt, oth_mt, tot_mt)
}

# FIX FOR NR
if(nrow(data_2894) > 0 && country_cd == "nr"){
  # some LL catch showing for NR but these must be errors in reported positions
  if (sum(data_2894$tot_mt) < 4) {
    data_2894 <- data.frame()
  }
  
}

ika_trips_file = paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")

if (file.exists(ika_trips_file)){
  data_ika_raw <- tryCatch(
    read.csv(paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")),
    error = function(e) NULL
  )
  
  if (exists("data_ika_raw")){
    if ("LandingSite" %in% colnames(data_ika_raw) && nrow(data_ika_raw) > 0){
        data_ika_raw <- data_ika_raw|>
          rename(landing_site_name = LandingSite) |>
          group_by(Country, landing_site_name, Year, Month) |>  
          summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop") |>
          filter(Year == report_year)
    }
    
  }
  
  if (is.null(data_ika_raw) || nrow(data_ika_raw) == 0 || is.null(data_ika_raw)) rm(data_ika_raw)
  }

if (!exists("data_ika_raw") && nrow(data_3615) == 0) {
  cat("No Ikasavea or Tails trip data found for this country/year. Document generation stopped.\n")
  knitr::knit_exit()
}

if (exists("data_ika_raw")){
  
   data_ika_trips <- data_ika_raw |>
    mutate(across(where(is.character), tolower)) |>
    janitor::clean_names() |>
    select(landing_site_name, month,
           number_trips)
  
  df_ika_wide <- data_ika_trips %>%
    complete(landing_site_name, month = 1:12, fill = list(number_trips = 0)) %>%
    mutate(month = month.abb[month]) %>%
    pivot_wider(names_from = month, values_from = number_trips) %>%
    select(landing_site_name, all_of(month.abb))
  
  data_ika_catch_kg <- data_ika_raw |>
    mutate(across(where(is.character), tolower)) |>
    janitor::clean_names() |>
    select(landing_site_name, month, skj = skj_kg, yft = yft_kg, bet = bet_kg)#, alb = alb_kg)
  
  data_ika_catch_kg <- data_ika_catch_kg |>
    pivot_longer(cols = c("skj", "bet", "yft"), names_to = "sp_code", values_to = "sp_kg")
  
  }

both_exist <- exists('data_ika_trips') && nrow(data_ika_trips) > 0 && exists('data_3615') && 
              is.data.frame(get('data_3615')) && nrow(data_3615) > 0

local_knowl_trips_file <-  str_c(this_yr_folder,  'additional_files/est_trips.csv')

```

\newpage

# Abstract

The commercial fishery in Tokelau is comprised of purse seine and longline vessels operating under various multilateral, regional and bilateral agreements. In 2025, there was no fishing activities recorded from longline vessels. For purse seine, it was observed a notable drop on the total catch (190.83 mt) in comparison to four-year average of 6,680.125 mt. Both fisheries are managed in full compliance with the vessel day schemes (VDS) established under the Palau Arrangement.

Coverage of Tokelau artisanal data continues to improve since the introduction of the SPC TAILS software. The total unraised tuna catch reported from this fishery in 2025 was 17.37 metric tons (mt) with skipjack (64%) and yellowfin (35%) tunas making up the majority of the catch composition.


# Background

Tokelau does not have a commercial fishing fleet however, it has a large artisanal fleet of about 200 small 10’ to 16’ motorised aluminium boats powered by 15-40HP outboard motors. These artisanal fishing boats operate within the 12nm Territorial Sea (TS), but generally do not venture out further than 5nm from shore. For food security purposes, Tokelau has also established a buffer zone (closed area) of 13nm extending from the 12nm TS closed to any foreign fishing.

Transhipment is permitted to occur in the Tokelau Exclusive Zone (EEZ), TS and Internal Waters where there are exceptional circumstances and prior approval provided. Tokelau did not approve any tranship-ment activity in her waters in 2025.

\newpage

# Coastal state reporting

Since 2012, `r country_name` has participated in the PNA purse seine vessel day scheme (PS VDS) established under the Parties to the Palau Arrangement, with an EEZ limit of 1,000 fishing days for the purse seine fishery. In 2016, `r country_name` also joined the PNA longline VDS and was allocated 5,000 fishing days, commencing the sale of these days the following year.

Offshore commercial fishing is undertaken by foreign vessels from distant water fishing nations and neighbouring Pacific Island States. In `r country_name` no new vessel was licensed to fish `r country_name` waters.

@tbl-data-eez-all-flags presents a catch summary by fishing gear over the last five years. Albacore is the most important species in the longline catch, with moderate catches of yellowfin and bigeye also taken. Longline activity in the `r country_name` EEZ has varied considerably over recent years: catches in some years were almost negligible. Longline activity was not recorded in the `r country_name` EEZ in `r report_year`. The share of catch by flag state for all previous 5 years is shown in @fig-data-eez-all-flags-perc.

Purse seine fishing has likewise been variable (@tbl-data-eez-all-flags), although effort in `r report_year` was substantially lower (16 fishing days) than the previous five-year average of around 150 fishing days. The fleet composition over this period is shown in @fig-data-eez-all-flags-perc, with vessels flagged to the United States and Tuvalu accounting for the vast majority of effort in `r report_year`. Annual catches for `r report_year` (@fig-data-eez-all-flags-hist) indicate that skipjack was the most commonly caught species across all flags.

```{r data-eez-all-flags}

if (nrow(data_2900) > 0){
  
  # Table - overall per year
  data_2900_summary <- data_2900 |>
    group_by(year) |>
    summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) |>
    mutate(across(year:fish_days, ~ round(.x, 0)),
           gear = "Purse seine") |>
    select(-sea_days)
  
  plot_data1ps <- data_2900 |>
    mutate(tot_mt = replace_na(tot_mt, 0)) |>          
    mutate(flag = fct_lump_n(flag, n = 5, w = tot_mt, other_level = "Other")) |>
    group_by(year, flag) |>
    summarise(tot_mt = sum(tot_mt, na.rm = TRUE), .groups = "drop") |>
    mutate(flag = fct_reorder(flag, tot_mt, .fun = sum, .desc = TRUE),
           gear = "Purse seine")
  
  # Country and sp catch 
  plot_data2ps <- data_2900 |>
    filter(year == report_year) |>
    mutate(across(c(skj_mt, bet_mt, yft_mt, oth_mt), \(x) replace_na(x, 0))) |>
    pivot_longer(c(skj_mt, bet_mt, yft_mt, oth_mt),
                 names_to = "species", values_to = "mt") |>
    mutate(species = recode(species,
                            skj_mt = "Skipjack tuna",
                            bet_mt = "Bigeye tuna",
                            yft_mt = "Yellowfin tuna",
                            oth_mt = "Other")) |>
    mutate(flag    = fct_reorder(flag, mt, .fun = sum, .desc = TRUE),
           species = fct_reorder(species, mt, .fun = sum, .desc = TRUE),
           gear = "Purse seine")
}

if (nrow(data_2894) > 0){
  
  # Table - overall per year
  data_2894_summary <- data_2894 |>
    group_by(year) |>
    summarise(vessels = n_distinct(vessels), 
              across(where(is.numeric), ~ sum(.x, na.rm = TRUE))
              ) |>
    mutate(across(year:days, ~ round(.x, 0)),
           gear = "Longline") |>
    rename(fish_days = days)
  
  plot_data1ll <- data_2894 |>
    mutate(tot_mt = replace_na(tot_mt, 0)) |>          
    mutate(flag = fct_lump_n(flag, n = 5, w = tot_mt, other_level = "Other")) |>
    group_by(year, flag) |>
    summarise(tot_mt = sum(tot_mt, na.rm = TRUE), .groups = "drop") |>
    mutate(flag = fct_reorder(flag, tot_mt, .fun = sum, .desc = TRUE),
           gear = "Longline")
  
  # Country and sp catch 
  plot_data2ll <- data_2894 |>
    filter(year == 2025) |>
    mutate(across(c(alb_mt, bet_mt, yft_mt, oth_mt), \(x) replace_na(x, 0))) |>
    pivot_longer(c(alb_mt, bet_mt, yft_mt, oth_mt),
                 names_to = "species", values_to = "mt") |>
    mutate(species = recode(species,
                            alb_mt = "Albacore tuna",
                            bet_mt = "Bigeye tuna",
                            oth_mt = "Other")) |>
    mutate(flag    = fct_reorder(flag, mt, .fun = sum, .desc = TRUE),
           species = fct_reorder(species, mt, .fun = sum, .desc = TRUE),
           gear = "Longline")

}

has_2894 <- exists("data_2894_summary")
has_2900 <- exists("data_2900_summary")

if (has_2894 && has_2900) {
  flag_summary <- bind_rows(data_2894_summary, data_2900_summary)
  plot_data1   <- rbind(plot_data1ll, plot_data1ps)
  plot_data2   <- rbind(plot_data2ll, plot_data2ps)
} else if (has_2894) {
  flag_summary <- data_2894_summary   # longline only
  plot_data1   <- plot_data1ll
  plot_data2   <- plot_data2ll
} else if (has_2900) {
  flag_summary <- data_2900_summary   # purse seine only
  plot_data1   <- plot_data1ps
  plot_data2   <- plot_data2ps
}else{
  flag_summary <- data.frame()
  plot_data1   <- data.frame()
  plot_data2   <- data.frame()
}


```

```{r}
#| echo: false
#| include: true
#| results: 'asis'
#| label: tbl-data-eez-all-flags
#| tbl-cap: !expr if(nrow(flag_summary) > 0) glue::glue("Annual catches in {country_name} EEZ for the last 5 years.") else ""

if (nrow(flag_summary) > 0) {
  flag_summary |>
  relocate(oth_mt, tot_mt, .after = last_col()) |>
  mutate(year = as.integer(year)) |>
  as_grouped_data(groups = "gear") |>
  flextable() |>
  set_header_labels(
    gear = "Gear", year = "Year", vessels = "Vessels", trips = "Trips",
    fish_days = "Days", alb_mt = "Alb (mt)", bet_mt = "Bet (mt)",
    yft_mt = "Yft (mt)", skj_mt = "Skj (mt)", oth_mt = "Others (mt)",
    tot_mt = "Total (mt)"
  ) |>
  colformat_double(j = c("alb_mt","bet_mt","yft_mt","skj_mt","oth_mt","tot_mt"),
                   digits = 0, na_str = "–") |>
  colformat_int(j = c("year","vessels","trips","fish_days")) |>
  bold(i = ~ !is.na(gear)) |>
  fontsize(size = 8, part = "all") |>
  fit_to_width(max_width = 6.5) |>
  theme_booktabs() |>
  align(align = "right", part = "all") |>
  align(j = "gear", align = "left", part = "all")
}
```

```{r}
#| echo: false
#| include: true
#| results: 'asis'
#| label: fig-data-eez-all-flags-perc
#| fig-cap: !expr if(nrow(plot_data1) > 0) glue::glue("Percentage of total catch (all species) in the {country_name} EEZ, by flag during the last 5 years.") else ""

if (nrow(plot_data1) > 0){
  # Percentage -  which country fished more in the last 5 years
  ggplot(plot_data1, aes(x = factor(year), y = tot_mt, fill = flag)) +
    geom_col(position = "fill", colour = "white", linewidth = 0.3) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_hue(labels = flag_labels, c = 50) +
    facet_wrap(~ gear) +
    labs(x = NULL, y = "Share of catch", fill = "Country") +
    theme_minimal() +
    theme(legend.position = "bottom")
  }
```

```{r}
#| echo: false
#| include: true
#| results: 'asis'
#| label: fig-data-eez-all-flags-hist
#| fig-cap: !expr if(nrow(plot_data2) > 0) glue::glue("Annual catch of tuna in the {country_name} EEZ, by flag during {report_year}.") else ""

if (nrow(plot_data2) > 0){
   ggplot(plot_data2, aes(x = flag, y = mt, fill = species)) +
    geom_col() +
    scale_x_discrete(labels = flag_labels) +  
    coord_flip() +
    scale_y_continuous(labels = scales::comma) +
    scale_fill_manual(values = sp_colours, na.value = "grey80") +
    facet_wrap(~ gear) +
    labs(x = NULL, y = "Catch (mt)", fill = "Species") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom",
          panel.grid.major.x = element_blank())
  }
```

\newpage

@fig-data-eez-ps-spatial shows the distribution of purse seine catch, and its species composition within the EEZ and the adjacent area in `r report_year`. This shows that catch is highest in the northern parts of the EEZ, and this pattern is consistent with the results presented in previous Part reports. Skipjack was the species most commonly caught in all spatial cells followed by Yellowfin tuna which also corroborates with the overall `r report_year` catch composition (@fig-data-eez-all-flags-perc) and results from previous years.

```{r}
#| echo: false
#| include: true
#| results: 'asis'
#| label: fig-data-eez-ps-spatial
#| fig-cap: !expr if(nrow(data_2904) > 0) glue::glue("Catch of the three main tuna species by purse  seine  vessels  fishing  in  the  {country_name}  EEZ,  in  each  1x1  degree  cell  in  the EEZ and  the immediate surrounding area, over {report_year}.") else ""


if (nrow(data_2904) > 0) {
  
  eez  <- get_country_eez(country_cd)
  land <- ne_countries(scale = "medium", returnclass = "sf")
  
  # Species lookup: short code → full name (must match sp_colours keys)
  species_map <- c(
    skj = "Skipjack tuna",
    yft = "Yellowfin tuna",
    bet = "Bigeye tuna"
  )
  
  catch_grid <- grid_catch(data_2904,
                           eez_code  = tolower(country_cd),
                           cell_size = 1,
                           species   = names(species_map))
  
  max_r <- 0.45
  catch_grid <- catch_grid |>
    mutate(radius = max_r * sqrt(total / max(total)))
  
  # Build fill values and labels from the lookup
  fill_values <- setNames(sp_colours[species_map], names(species_map))
  fill_labels <- setNames(species_map, names(species_map))
  
  bbox <- st_bbox(eez)
  
  ggplot() +
    geom_sf(data = land, fill = "grey90", colour = "grey60", linewidth = 0.2) +
    geom_sf(data = eez,  fill = "grey50", colour = "black", alpha = 0.4) +
    geom_scatterpie(
      data   = catch_grid,
      aes(x = lon_bin, y = lat_bin, r = radius),
      cols   = names(species_map),
      colour = NA,
      alpha  = 0.9
    ) +
    scale_fill_manual(
      values = fill_values,
      labels = fill_labels,
      name   = "Species"
    ) +
    coord_sf(
      xlim   = c(bbox["xmin"], bbox["xmax"]),
      ylim   = c(bbox["ymin"], bbox["ymax"]),
      expand = FALSE
    ) +
    labs(x = "Longitude", y = "Latitude") +
    theme_bw() +
    theme(legend.position = "bottom",,
          panel.background = element_rect(fill = "white"))
}

```

\newpage

# Flag state reporting

Although `r country_name` is not a flag state, this section presents annual catch estimates derived from artisanal fishery data. To improve the estimate of annual production, catch landings recorded in *TAILS* are raised using estimated fishing activity --- the number of boats and canoes estimated to fish daily at each landing site. @tbl-raised-ace presents the raised and unraised estimates per species for 2025, and @tbl-raised-ace-perc shows the percentage contribution of each species.

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: Monthly-effort
#| tbl-cap: !expr if(nrow(data_3615) > 0) glue::glue("Number of trips per month and landing site reported in Tails") else ""

# Check if data exists
if (!(nrow(data_3615) == 0)) {
  
  names(data_3615) <- gsub(x = names(data_3615), pattern = "_trips$", replacement = "")  
  
  data_3615 <- data_3615 |>
    rename_with(~ month.abb[match(., tolower(month.abb))], 
              .cols = any_of(tolower(month.abb))) |> 
    select(-entity)
  
  flextable(data_3615) |>  
  set_header_labels(landing_site_name = "Landing Site") |>
  width(j=1, width=1.2) |> 
  fontsize(size = 8, part = "all") |>
  fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: Monthly-effort-ika
#| tbl-cap: !expr if(exists('data_ika_trips')) glue::glue("Number of trips per month and landing site reported in Ikasavea") else ""
if (exists('data_ika_trips')){
  flextable(df_ika_wide) |> 
  set_header_labels(landing_site_name = "Landing Site") |>

  width(j=1, width=1.2) |> 
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
  }
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr both_exist

if (both_exist) {
  cat("### Differences in Landing Sites: Ikasavea + Tails\n\n")
  cat("Artisanal data is collected using Tails and in Ikasavea platforms. The summary table below of all landing sites common to T2 and Ikasavea as well the ones unique to each of these sources (if any).")
}
```

```{r}
#| echo: false
#| include: false
#| eval: !expr both_exist
#| results: 'asis'
#| label: diff_t2_ika

if (both_exist) {

  ls_t2  <- tolower(trimws(unique(data_3615$landing_site_name)))
  ls_ika <- tolower(trimws(unique(df_ika_wide$landing_site_name)))
  
  in_both    <- intersect(ls_t2, ls_ika)
  only_t2    <- setdiff(ls_t2, ls_ika)
  only_ika   <- setdiff(ls_ika, ls_t2)
  
  if (length(only_t2) == 0 && length(only_ika) == 0) {
    cat("✅ All landing site names match perfectly between both sources.\n")
  } else {
    
    # Summary table
    summary_df <- data.frame(
      ` ` = c("Matched in both", "Only in T2 (data_3615)", "Only in Ikasavea"),
      `Count` = c(length(in_both), length(only_t2), length(only_ika)),
      check.names = FALSE
    )
    
    # cat("### Landing site name comparison\n\n")
      flextable(summary_df) |>  
        set_header_labels(landing_site_name = "Landing Site") |>
         width(j=1, width=1.2) |> 
        fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)

    
    # Table of mismatches
    if (length(only_t2) > 0 || length(only_ika) > 0) {
      max_len <- max(length(only_t2), length(only_ika))
      mismatch_df <- data.frame(
        `Only in T2 (data_3615)` = c(only_t2,  rep("", max_len - length(only_t2))),
        `Only in Ikasavea`       = c(only_ika, rep("", max_len - length(only_ika))),
        `In both` = c(in_both, rep("", max_len - length(in_both))),
        check.names = FALSE
      )
      # cat("\n### Unmatched landing site names\n\n")
      flextable(mismatch_df) |>  
        set_header_labels(landing_site_name = "Landing Site") |>
         width(j=1, width=1.2) |> 
        fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
    }
  }
}

```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr both_exist

if (both_exist) {
  cat("### Trips: Ikasavea + Tails\n\n")
  cat("The final number of trips combining what was reported on these two platforms is presented in the table below.")
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: tails_ika_summary

# ika and tails combined
if (both_exist) {
  prep <- function(df) {
    df %>%
      mutate(landing_site_name = tolower(trimws(landing_site_name))) %>%
      pivot_longer(-landing_site_name, names_to = "month", values_to = "trips") |>
      mutate(month = tolower(gsub("_trips", "", month)))
  }
  
  combined_t2_ika <- bind_rows(prep(data_3615), prep(df_ika_wide)) %>%
    group_by(landing_site_name, month) %>%
    summarise(trips = sum(trips), .groups = "drop") %>%
    mutate(month = factor(month, levels = tolower(month.abb))) %>%
    pivot_wider(names_from = month, values_from = trips) %>%
    arrange(landing_site_name) |>
    select(landing_site_name, tolower(month.abb))
  
  ft <- flextable(combined_t2_ika) |>  

    width(j=1, width=1.2) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
  
  # flextable::htmltools_value(ft)
  ft

} else {
  combined_t2_ika <- data.frame()
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr nrow(data_3615) > 0 && file.exists(local_knowl_trips_file)

if (nrow(data_3615) > 0 && file.exists(local_knowl_trips_file)) {
  cat("### Trips: Local Knowledge\n\n")
  cat("Member-estimated **trips** per month at each landing site are presented in the table below. These estimates are used to compare with what is reported in the ereportng tools to produce a 'best' estimate of effort.")
}
```

```{r trips-est-mem}
#| echo: false
#| include: false
#| results: 'asis'
#| label: mem-trips
#| tbl-cap: !expr if(nrow(data_3615) > 0) glue::glue("Member reported estimated trips per month per landing site") else ""

if (file.exists(local_knowl_trips_file)){
  local_knowl_trips = read.csv(local_knowl_trips_file)
  flextable(local_knowl_trips)  |>  
    fontsize(size = 8, part = "all") |> 
    width(j=1, width=2) |>
        fit_to_width(max_width = 6.5)

}

```

<!-- ### Best estimate of fishing trips -->

<!-- Estimate monthly trips are based on the **max** *value either reported by the member or reported in the ereporting tools Tails and Ikasavea*. In case ereporting data is collected through Tails and Ikasavea tools the combination of trips in both sources will be used as your best estimate monthly trips (tab above). -->

```{r prep-ereporting_trips, echo = F}

# check logic to be applied for ereporting: tails, ikasavea or tails + ikasavea?

if (nrow(combined_t2_ika)>0){ # it means tails + ikasavea
  
  ereporting_trips <- combined_t2_ika |>
    mutate(landing_site_name = str_squish(landing_site_name)) |>
    data.frame()
  ereporting_trips_tag <- c("Trips: Ikasavea + Tails")
  
  }else if (!exists("data_ika_trips") && nrow(data_3615) > 0){ # so it means you only have tails
    ereporting_trips <- data_3615 |>
      mutate(landing_site_name = str_squish(landing_site_name)) |>
      data.frame()
    ereporting_trips_tag <- c("Trips: Tails")
      
  }else if(nrow(data_3615) == 0 && nrow(data_ika_trips) > 0){ # it means you only have ikasavea 
    ereporting_trips <- df_ika_wide |>
      data.frame()
    ereporting_trips_tag <- c("Trips: Ikasavea")
  }else{
    ereporting_trips <- data.frame()
    ereporting_trips_tag <- c("")
  }
```

```{r max.trips}
#| echo: false
#| include: false
#| results: 'asis'
#| label: est-trips-monthly
#| tbl-cap: !expr if(nrow(ereporting_trips) > 0) glue::glue("Best estimate of trips per month at each landing site") else ""
 
if (nrow(ereporting_trips) > 0){
  
  if (file.exists(local_knowl_trips_file)){
    
    # check if columns are correct, if not, flag the issue
    if (!all(c("landing_site_name", "est.trips") %in% colnames(local_knowl_trips))) {
      cat("Your estimated trip file has to have the following columns: 'landing_site_name', 'est.trips'")
      best_estimate_trips <- ereporting_trips
        
    }else{
      
      # if all columns are in the df, groom local_knowl_trips and produce best_estimate_trips
      local_knowl_trips <- local_knowl_trips |>
        mutate(across(where(is.character), tolower)) |>
        mutate(landing_site_name = str_squish(landing_site_name))

      best_estimate_trips = ereporting_trips |> 
        full_join(local_knowl_trips |> 
                    select(landing_site_name, est.trips))
      
      best_estimate_trips[is.na(best_estimate_trips)] <- 0
      best_estimate_trips <- data.frame(best_estimate_trips)
      for( i in 2:13){
        best_estimate_trips[ ,i] = ifelse( best_estimate_trips[ ,i] > best_estimate_trips[ ,14],  
                                           best_estimate_trips[ ,i] , best_estimate_trips[ ,14] )
        }

      # Get rid of sites that only have NAs
      best_estimate_trips$num_na_cols <- rowSums(best_estimate_trips[, 2:13] )

      best_estimate_trips <- best_estimate_trips |>  filter(num_na_cols>0) |>
        select(-c(est.trips, num_na_cols))
    }
    }
  else{
    cat(paste0("There is no member-estimated trips available, so the best estimate fishing is the table above: <b>", ereporting_trips_tag,
             "</b>"))    
    best_estimate_trips <- ereporting_trips
  }
  }

flextable(best_estimate_trips)  |>    
  set_header_labels(landing_site_name = "Landing Site") |>

  fontsize(size = 8, part = "all") |>
  fit_to_width(max_width = 6.5)

```

<!-- ## Catch Estimates -->

```{r prep-catch-data}
# ── Month number → abbreviated name (keeps columns in calendar order) ─────────
month_labels <- c("Jan","Feb","Mar","Apr","May","Jun",
                  "Jul","Aug","Sep","Oct","Nov","Dec")

# ── Helper: pivot one subset to wide format ───────────────────────────────────
make_wide <- function(data) {
  data %>%
    # Sum sp_kg for each landing_site × month combination
    group_by(landing_site_name, month) %>%
    summarise(sp_kg = sum(sp_kg, na.rm = TRUE), .groups = "drop") %>%
    # Convert month numbers to abbreviated names
    mutate(month = month_labels[month]) %>%
    # Ensure all 12 months are present before pivoting
    complete(
      landing_site_name,
      month = month_labels,
      fill  = list(sp_kg = 0)
    ) %>%
    # Pivot to wide
    pivot_wider(
      names_from  = month,
      values_from = sp_kg,
      values_fill = 0          # missing combos → 0
    ) %>%
    # Reorder columns so months appear Jan → Dec (only those present in data)
    select(landing_site_name, any_of(month_labels))
}

# helper to combine ika and tails
combine_wide <- function(df1, df2) {
  if (is.null(df1) && is.null(df2)) return(NULL)
  
  month_cols <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
  
  drop_zero_rows <- function(df) {
    if (is.null(df)) return(NULL)
    df |> filter(if_any(all_of(month_cols), ~ . != 0))
  }
  
  df1 <- drop_zero_rows(df1)
  df2 <- drop_zero_rows(df2)
  
  if (is.null(df1) || nrow(df1) == 0) return(df2)
  if (is.null(df2) || nrow(df2) == 0) return(df1)
  
  full_join(df1, df2, by = "landing_site_name", suffix = c("_a", "_b")) |>
    mutate(across(all_of(paste0(month_cols, "_a")), 
                  ~ coalesce(., 0L))) |>
    mutate(across(all_of(paste0(month_cols, "_b")), 
                  ~ coalesce(., 0L))) |>
    mutate(Jan = Jan_a + Jan_b, Feb = Feb_a + Feb_b,
           Mar = Mar_a + Mar_b, Apr = Apr_a + Apr_b,
           May = May_a + May_b, Jun = Jun_a + Jun_b,
           Jul = Jul_a + Jul_b, Aug = Aug_a + Aug_b,
           Sep = Sep_a + Sep_b, Oct = Oct_a + Oct_b,
           Nov = Nov_a + Nov_b, Dec = Dec_a + Dec_b) |>
    select(landing_site_name, all_of(month_cols))
}

# Check if tails data exists - if yes, reshape and create a df per sp
if (nrow(data_3614) > 0) {
  df_by_species <- data_3614 |>
    mutate(landing_site_name = str_squish(landing_site_name)) |>
    group_by(sp_code) %>%
    group_map(~ make_wide(.x), .keep = FALSE) %>%
    setNames(sort(unique(data_3614$sp_code))) 
  
  df_bet <- df_by_species[["bet"]]
  df_skj <- df_by_species[["skj"]]
  df_yft <- df_by_species[["yft"]]
}

# Check if ika data exists - if yeas, reshape and create a df per sp
if(exists('data_ika_catch_kg') && is.data.frame(get('data_ika_catch_kg'))){
  
  df_by_species_ika <- data_ika_catch_kg %>%
    mutate(landing_site_name = str_squish(landing_site_name)) |>

    group_by(sp_code) %>%
    group_map(~ make_wide(.x), .keep = FALSE) %>%
    setNames(sort(unique(data_ika_catch_kg$sp_code))) 
  
  df_bet_ika <- df_by_species_ika[["bet"]]
  df_skj_ika <- df_by_species_ika[["skj"]]
  df_yft_ika <- df_by_species_ika[["yft"]]
  
  
  # If required (if ika and tails exist), combine catch values from both sources
  if (nrow(data_3614) > 0){
    df_bet_total <- combine_wide(df_bet, df_bet_ika)
    df_skj_total <- combine_wide(df_skj, df_skj_ika)
    df_yft_total <- combine_wide(df_yft, df_yft_ika)
  } else{
    # Just use the ika source if t2 data doesn't exist
    df_bet_total <- df_bet_ika
    df_skj_total <- df_skj_ika
    df_yft_total <- df_yft_ika
    
  }
} else {
  # Just use catch from tails if ika data doesn't exist
  df_bet_total <- df_bet
  df_skj_total <- df_skj
  df_yft_total <- df_yft
}
```

```{r prep-unr-data}
# Calculate total effort for the year by landing site using the ereporting trips
tot_trips_ereport = data.frame(landing_site_name = ereporting_trips$landing_site_name,
                    tot_trips = rowSums(ereporting_trips[,-1],na.rm=T)) |>
                    mutate(landing_site_name = str_squish(landing_site_name))
# Reshape trips
best_estimate_trips_long <- best_estimate_trips |>
  janitor::clean_names() |>
  mutate(landing_site_name = str_squish(landing_site_name)) |>
  pivot_longer(cols = 2:13, names_to = "months", values_to = "tot_trips_best")

ereporting_trips_long <- ereporting_trips |>
  janitor::clean_names() |>
  mutate(landing_site_name = str_squish(landing_site_name)) |>
  pivot_longer(cols = 2:13, names_to = "months", values_to = "tot_trips_spc")

# Check if data exists
species = c("skj", "bet", "yft")

unr_list <- list()
raised_list <- list()
for (sp in species){
  df_name <- paste0("df_", sp, "_total")
  if (exists(df_name) && is.data.frame(get(df_name)) && nrow(get(df_name)) > 0) {
    df <- get(df_name)
    # Now calculate monthly average catch per trip
    avg_sp = data.frame(landing_site_name = df$landing_site_name,
                    tot_catch = rowSums(df[,-1],na.rm=T)) |>
                  left_join(tot_trips_ereport) |>
      mutate(cpd = round(tot_catch/tot_trips,0))
    
    unr_sp <- df  |> 
      left_join(avg_sp |> 
                  select(landing_site_name, 
                         cpd))
    
    # Add result to unraised list
    unr_list[[sp]] <- unr_sp # 
    
    # Calculate raised
    unr_sp_long <- unr_sp |>
      janitor::clean_names() |>
      mutate(landing_site_name = str_squish(landing_site_name)) |>
      pivot_longer(cols = 2:13, names_to = "months", values_to = "unr_catch")
  
    raised_df <- unr_sp_long |>
      left_join(ereporting_trips_long, by = c("landing_site_name", "months")) |>
      left_join(best_estimate_trips_long, by = c("landing_site_name", "months")) |>
      mutate(raised = round(case_when(
        is.na(tot_trips_spc) | tot_trips_spc == 0 ~ cpd * tot_trips_best,  # fallback: no sampling data
        .default = (unr_catch / tot_trips_spc) * tot_trips_best            # main: monthly CPUE × best trips
      ), 0))
            # mutate(raised = round(case_when(unr_catch == 0 & tot_trips_spc == 0 ~ cpd * tot_trips_best,
      #                           .default = (unr_catch/tot_trips_spc) * tot_trips_best), 0))

    raised_df_wide <- raised_df |>
      select(landing_site_name, months, raised) |>
      pivot_wider(id_cols = 1, names_from = "months", values_from = "raised")
    
    # Add result to raised list
    raised_list[[sp]] <- raised_df_wide
    
  }
}
```

<!-- **Unraised** tuna catches reported in ereporting tools are illustrated in the tables below, in kilograms (kgs). -->

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: data_skj
#| tbl-cap: !expr if ("skj" %in% names(unr_list)) glue::glue("Unraised skipjack catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("skj" %in% names(unr_list)) { 
    flextable(unr_list$skj|> 
                mutate(across(where(is.numeric), as.integer))) |>
    set_header_labels(landing_site_name = "Landing Site") |>
    colformat_num(j = 2:13, digits = 0) |>  
    width(j=1, width=1.2) |> 
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 17, unit = "cm")
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: df_yft_total
#| tbl-cap: !expr if ("yft" %in% names(unr_list)) glue::glue("Unraised yellowfin catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("yft" %in% names(unr_list)) { 
    flextable(unr_list$yft|> 
                mutate(across(where(is.numeric), as.integer))) |>
    set_header_labels(landing_site_name = "Landing Site") |>

    width(j=1, width=1.2) |> 
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 17, unit = "cm")
}

```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: df_bet_total
#| tbl-cap: !expr if ("bet" %in% names(unr_list)) glue::glue("Unraised bigeye catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("bet" %in% names(unr_list)) { 
    flextable(unr_list$bet|> 
                mutate(across(where(is.numeric), as.integer))) |>  
    set_header_labels(landing_site_name = "Landing Site") |>
    colformat_num(j = 2:13, digits = 0) |>  

    width(j=1, width=1.2) |> 
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 17, unit = "cm")
    }

```

<!-- **Raised** catches are produced using the average catch per trip, raised by the best estimate of trips, to produce a representative estimate of total artisanal tuna catches. It should be **noted** that if there are sites with no catches reported, these are removed from the totals, even if effort was indicated by the member. This should be revisited - and an error message is noted in this document when that occurs. -->

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: raise-skj
#| tbl-cap: !expr if ("skj" %in% names(raised_list)) glue::glue("Raised skipjack catch (kgs) reported in ereporting tools.") else ""

# Remove rows where no catch was reported - as no average exists...

if ("skj" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$skj 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_skj = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |>
    set_header_labels(landing_site_name = "Landing Site") |>

    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0

if (exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0) {
  cat("Missing skipjack catch with effort are reported below.\n\n")
}
```

```{r missing_skj, include=FALSE}
# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0){
  flextable(missing_unraised_skj |> select(-no_catch)) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

  colformat_num(j = 2:13, digits = 0) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: raise-yft
#| tbl-cap: !expr if ("yft" %in% names(raised_list)) glue::glue("Raised yellowfin catch (kgs) reported in ereporting tools.") else ""
# Remove rows where no catch was reported - as no average exists...

if ("yft" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$yft 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_yft = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |>
    set_header_labels(landing_site_name = "Landing Site") |>

    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0

if (exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0) {
  cat("Missing yellowfin catch with effort are reported below.\n\n")
}
```

```{r missing_yft, include=FALSE}

# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0){
  flextable(missing_unraised_yft |> select(-no_catch)) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

  colformat_num(j = 2:13, digits = 0) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: raise-bet
#| tbl-cap: !expr if ("bet" %in% names(raised_list)) glue::glue("Raised bigeye catch (kgs) reported in ereporting tools.") else ""

# Remove rows where no catch was reported - as no average exists...

if ("bet" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$bet 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_bet = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0

if (exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0) {
  cat("Missing bigeye catch with effort are reported below.\n\n")
}
```

```{r missing_bet, include=FALSE}
# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0){
  flextable(missing_unraised_bet|> select(-no_catch)) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

  colformat_num(j = 2:13, digits = 0) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

<!-- ## Final Estimates -->

<!-- Here we take the raised catch estimates for the three key tuna species and sum them across all landing sites and months to produce an annual estimate in metric tonnes (mt). -->

```{r}
#| label: tbl-raised-ace
#| tbl-cap: Raised and unraised annual artisanal catch estimates in mt

ace <- data.frame()

for (sp in species) {
  
  if (sp %in% names(unr_list)) {
    
    unraised <- unr_list[[sp]] |>
      data.frame() |>
      ungroup() |>
      select(-c(1, 14)) |>
      summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) |>
      rowSums()
    
    raised <- raised_list[[sp]] |>
      data.frame() |>
      ungroup() |>
      select(-1) |>
      summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) |>
      rowSums()
    
  } else {
    unraised <- 0
    raised   <- 0
  }
  
  ace <- bind_rows(ace, data.frame(
    sp       = toupper(sp),
    unraised = round(unraised / 1000, 2),
    raised   = round(raised / 1000, 2)
  ))
}

# Produce final output
ace_final = ace |>
  gather(key='Method', value='ACE', -1) |>
  spread(key = sp, value=ACE)

flextable(ace_final) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)

```

```{r}
#| label: tbl-raised-ace-perc
#| tbl-cap: Percentage of the artisanal tuna ACE associated with each of the key species

ace_final$tot = rowSums(ace_final[,-1])
ace_final <- ace_final |>  mutate(SKJ = round(SKJ/tot*100,2),
                      YFT = round(YFT/tot*100,2),
                      BET = round(BET/tot*100,2)) |>
  mutate(across(BET:YFT, ~replace_na(.,0))) |>
  select(-tot)

flextable(ace_final) |>
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)

```

# Socio-economic factors

Whilst revenue from fishing license contributes largely to Tokelau’s annual budget, Artisanal Fisheries is the main source of protein for the local communities.  In many occasions, especially if the catch is in abundance, the “inati system” will take effect.  This is a traditional system where resources are being distributed to ensure that every household on island receives a share of the resources.  A tradition that is rooted in the cultural value of faka-Tokelau (Tokelauan way of life) and it serves as a powerful communal safety net that guarantees that nobody is being left behind.

# Disposal of catch

Licensed vessels disposes of catches at their designated ports.  For inshore fisheries, catches are disposed at the entry point (channels) and where fishermen will dock.

# Onshore developments

As part of our reporting obligations, for inshore fisheries, the Forum Fisheries Agency provided tablets to assist in reporting data using ER and sending off data using TAILS.

# Status of tuna fishery data collection systems

Artisanal data is collected on a daily basis on all three atolls. Data is then being sent using TAILS to SPC where they will be entered into the TUFMAN 2 system and database.

## TUFMAN 2 -- Artisanal Fisheries Component

The tuna research and statistical needs of Tokelau are largely met by the assistance and support of SPCs’ Oceanic Fisheries Programme. Part of this assistance is the implementation of the Tuna Management System TUFMAN, now upgraded to TUFMAN 2, in Tokelau. It is now fully functional with ongoing assistance with training as new systems are developed. This initiative will empower the Fisheries Management Agency with a tool to help provide advice to local government and to help fulfil obligations and commitments to the Commission and other regional organizations. Whilst TUFMAN 2 manages the commercial data (offshore fisheries) TUF-ART has been implemented for managing the artisanal fishery (inshore fisheries). The new artisanal fisheries component of TUFMAN 2 was introduced to Tokelau fisheries oﬀicers in late 2016. One of the major improvements to the data system was the update to TAILS software in 2018.

## Monitoring

All foreign fishing vessels that are licensed to fish in Tokelau waters are required to carry an FFA approved Automatic Location Communicator (ALC) or Mobile Transceiver (MTU) and whilst within Tokelau waters, they are monitored by the FFA installed Vessel Monitoring Systems (VMS). This is a critical part of MCS and a workstation was installed in Tokelau in 2005. Tokelau aims to strengthen MCS capabilities through participation in regional and sub-regional strategies and through regional infrastructures. Due to limited resources available, Tokelau must participate in regional and sub-regional collaborations to combat IUU in the Tokelau EEZ.

As a participant in the PNA VDS established under the Palau Arrangement, Tokelau is now implementing the VDS for all their purse seine fisheries and the longline fisheries.  All vessels licensed to fish against Tokelau’s vessel days within its EEZ are also monitored through the PNA VDS systems.

Tokelau appreciates the value of an Observer Programme and will consider appropriate collaborative effort in the region. Required observer services are provided through other FFA members’ country observer programmes and the PNA Observer Programme. Tokelau has no port and wharfage infrastructure so there is no Port Sampling Programme and no port sampling is carried out in Tokelau.

Tokelau is committed to working cooperatively with respect to monitoring and enforcement, and the sharing of MCS assets and information with her counterparts to strengthen the MCS arrangements for Tokelau’s EEZ. Surface and air patrols/surveillance of Tokelau’s EEZ are undertaken by New Zealand and FFA coordinated patrols, with input and assistance from the Tokelau Fisheries Management Agency.


\newpage

# Addendum

```{r loadaddendum}
#| output: asis
#| echo: false

# Set values the child can read
applicable_measures <- "none"

cat(knitr::knit_child("template_addendum.qmd", quiet = TRUE))
```

\newpage

# Tokelau Contextual Statement

The following points should be taken into account when considering Tokelau’s completion of her AR Prt1 & AR Prt2 Reporting Template. Tokelau will refer to these points by their number when responding to fields in the template. 

**1.** Tokelau is a small island developing territory (administered by New Zealand).  

**2.** Tokelau consists of three atolls with a population of 1400 people. The Tokelau EEZ is geographically situated between approximately 6 degrees and 11degrees South of latitude and between 167 and 176 degrees West of longitude 

**3.** New Zealand signed UNCLOS on behalf of Tokelau and therefore New Zealand is responsible at international law for the Tokelau EEZ. Some of the administrative functions relating to the management of the Tokelau EEZ have been delegated by New Zealand to the government of Tokelau.  

**4.** Tokelau is a participating territory in the Western and Central Pacific Fisheries Commission. 

**5.** Tokelau is not a flag state. Tokelau does not authorize vessel to fish on the High Seas. Tokelau does not charter Foreign Fishing Vessels. 

**6.** Tokelau has no ports and is therefore not a Port State. 

**7.** While transshipments in the EEZ is permitted by law in certain circumstances, as a matter of policy Tokelau does not currently allow transshipments to occur in its EEZ, Territorial Sea or Internal Waters. 

**8.** Tokelau has very limited fisheries scientific, management or MCS capability. However New Zealand, FFA, SPC, TVM and PNAO provide fisheries services to support Tokelau meet its fisheries obligations. 

**9.** Tokelau does not have a domestic commercial fishing industry. 

**10.** The fisheries of the Tokelau EEZ are managed by way of an Offshore Fisheries Management Plan that sets tonnage zone limits for each of thetuna species supported by numerical limits on purse seine and long line licenses. On the 1stJanuary 2013 the management of the Tokelau purse seine fishery has been managed by way of the Vessel Day Scheme established under the Palau Arrangement. 

Tokelau has a very small artisanal fishery that is carried out within its 12 nautical mile Territorial Sea and Internal Waters. This artisanal fishery is managed by the elders of each atoll with the primary objective of maintaining the food security and cultural wellbeing of each of the 3 island communities.
