---
format: 
  pdf:
    geometry: "margin=2cm"
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"
    include-in-header:
      text: |
        \usepackage{float}
        \floatplacement{figure}{H}
        \floatplacement{table}{H}
editor: source
toc: false
embed-resources: true 
execute: 
  echo: false
  message: false
  warning: false
  fig-width: 12
  fig-height: 8
params:
  flag: 'KI'
  year: 2025
  author: "AUTHOR"
---

This section reports on fishing activity in your EEZ conducted by foreign countries (Distant Water Fishing Nations).

```{r}
# Load necessary libraries
library(DBI)
library(odbc)
library(tidyverse)
library(flextable)
library(scatterpie)
library(mregions2)
library(countrycode)
source("utils.R")

sp_colours <- setNames(object = c("seagreen3","red3","black","cyan1","steelblue","orange",
                                  "pink","wheat","purple3","grey40","dodgerblue4","purple4",
                                  "#66CCFF","salmon4","magenta3","gold", "gray"),
                         nm = c("Albacore","Bigeye tuna","Black marlin","Blue marlin","Blue shark",
                                "Hammerhead sharks nei","Mako sharks","Oceanic whitetip shark",
                                "Pacific bluefin tuna","Silky shark","Skipjack tuna",
                                "Striped marlin","Swordfish","Thresher sharks nei",
                                "Whale shark","Yellowfin tuna", "Other"))
# set theme
theme_set(
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.text     = element_text(size = 8),
    legend.title    = element_text(size = 9),
    legend.key.size = unit(0.4, "cm")
  )
)

# Create a vector of the key species to plot for each gear
ps_species <- c("skj", "yft", "bet", "alb", "pbf")
ll_species <- c("skj", "yft", "bet", "alb", "pbf", "blm", "bum", "mls", "swo")
pl_species <- c("skj", "yft", "bet", "alb", "pbf")

# adjust path as needed
flag_lookup <- read_csv("config/flag_names.csv") |>
  select(country_code, country_name, country_short) |>
  mutate(country_name = na_if(country_name, "NULL"),
         country_short = na_if(country_short, "NULL")) |>
  # use short name when available, fall back to long
  mutate(name_pretty = coalesce(str_to_title(tolower(str_trim(country_short))), str_to_title(country_name)))

flag_labels <- flag_lookup |>
  mutate(flag_code_lower = tolower(country_code)) |>
  select(flag_code_lower, name_pretty) |>
  deframe()  # from tibble: turns 2-col df into a named vector

# Helper: get a country's 200NM EEZ as an sf object, keyed by ISO-2 code
get_country_eez <- function(iso2_code) {
  iso3 <- countrycode(iso2_code, "iso2c", "iso3c", warn = FALSE)
  
  if (is.na(iso3)) {
    stop("Could not resolve ISO-2 code '", iso2_code, "' to ISO-3.")
  }
  
  eez <- mrp_get(
    "eez",
    cql_filter = sprintf("iso_ter1 = '%s' AND pol_type = '200NM'", iso3)
  )
  
  if (nrow(eez) == 0) {
    warning("No EEZ found for '", iso2_code, "' (ISO-3: ", iso3, ").")
  }
  
  eez
}

# Helper: build a gridded species-composition catch summary for one EEZ
grid_catch <- function(data, eez_code, cell_size = 1,
                       species = c("skj", "yft", "bet")) {
  
  species_cols <- paste0(species, "_mt")
  half <- cell_size / 2
  
  data |>
    # filter(eez == eez_code) |>
    mutate(
      lat_bin = floor(latd / cell_size) * cell_size + half,
      lon_bin = floor(lond / cell_size) * cell_size + half
    ) |>
    group_by(lon_bin, lat_bin) |>
    summarise(across(all_of(species_cols), \(x) sum(x, na.rm = TRUE)),
              .groups = "drop") |>
    rename_with(\(x) str_remove(x, "_mt$"), all_of(species_cols)) |>
    mutate(total = rowSums(across(all_of(species)))) |>
    filter(total > 0)
}

# Detect context: standalone render vs. knit_child from parent
# - Standalone: params$flag exists (set via execute_params)
# - Child:      country_cd is already defined by the parent
if (!exists("country_cd") || length(country_cd) == 0 || !nzchar(country_cd)) {
  country_cd <- tolower(params$flag)
}

if (!exists("report_year") || length(report_year) == 0) {
  report_year <- params$year
}

if (!exists("this_yr_folder") || length(this_yr_folder) == 0 || !nzchar(this_yr_folder)) {
  this_yr_folder <- paste0("./data/report_", as.character(report_year), "_", country_cd, "/") |>
    tolower()
}

# Sanity check — fail loudly if anything is still missing
stopifnot(
  length(country_cd)     == 1, nzchar(country_cd),
  length(report_year)    == 1,
  length(this_yr_folder) == 1, nzchar(this_yr_folder)
)

# Define years to use for tables and figures
yrs_long <- (report_year - 4):report_year
yr_range <- paste(min(yrs_long), max(yrs_long), sep = "-")
'%nin%' <- function(x, y) !x %in% y

country_name <- flag_labels[[country_cd]]


parse_coord <- function(x, type = c("lat", "lon")) {
  type <- match.arg(type)
  
  # split into the numeric body and the hemisphere letter
  hem <- str_sub(x, -1)
  num <- str_sub(x, 1, -2)
  
  # last two chars are minutes, the rest is degrees
  mins <- as.numeric(str_sub(num, -2))
  degs <- as.numeric(str_sub(num, 1, -3))
  
  decimal <- degs + mins / 60
  
  # apply sign based on hemisphere
  sign <- dplyr::case_when(
    hem %in% c("n", "e") ~  1,
    hem %in% c("s", "w") ~ -1,
    TRUE                 ~ NA_real_
  )
  
  decimal * sign
}

```

```{r data-prep}
# Load all raw data - from t2
data_3615 <- read_and_clean(this_yr_folder, country_cd, r_code = 3615) |>
      mutate(across(where(is.character), tolower))

data_3527 <- read_and_clean(this_yr_folder, country_cd, r_code = 3527)|>
      mutate(across(where(is.character), tolower))

data_3614 <- read_and_clean(this_yr_folder, country_cd, r_code = 3614)|>
      mutate(across(where(is.character), tolower))

data_2900 <- read_and_clean(this_yr_folder, country_cd, r_code = 2900)|>
      mutate(across(where(is.character), tolower)) |>
  filter(!(flag == country_cd))  # PS catch EEZ per country

data_2894<- read_and_clean(this_yr_folder, country_cd, r_code = 2894)|>
      mutate(across(where(is.character), tolower))|>
      filter(!(flag == country_cd))# LL catch EEZ (I think…)

data_2904<- read_and_clean(this_yr_folder, country_cd, r_code = 2904)|>
      mutate(across(where(is.character), tolower)) |>
      filter(!(flag == country_cd))  # PS catch spatial EEZ

data_3375<- read_and_clean(this_yr_folder, country_cd, r_code = 3375)|>
      mutate(across(where(is.character), tolower))|>
      filter(!(flag == country_cd)) # LL catch spatial EEZ

# FIX FOR TO
if(nrow(data_3375) > 0 && country_cd == "to"){
  # there are some issues with WS vessels positions - they are wrongly assigning catches in TO EEZ
  data_3375 <- data_3375 |> 
    filter(!(flag == "ws"))
}


if(nrow(data_2904) > 0){
  data_2904 <- data_2904 |>
    rename(year = t2col0) |>
    filter(year == report_year) 
}

if (nrow(data_2900) > 0){
  data_2900 <- data_2900 |>
    mutate(year = as.numeric(substr(t2col1, 1, 4))) |>
    filter(year >= report_year - 4) |>
    filter(year <= report_year) |>
    select(-fish_days_aws_only, - t2col1)
}

if (nrow(data_2894) > 0){
  data_2894 <- data_2894 |>
    mutate(year = as.numeric(substr(t2col4, 1, 4))) |>
    filter(year >= report_year - 4) |>
    filter(year <= report_year) |>
    select(flag, year, vessels = vessel_name, trips, days, days, alb_mt, bet_mt, yft_mt, oth_mt, tot_mt)
}

# FIX FOR NR
if(nrow(data_2894) > 0 && country_cd == "nr"){
  # some LL catch showing for NR but these must be errors in reported positions
  if (sum(data_2894$tot_mt) < 4) {
    data_2894 <- data.frame()
  }
  
}

ika_trips_file = paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")

if (file.exists(ika_trips_file)){
  data_ika_raw <- tryCatch(
    read.csv(paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")),
    error = function(e) NULL
  )
  
  if (exists("data_ika_raw")){
    if ("LandingSite" %in% colnames(data_ika_raw) && nrow(data_ika_raw) > 0){
        data_ika_raw <- data_ika_raw|>
          rename(landing_site_name = LandingSite) |>
          group_by(Country, landing_site_name, Year, Month) |>  
          summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop") |>
          filter(Year == report_year)
    }
    
  }
  
  if (is.null(data_ika_raw) || nrow(data_ika_raw) == 0 || is.null(data_ika_raw)) rm(data_ika_raw)
  }

if (!exists("data_ika_raw") && nrow(data_3615) == 0) {
  cat("No Ikasavea or Tails trip data found for this country/year. Document generation stopped.\n")
  knitr::knit_exit()
}

if (exists("data_ika_raw")){
  
   data_ika_trips <- data_ika_raw |>
    mutate(across(where(is.character), tolower)) |>
    janitor::clean_names() |>
    select(landing_site_name, month,
           number_trips)
  
  df_ika_wide <- data_ika_trips %>%
    complete(landing_site_name, month = 1:12, fill = list(number_trips = 0)) %>%
    mutate(month = month.abb[month]) %>%
    pivot_wider(names_from = month, values_from = number_trips) %>%
    select(landing_site_name, all_of(month.abb))
  
  data_ika_catch_kg <- data_ika_raw |>
    mutate(across(where(is.character), tolower)) |>
    janitor::clean_names() |>
    select(landing_site_name, month, skj = skj_kg, yft = yft_kg, bet = bet_kg)#, alb = alb_kg)
  
  data_ika_catch_kg <- data_ika_catch_kg |>
    pivot_longer(cols = c("skj", "bet", "yft"), names_to = "sp_code", values_to = "sp_kg")
  
  }

both_exist <- exists('data_ika_trips') && nrow(data_ika_trips) > 0 && exists('data_3615') && 
              is.data.frame(get('data_3615')) && nrow(data_3615) > 0

local_knowl_trips_file <-  str_c(this_yr_folder,  'additional_files/est_trips.csv')

```

## Fishing Licenses

**NOTE:** Here, members should report the Number of Licenses issued for fishing in their EEZ by vessel gear type in the last 5 years. *This information should be sourced from IFIMS.*


## Catch and Effort Trends

```{r data-eez-all-flags}

if (nrow(data_2900) > 0){
  
  # Table - overall per year
  data_2900_summary <- data_2900 |>
    group_by(year) |>
    summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) |>
    ungroup() |>
    mutate(across(year:fish_days, ~ round(.x, 0)),
           gear = "Purse seine") |>
    select(-sea_days)
  
  plot_data1ps <- data_2900 |>
    mutate(tot_mt = replace_na(tot_mt, 0)) |>          
    mutate(flag = fct_lump_n(flag, n = 5, w = tot_mt, other_level = "Other")) |>
    group_by(year, flag) |>
    summarise(tot_mt = sum(tot_mt, na.rm = TRUE), .groups = "drop") |>
    mutate(flag = fct_reorder(flag, tot_mt, .fun = sum, .desc = TRUE),
           gear = "Purse seine") |> 
    select(year, flag, tot_mt, gear)
  
  # Country and sp catch 
  plot_data2ps <- data_2900 |>
    filter(year == report_year) |>
    mutate(across(c(skj_mt, bet_mt, yft_mt, oth_mt), \(x) replace_na(x, 0))) |>
    pivot_longer(c(skj_mt, bet_mt, yft_mt, oth_mt),
                 names_to = "species", values_to = "mt") |>
    mutate(species = recode(species,
                            skj_mt = "Skipjack tuna",
                            bet_mt = "Bigeye tuna",
                            yft_mt = "Yellowfin tuna",
                            oth_mt = "Other")) |>
    mutate(flag    = fct_reorder(flag, mt, .fun = sum, .desc = TRUE),
           species = fct_reorder(species, mt, .fun = sum, .desc = TRUE),
           gear = "Purse seine") |> 
    select(year, flag, species, mt, gear)
}

if (nrow(data_2894) > 0){
  
  # Table - overall per year
  data_2894_summary <- data_2894 |>
    group_by(year) |>
    summarise(vessels = n_distinct(vessels), 
              across(where(is.numeric), ~ sum(.x, na.rm = TRUE))
              ) |>
    mutate(across(year:days, ~ round(.x, 0)),
           gear = "Longline") |>
    rename(fish_days = days)
  
  plot_data1ll <- data_2894 |>
    mutate(tot_mt = replace_na(tot_mt, 0)) |>          
    mutate(flag = fct_lump_n(flag, n = 5, w = tot_mt, other_level = "Other")) |>
    group_by(year, flag) |>
    summarise(tot_mt = sum(tot_mt, na.rm = TRUE), .groups = "drop") |>
    mutate(flag = fct_reorder(flag, tot_mt, .fun = sum, .desc = TRUE),
           gear = "Longline") |> 
    select(year, flag, tot_mt, gear)
  
  # Country and sp catch 
  plot_data2ll <- data_2894 |>
    filter(year == report_year) |>
    mutate(across(c(alb_mt, bet_mt, yft_mt, oth_mt), \(x) replace_na(x, 0))) |>
    pivot_longer(c(alb_mt, bet_mt, yft_mt, oth_mt),
                 names_to = "species", values_to = "mt") |>
    mutate(species = recode(species,
                            yft_mt = "Yellowfin tuna",
                            alb_mt = "Albacore",
                            bet_mt = "Bigeye tuna",
                            oth_mt = "Other")) |>
    mutate(flag    = fct_reorder(flag, mt, .fun = sum, .desc = TRUE),
           species = fct_reorder(species, mt, .fun = sum, .desc = TRUE),
           gear = "Longline") |>
   select(year, flag, species, mt, gear)
  }

has_2894 <- exists("data_2894_summary")
has_2900 <- exists("data_2900_summary")

if (has_2894 && has_2900) {
  flag_summary <- bind_rows(data_2894_summary, data_2900_summary)
  plot_data1   <- bind_rows(plot_data1ll, plot_data1ps)
  plot_data2   <- bind_rows(plot_data2ll, plot_data2ps)
} else if (has_2894) {
  flag_summary <- data_2894_summary   # longline only
  plot_data1   <- plot_data1ll
  plot_data2   <- plot_data2ll
} else if (has_2900) {
  flag_summary <- data_2900_summary   # purse seine only
  plot_data1   <- plot_data1ps
  plot_data2   <- plot_data2ps
}else{
  flag_summary <- data.frame()
  plot_data1   <- data.frame()
  plot_data2   <- data.frame()
}

# remove any rows where fishing did not happen
if(nrow(flag_summary) > 0){
  flag_summary <- flag_summary |>
    filter(fish_days > 0)
}
if(nrow(plot_data1) > 0){
  plot_data1 <- plot_data1 |>
    filter(tot_mt > 0)
  
  if(country_cd == 'to'){
    plot_data1 <- plot_data1 |>
      filter(!(flag == "ws"))
    }
  }
if(nrow(plot_data2) > 0){
  plot_data2 <- plot_data2 |>
    filter(mt > 0) 
  
  if(country_cd == 'to'){
    plot_data2 <- plot_data2 |>
      filter(!(flag == "ws"))
    }
  }

```

```{r}
#| echo: false
#| include: true
#| results: 'asis'
#| label: tbl-data-eez-all-flags
#| tbl-cap: !expr if(nrow(flag_summary) > 0) glue::glue("Annual catches in {country_name} EEZ per gear (LL = Longline; PS = Purse seine) for the last 5 years. Source - Tufman 2, reports 2900 and 2894.") else ""

if (nrow(flag_summary) > 0) {
  
  # Define the full set of possible columns and their labels
  all_mt_cols  <- c("alb_mt", "bet_mt", "yft_mt", "skj_mt", "oth_mt", "tot_mt")
  all_int_cols <- c("year", "vessels", "trips", "fish_days")
  
  all_labels <- list(
    gear      = "Gear",
    year      = "Year",
    vessels   = "Vessels",
    trips     = "Trips",
    fish_days = "Days",
    alb_mt    = "Alb (mt)",
    bet_mt    = "Bet (mt)",
    yft_mt    = "Yft (mt)",
    skj_mt    = "Skj (mt)",
    oth_mt    = "Others (mt)",
    tot_mt    = "Total (mt)"
  )
  
  # Filter to only columns that actually exist in this run's data
  mt_cols     <- intersect(all_mt_cols,  names(flag_summary))
  int_cols    <- intersect(all_int_cols, names(flag_summary))
  labels_use  <- all_labels[intersect(names(all_labels), names(flag_summary))]
  
  # Build relocate list dynamically — only push columns to the end if they exist
  relocate_cols <- intersect(c("oth_mt", "tot_mt"), names(flag_summary))
  
  ft <- flag_summary |>
    relocate(all_of(relocate_cols), .after = last_col()) |>
    mutate(year = as.integer(year)) |>
      # apply sign based on hemisphere
    mutate(gear = dplyr::case_when(
      gear %in% c("Longline") ~  "LL",
      gear %in% c("Purse seine") ~ "PS",
      TRUE ~ gear
    )) |>
    as_grouped_data(groups = "gear") |>
    flextable() |>
    set_header_labels(values = labels_use) |>
    bold(i = ~ !is.na(gear)) |>
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5) |>
    theme_booktabs() |>
    align(align = "right", part = "all") |>
    align(j = "gear", align = "left", part = "all")
  
  if (length(mt_cols) > 0) {
    ft <- ft |> colformat_double(j = mt_cols, digits = 0, na_str = "–")
  }
  if (length(int_cols) > 0) {
    ft <- ft |> colformat_int(j = int_cols)
  }
  
  ft
}
```

```{r}
#| echo: false
#| include: true
#| results: 'asis'
#| label: fig-data-eez-all-flags-perc
#| fig-cap: !expr if(nrow(plot_data1) > 0) glue::glue("Percentage of total catch (all species) in the {country_name} EEZ, by flag during the last 5 years. Source - Tufman 2, reports 2900 and 2894.") else ""

if (nrow(plot_data1) > 0) {

  plot_data1 <- plot_data1 |> 
    mutate(flag = toupper(flag),
           flag = forcats::fct_reorder(flag, tot_mt, .fun = sum, .desc = TRUE)
    )


  flags <- sort(unique(plot_data1$flag))

  flag_labels <- setNames(stringr::str_to_title(flag_lookup$name_pretty),
                        flag_lookup$country_code)
  flag_labels <- c(flag_labels, OTHER = "Other")
  
  # fall back to the code itself for anything not in the lookup (e.g. OTHER)
  flag_labels[is.na(flag_labels)] <- names(flag_labels)[is.na(flag_labels)]

  ggplot(plot_data1, aes(x = factor(year), y = tot_mt, fill = flag)) +
    geom_col(position = "fill", colour = "white", linewidth = 0.3) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_hue(labels = flag_labels, c = 50) +
    facet_wrap(~ gear) +
    labs(x = NULL, y = "Share of catch", fill = "Country")
}
```

```{r}
#| echo: false
#| include: true
#| results: 'asis'
#| label: fig-data-eez-all-flags-hist
#| fig-cap: !expr if(nrow(plot_data2) > 0) glue::glue("Annual catch of tuna in the {country_name} EEZ by flag during {report_year}. Source - Tufman 2, reports 2900 and 2894.") else ""

if (nrow(plot_data2) > 0) {

  # labels built from the full lookup, so every flag in any plot is covered
  flag_labels <- setNames(stringr::str_to_title(flag_lookup$name_pretty),
                        flag_lookup$country_code)
  # quick guard: any species missing from the colour palette?
  missing_sp <- setdiff(unique(plot_data2$species), names(sp_colours))
  if (length(missing_sp) > 0) {
    warning("Species missing from sp_colours: ", paste(missing_sp, collapse = ", "))
  }

  plot_data2 |>
    mutate(
      flag    = toupper(flag),
      # order flags by total catch (ascending = biggest at top after flip)
      flag    = forcats::fct_reorder(flag, mt, .fun = sum),
      # push "Other" to the end of the legend
      species = forcats::fct_relevel(factor(species), "Other", after = Inf)
    ) |>
    ggplot(aes(x = mt, y = flag, fill = species)) +
    geom_col(position = position_stack(reverse = TRUE)) +
    scale_y_discrete(labels = flag_labels) +
    scale_x_continuous(labels = scales::comma) +
    scale_fill_manual(values = sp_colours, na.value = "grey80") +
    facet_wrap(~ gear) +
    labs(x = "Catch (mt)", y = NULL, fill = "Species")
}
```

\newpage

```{r}
#| echo: false
#| include: true
#| results: 'asis'
#| fig-width: 10
#| fig-height: 8
#| label: fig-data-eez-ps-spatial
#| fig-pos: 'H'
#| fig-cap: !expr if(nrow(data_2904) > 0) glue::glue("Catch of the three main tuna species by purse  seine  vessels  fishing  in  the  {country_name}  EEZ,  in  each  1x1  degree  cell  in  the EEZ and  the immediate surrounding area during {report_year}. Source - Tufman 2, report 2904.") else ""
if (nrow(data_2904) > 0) {

  # ── EEZ: shift to 0-360 longitude space so EEZs crossing the dateline
  #    (FJ, TV, KI, ...) get a sensible bbox, then dissolve the two
  #    antimeridian halves so geom_sf doesn't draw a seam at 180.
  #    IMPORTANT: drop the CRS *before* st_union(). With a CRS attached,
  #    st_union() under s2 (the default in a fresh render) is spherical
  #    and normalises longitudes back to [-180, 180], silently undoing
  #    the shift — it only works interactively if s2 happens to be
  #    disabled in the session. With no CRS the union is planar and the
  #    0-360 coordinates survive. st_shift_longitude() is the only step
  #    that needs the CRS, so it runs first. All no-ops for EEZs that
  #    don't cross the dateline.
  eez <- get_country_eez(country_cd) |>
    st_make_valid() |>
    st_shift_longitude() |>
    st_set_crs(NA) |>
    st_union() |>
    st_make_valid()

  bbox <- st_bbox(eez)

  # ── Land: shift, crop to the (padded) map extent, then drop the CRS
  #    to match the EEZ layer. The CRS-less layers also make coord_sf
  #    honour scale_x_continuous breaks past 180 — with a geographic CRS
  #    it routes breaks through st_graticule(), which normalises to
  #    [-180, 180] and silently drops anything >= 180 (the missing
  #    x-label problem).
  land <- suppressWarnings(
    ne_countries(scale = "medium", returnclass = "sf") |>
      st_make_valid() |>
      st_shift_longitude() |>
      st_crop(bbox + c(-2, -2, 2, 2)) |>
      st_set_crs(NA)
  )

  # ── Species lookup: short code → full name (must match sp_colours keys)
  species_map <- c(
    # alb = "Albacore tuna",
    skj = "Skipjack tuna",
    yft = "Yellowfin tuna",
    bet = "Bigeye tuna"
  )

  # warn loudly (build log) if the palette is missing a species
  missing_sp <- setdiff(species_map, names(sp_colours))
  if (length(missing_sp) > 0) {
    warning("Species missing from sp_colours: ",
            paste(missing_sp, collapse = ", "))
  }

  catch_grid <- grid_catch(data_2904,
                           eez_code  = tolower(country_cd),
                           cell_size = 1,
                           species   = names(species_map))

  # Align pie longitudes with the 0-360 convention (no-op if already 0-360)
  catch_grid <- catch_grid |>
    mutate(lon_bin = ifelse(lon_bin < 0, lon_bin + 360, lon_bin)) |>
    filter(total > 0)

  # only draw if there is anything to draw (grid_catch can return zero
  # usable cells even when data_2904 has rows)
  if (nrow(catch_grid) > 0) {

    max_r <- 0.45
    catch_grid <- catch_grid |>
      mutate(radius = max_r * sqrt(total / max(total)))

    # sanity check: pies should fall within the map extent; if not, the
    # longitude conventions of grid_catch and the shifted EEZ disagree
    out_of_extent <- catch_grid |>
      filter(lon_bin < bbox["xmin"] - 2 | lon_bin > bbox["xmax"] + 2 |
             lat_bin < bbox["ymin"] - 2 | lat_bin > bbox["ymax"] + 2)
    if (nrow(out_of_extent) > 0) {
      warning(country_cd, ": ", nrow(out_of_extent),
              " grid cell(s) fall outside the EEZ map extent and will not be visible.")
    }

    # ── Fill values and labels from the lookup
    fill_values <- setNames(sp_colours[species_map], names(species_map))
    fill_labels <- setNames(species_map, names(species_map))

    # ── Axis breaks: ~8 longitude labels whatever the EEZ width, labelled
    #    conventionally either side of the dateline; latitude labels are
    #    needed explicitly now that the CRS (and its graticule) is gone.
    x_span <- as.numeric(bbox["xmax"] - bbox["xmin"])
    step   <- max(1, round(x_span / 8))
    xbr    <- seq(floor(bbox["xmin"]), ceiling(bbox["xmax"]), by = step)
    lon_lab <- function(x) {
      x <- x %% 360
      ifelse(x == 180, "180\u00B0",
             ifelse(x > 180, paste0(360 - x, "\u00B0W"),
                    paste0(x, "\u00B0E")))
    }
    lat_lab <- function(y) {
      ifelse(y < 0, paste0(abs(y), "\u00B0S"),
             ifelse(y > 0, paste0(y, "\u00B0N"), "0\u00B0"))
    }

    # ── Aspect ratio: with no CRS, coord_sf assumes projected coordinates
    #    (1 deg lon = 1 deg lat on screen). Correct for longitude
    #    convergence at the EEZ's mid-latitude so shapes aren't stretched
    #    E-W — matters for the higher-latitude PICTs (CK, PF, NU, ...).
    mid_lat <- mean(c(bbox["ymin"], bbox["ymax"]))
    y_span  <- as.numeric(bbox["ymax"] - bbox["ymin"])
    asp     <- (y_span / x_span) / cos(mid_lat * pi / 180)

    p <- ggplot() +
      geom_sf(data = land, fill = "grey90", colour = "grey60", linewidth = 0.2) +
      geom_sf(data = eez,  fill = "grey50", colour = "black", alpha = 0.4) +
      geom_scatterpie(
        data   = catch_grid,
        aes(x = lon_bin, y = lat_bin, r = radius),
        cols   = names(species_map),
        colour = NA,
        alpha  = 0.9
      ) +
      scale_fill_manual(
        values = fill_values,
        labels = fill_labels,
        name   = "Species"
      ) +
      scale_x_continuous(breaks = xbr, labels = lon_lab) +
      scale_y_continuous(labels = lat_lab) +
      coord_sf(
        xlim   = c(bbox["xmin"], bbox["xmax"]),
        ylim   = c(bbox["ymin"], bbox["ymax"]),
        expand = FALSE
      ) +
      labs(x = "Longitude", y = "Latitude") +
      theme_bw() +
      theme(legend.position = "bottom",
            aspect.ratio = asp,
            axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 10),
            axis.text.y = element_text(size = 10),
            panel.background = element_rect(fill = "white"))

    print(p)

  } else {
    message(country_cd, ": no positive-catch grid cells — spatial figure skipped.")
  }
}
```

```{r}
#| echo: false
#| include: true
#| results: 'asis'
#| fig-width: 8
#| fig-height: 9
#| out-width: "90%"
#| fig-pos: 'H'
#| label: fig-data-eez-ll-spatial
#| fig-cap: !expr if(nrow(data_3375) > 0) glue::glue("Catch of the three main tuna species by longline  vessels  fishing  in  the  {country_name}  EEZ,  in  each  1x1  degree  cell  in  the EEZ and  the immediate surrounding area during {report_year}. Source - Tufman 2, report 3375.") else ""
if (nrow(data_3375) > 0 && nrow(data_2894) > 0) {

  # ── Species lookup: short code → full name (must match sp_colours keys)
  species_map <- c(
    alb = "Albacore",
    skj = "Skipjack tuna",
    yft = "Yellowfin tuna",
    bet = "Bigeye tuna"
  )

  # warn loudly (build log) if the palette is missing a species
  missing_sp <- setdiff(species_map, names(sp_colours))
  if (length(missing_sp) > 0) {
    warning("Species missing from sp_colours: ",
            paste(missing_sp, collapse = ", "))
  }

  data_3375 <- data_3375 |>
    mutate(
      latd = parse_coord(latitude,  "lat"),
      lond = parse_coord(longitude, "lon")
    ) |>
    select(flag, vessel_name, log_date, latd, lond, species, mt) |>
    filter(species %in% names(species_map)) |>
    pivot_wider(names_from = species, values_from = mt) |>
    # current year
    mutate(year = substr(log_date, 7, 10)) |>
    filter(year %in% report_year) |>
    # name-based rename: pivot_wider orders columns by first appearance in
    # the data, NOT in a fixed order, so a positional colnames()[6:9] <- ...
    # can silently mislabel species. any_of() also tolerates a species
    # having no rows at all for this country/year.
    rename_with(\(x) paste0(x, "_mt"), any_of(names(species_map)))

  # ── EEZ: shift to 0-360 longitude space so EEZs crossing the dateline
  #    (FJ, TV, KI, ...) get a sensible bbox, then dissolve the two
  #    antimeridian halves so geom_sf doesn't draw a seam at 180.
  #    IMPORTANT: drop the CRS *before* st_union(). With a CRS attached,
  #    st_union() under s2 (the default in a fresh render) is spherical
  #    and normalises longitudes back to [-180, 180], silently undoing
  #    the shift — it only works interactively if s2 happens to be
  #    disabled in the session. With no CRS the union is planar and the
  #    0-360 coordinates survive. st_shift_longitude() is the only step
  #    that needs the CRS, so it runs first. All no-ops for EEZs that
  #    don't cross the dateline.
  eez <- get_country_eez(country_cd) |>
    st_make_valid() |>
    st_shift_longitude() |>
    st_set_crs(NA) |>
    st_union() |>
    st_make_valid()

  bbox <- st_bbox(eez)

  # ── Land: shift, crop to the (padded) map extent, then drop the CRS
  #    to match the EEZ layer. The CRS-less layers also make coord_sf
  #    honour scale_x_continuous breaks past 180 — with a geographic CRS
  #    it routes breaks through st_graticule(), which normalises to
  #    [-180, 180] and silently drops anything >= 180 (the missing
  #    x-label problem).
  land <- suppressWarnings(
    ne_countries(scale = "medium", returnclass = "sf") |>
      st_make_valid() |>
      st_shift_longitude() |>
      st_crop(bbox + c(-2, -2, 2, 2)) |>
      st_set_crs(NA)
  )

  catch_grid <- grid_catch(data = data_3375,
                           eez_code  = tolower(country_cd),
                           cell_size = 1,
                           species   = names(species_map))

  # Align pie longitudes with the 0-360 convention (no-op if already 0-360)
  catch_grid <- catch_grid |>
    mutate(lon_bin = ifelse(lon_bin < 0, lon_bin + 360, lon_bin)) |>
    filter(total > 0)

  # only draw if there is anything to draw (grid_catch can return zero
  # usable cells even when data_3375 has rows)
  if (nrow(catch_grid) > 0) {

    max_r <- 0.45
    catch_grid <- catch_grid |>
      mutate(radius = max_r * sqrt(total / max(total)))

    # sanity check: pies should fall within the map extent; if not, the
    # longitude conventions of grid_catch and the shifted EEZ disagree
    out_of_extent <- catch_grid |>
      filter(lon_bin < bbox["xmin"] - 2 | lon_bin > bbox["xmax"] + 2 |
             lat_bin < bbox["ymin"] - 2 | lat_bin > bbox["ymax"] + 2)
    if (nrow(out_of_extent) > 0) {
      warning(country_cd, ": ", nrow(out_of_extent),
              " grid cell(s) fall outside the EEZ map extent and will not be visible.")
    }

    # ── Fill values and labels from the lookup
    fill_values <- setNames(sp_colours[species_map], names(species_map))
    fill_labels <- setNames(species_map, names(species_map))

    # ── Axis breaks: ~8 longitude labels whatever the EEZ width, labelled
    #    conventionally either side of the dateline; latitude labels are
    #    needed explicitly now that the CRS (and its graticule) is gone.
    x_span <- as.numeric(bbox["xmax"] - bbox["xmin"])
    step   <- max(1, round(x_span / 8))
    xbr    <- seq(floor(bbox["xmin"]), ceiling(bbox["xmax"]), by = step)
    lon_lab <- function(x) {
      x <- x %% 360
      ifelse(x == 180, "180\u00B0",
             ifelse(x > 180, paste0(360 - x, "\u00B0W"),
                    paste0(x, "\u00B0E")))
    }
    lat_lab <- function(y) {
      ifelse(y < 0, paste0(abs(y), "\u00B0S"),
             ifelse(y > 0, paste0(y, "\u00B0N"), "0\u00B0"))
    }

    # ── Aspect ratio: with no CRS, coord_sf assumes projected coordinates
    #    (1 deg lon = 1 deg lat on screen). Correct for longitude
    #    convergence at the EEZ's mid-latitude so shapes aren't stretched
    #    E-W — matters for the higher-latitude PICTs (CK, PF, NU, ...).
    mid_lat <- mean(c(bbox["ymin"], bbox["ymax"]))
    y_span  <- as.numeric(bbox["ymax"] - bbox["ymin"])
    asp     <- (y_span / x_span) / cos(mid_lat * pi / 180)

    p <- ggplot() +
      geom_sf(data = land, fill = "grey90", colour = "grey60", linewidth = 0.2) +
      geom_sf(data = eez,  fill = "grey50", colour = "black", alpha = 0.4) +
      geom_scatterpie(
        data   = catch_grid,
        aes(x = lon_bin, y = lat_bin, r = radius),
        cols   = names(species_map),
        colour = NA,
        alpha  = 0.9
      ) +
      scale_fill_manual(
        values = fill_values,
        labels = fill_labels,
        name   = "Species"
      ) +
      scale_x_continuous(breaks = xbr, labels = lon_lab) +
      scale_y_continuous(labels = lat_lab) +
      coord_sf(
        xlim   = c(bbox["xmin"], bbox["xmax"]),
        ylim   = c(bbox["ymin"], bbox["ymax"]),
        expand = FALSE
      ) +
      labs(x = "Longitude", y = "Latitude") +
      theme_bw() +
      theme(legend.position = "bottom",
            aspect.ratio = asp,
            axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 10),
            axis.text.y = element_text(size = 10),
            panel.background = element_rect(fill = "white"))

    print(p)

  } else {
    message(country_cd, ": no positive-catch grid cells — longline spatial figure skipped.")
  }
}
```

\newpage

## Annual catch estimates - artisanal

@tbl-raised-ace presents the raised and unraised estimates per species for `r report_year`, and @tbl-raised-ace-perc shows the percentage contribution of each species. Data is sourced from Tails and Ikasavea (landing surveys). Raised estimates are calculated based on the local knowledge of average monthly trips per landing sites.

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: Monthly-effort
#| tbl-cap: !expr if(nrow(data_3615) > 0) glue::glue("Number of trips per month and landing site reported in Tails") else ""

# Check if data exists
if (!(nrow(data_3615) == 0)) {
  
  names(data_3615) <- gsub(x = names(data_3615), pattern = "_trips$", replacement = "")  
  
  data_3615 <- data_3615 |>
    rename_with(~ month.abb[match(., tolower(month.abb))], 
              .cols = any_of(tolower(month.abb))) |> 
    select(-entity)
  
  flextable(data_3615) |>  
  set_header_labels(landing_site_name = "Landing Site") |>
  width(j=1, width=1.2) |> 
  fontsize(size = 8, part = "all") |>
  fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: Monthly-effort-ika
#| tbl-cap: !expr if(exists('data_ika_trips')) glue::glue("Number of trips per month and landing site reported in Ikasavea") else ""
if (exists('data_ika_trips')){
  flextable(df_ika_wide) |> 
  set_header_labels(landing_site_name = "Landing Site") |>

  width(j=1, width=1.2) |> 
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
  }
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr both_exist

if (both_exist) {
  cat("### Differences in Landing Sites: Ikasavea + Tails\n\n")
  cat("Artisanal data is collected using Tails and in Ikasavea platforms. The summary table below of all landing sites common to T2 and Ikasavea as well the ones unique to each of these sources (if any).")
}
```

```{r}
#| echo: false
#| include: false
#| eval: !expr both_exist
#| results: 'asis'
#| label: diff_t2_ika

if (both_exist) {

  ls_t2  <- tolower(trimws(unique(data_3615$landing_site_name)))
  ls_ika <- tolower(trimws(unique(df_ika_wide$landing_site_name)))
  
  in_both    <- intersect(ls_t2, ls_ika)
  only_t2    <- setdiff(ls_t2, ls_ika)
  only_ika   <- setdiff(ls_ika, ls_t2)
  
  if (length(only_t2) == 0 && length(only_ika) == 0) {
    cat("✅ All landing site names match perfectly between both sources.\n")
  } else {
    
    # Summary table
    summary_df <- data.frame(
      ` ` = c("Matched in both", "Only in T2 (data_3615)", "Only in Ikasavea"),
      `Count` = c(length(in_both), length(only_t2), length(only_ika)),
      check.names = FALSE
    )
    
    # cat("### Landing site name comparison\n\n")
      flextable(summary_df) |>  
        set_header_labels(landing_site_name = "Landing Site") |>
         width(j=1, width=1.2) |> 
        fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)

    
    # Table of mismatches
    if (length(only_t2) > 0 || length(only_ika) > 0) {
      max_len <- max(length(only_t2), length(only_ika))
      mismatch_df <- data.frame(
        `Only in T2 (data_3615)` = c(only_t2,  rep("", max_len - length(only_t2))),
        `Only in Ikasavea`       = c(only_ika, rep("", max_len - length(only_ika))),
        `In both` = c(in_both, rep("", max_len - length(in_both))),
        check.names = FALSE
      )
      # cat("\n### Unmatched landing site names\n\n")
      flextable(mismatch_df) |>  
        set_header_labels(landing_site_name = "Landing Site") |>
         width(j=1, width=1.2) |> 
        fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
    }
  }
}

```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr both_exist

if (both_exist) {
  cat("### Trips: Ikasavea + Tails\n\n")
  cat("The final number of trips combining what was reported on these two platforms is presented in the table below.")
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: tails_ika_summary

# ika and tails combined
if (both_exist) {
  prep <- function(df) {
    df %>%
      mutate(landing_site_name = tolower(trimws(landing_site_name))) %>%
      pivot_longer(-landing_site_name, names_to = "month", values_to = "trips") |>
      mutate(month = tolower(gsub("_trips", "", month)))
  }
  
  combined_t2_ika <- bind_rows(prep(data_3615), prep(df_ika_wide)) %>%
    group_by(landing_site_name, month) %>%
    summarise(trips = sum(trips), .groups = "drop") %>%
    mutate(month = factor(month, levels = tolower(month.abb))) %>%
    pivot_wider(names_from = month, values_from = trips) %>%
    arrange(landing_site_name) |>
    select(landing_site_name, tolower(month.abb))
  
  ft <- flextable(combined_t2_ika) |>  

    width(j=1, width=1.2) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
  
  # flextable::htmltools_value(ft)
  ft

} else {
  combined_t2_ika <- data.frame()
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr nrow(data_3615) > 0 && file.exists(local_knowl_trips_file)

if (nrow(data_3615) > 0 && file.exists(local_knowl_trips_file)) {
  cat("### Trips: Local Knowledge\n\n")
  cat("Member-estimated **trips** per month at each landing site are presented in the table below. These estimates are used to compare with what is reported in the ereportng tools to produce a 'best' estimate of effort.")
}
```

```{r trips-est-mem}
#| echo: false
#| include: false
#| results: 'asis'
#| label: mem-trips
#| tbl-cap: !expr if(nrow(data_3615) > 0) glue::glue("Member reported estimated trips per month per landing site") else ""

if (file.exists(local_knowl_trips_file)){
  local_knowl_trips = read.csv(local_knowl_trips_file)
  flextable(local_knowl_trips)  |>  
    fontsize(size = 8, part = "all") |> 
    width(j=1, width=2) |>
        fit_to_width(max_width = 6.5)

}

```

```{r prep-ereporting_trips, echo = F}

# check logic to be applied for ereporting: tails, ikasavea or tails + ikasavea?

if (nrow(combined_t2_ika)>0){ # it means tails + ikasavea
  
  ereporting_trips <- combined_t2_ika |>
    mutate(landing_site_name = str_squish(landing_site_name)) |>
    data.frame()
  ereporting_trips_tag <- c("Trips: Ikasavea + Tails")
  
  }else if (!exists("data_ika_trips") && nrow(data_3615) > 0){ # so it means you only have tails
    ereporting_trips <- data_3615 |>
      mutate(landing_site_name = str_squish(landing_site_name)) |>
      data.frame()
    ereporting_trips_tag <- c("Trips: Tails")
      
  }else if(nrow(data_3615) == 0 && nrow(data_ika_trips) > 0){ # it means you only have ikasavea 
    ereporting_trips <- df_ika_wide |>
      data.frame()
    ereporting_trips_tag <- c("Trips: Ikasavea")
  }else{
    ereporting_trips <- data.frame()
    ereporting_trips_tag <- c("")
  }
```

```{r max.trips}
#| echo: false
#| include: false
#| results: 'asis'
#| label: est-trips-monthly
#| tbl-cap: !expr if(nrow(ereporting_trips) > 0) glue::glue("Best estimate of trips per month at each landing site") else ""
 
if (nrow(ereporting_trips) > 0){
  
  if (file.exists(local_knowl_trips_file)){
    
    # check if columns are correct, if not, flag the issue
    if (!all(c("landing_site_name", "est.trips") %in% colnames(local_knowl_trips))) {
      cat("Your estimated trip file has to have the following columns: 'landing_site_name', 'est.trips'")
      best_estimate_trips <- ereporting_trips
        
    }else{
      
      # if all columns are in the df, groom local_knowl_trips and produce best_estimate_trips
      local_knowl_trips <- local_knowl_trips |>
        mutate(across(where(is.character), tolower)) |>
        mutate(landing_site_name = str_squish(landing_site_name))

      best_estimate_trips = ereporting_trips |> 
        full_join(local_knowl_trips |> 
                    select(landing_site_name, est.trips))
      
      best_estimate_trips[is.na(best_estimate_trips)] <- 0
      best_estimate_trips <- data.frame(best_estimate_trips)
      for( i in 2:13){
        best_estimate_trips[ ,i] = ifelse( best_estimate_trips[ ,i] > best_estimate_trips[ ,14],  
                                           best_estimate_trips[ ,i] , best_estimate_trips[ ,14] )
        }

      # Get rid of sites that only have NAs
      best_estimate_trips$num_na_cols <- rowSums(best_estimate_trips[, 2:13] )

      best_estimate_trips <- best_estimate_trips |>  filter(num_na_cols>0) |>
        select(-c(est.trips, num_na_cols))
    }
    }
  else{
    cat(paste0("There is no member-estimated trips available, so the best estimate fishing is the table above: <b>", ereporting_trips_tag,
             "</b>"))    
    best_estimate_trips <- ereporting_trips
  }
  }

flextable(best_estimate_trips)  |>    
  set_header_labels(landing_site_name = "Landing Site") |>

  fontsize(size = 8, part = "all") |>
  fit_to_width(max_width = 6.5)

```

```{r prep-catch-data}
# ── Month number → abbreviated name (keeps columns in calendar order) ─────────
month_labels <- c("Jan","Feb","Mar","Apr","May","Jun",
                  "Jul","Aug","Sep","Oct","Nov","Dec")

# ── Helper: pivot one subset to wide format ───────────────────────────────────
make_wide <- function(data) {
  data %>%
    # Sum sp_kg for each landing_site × month combination
    group_by(landing_site_name, month) %>%
    summarise(sp_kg = sum(sp_kg, na.rm = TRUE), .groups = "drop") %>%
    # Convert month numbers to abbreviated names
    mutate(month = month_labels[month]) %>%
    # Ensure all 12 months are present before pivoting
    complete(
      landing_site_name,
      month = month_labels,
      fill  = list(sp_kg = 0)
    ) %>%
    # Pivot to wide
    pivot_wider(
      names_from  = month,
      values_from = sp_kg,
      values_fill = 0          # missing combos → 0
    ) %>%
    # Reorder columns so months appear Jan → Dec (only those present in data)
    select(landing_site_name, any_of(month_labels))
}

# helper to combine ika and tails
combine_wide <- function(df1, df2) {
  if (is.null(df1) && is.null(df2)) return(NULL)
  
  month_cols <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
  
  drop_zero_rows <- function(df) {
    if (is.null(df)) return(NULL)
    df |> filter(if_any(all_of(month_cols), ~ . != 0))
  }
  
  df1 <- drop_zero_rows(df1)
  df2 <- drop_zero_rows(df2)
  
  if (is.null(df1) || nrow(df1) == 0) return(df2)
  if (is.null(df2) || nrow(df2) == 0) return(df1)
  
  full_join(df1, df2, by = "landing_site_name", suffix = c("_a", "_b")) |>
    mutate(across(all_of(paste0(month_cols, "_a")), 
                  ~ coalesce(., 0L))) |>
    mutate(across(all_of(paste0(month_cols, "_b")), 
                  ~ coalesce(., 0L))) |>
    mutate(Jan = Jan_a + Jan_b, Feb = Feb_a + Feb_b,
           Mar = Mar_a + Mar_b, Apr = Apr_a + Apr_b,
           May = May_a + May_b, Jun = Jun_a + Jun_b,
           Jul = Jul_a + Jul_b, Aug = Aug_a + Aug_b,
           Sep = Sep_a + Sep_b, Oct = Oct_a + Oct_b,
           Nov = Nov_a + Nov_b, Dec = Dec_a + Dec_b) |>
    select(landing_site_name, all_of(month_cols))
}

# Check if tails data exists - if yes, reshape and create a df per sp
if (nrow(data_3614) > 0) {
  df_by_species <- data_3614 |>
    mutate(landing_site_name = str_squish(landing_site_name)) |>
    group_by(sp_code) %>%
    group_map(~ make_wide(.x), .keep = FALSE) %>%
    setNames(sort(unique(data_3614$sp_code))) 
  
  df_bet <- df_by_species[["bet"]]
  df_skj <- df_by_species[["skj"]]
  df_yft <- df_by_species[["yft"]]
}

# Check if ika data exists - if yeas, reshape and create a df per sp
if(exists('data_ika_catch_kg') && is.data.frame(get('data_ika_catch_kg'))){
  
  df_by_species_ika <- data_ika_catch_kg %>%
    mutate(landing_site_name = str_squish(landing_site_name)) |>

    group_by(sp_code) %>%
    group_map(~ make_wide(.x), .keep = FALSE) %>%
    setNames(sort(unique(data_ika_catch_kg$sp_code))) 
  
  df_bet_ika <- df_by_species_ika[["bet"]]
  df_skj_ika <- df_by_species_ika[["skj"]]
  df_yft_ika <- df_by_species_ika[["yft"]]
  
  
  # If required (if ika and tails exist), combine catch values from both sources
  if (nrow(data_3614) > 0){
    df_bet_total <- combine_wide(df_bet, df_bet_ika)
    df_skj_total <- combine_wide(df_skj, df_skj_ika)
    df_yft_total <- combine_wide(df_yft, df_yft_ika)
  } else{
    # Just use the ika source if t2 data doesn't exist
    df_bet_total <- df_bet_ika
    df_skj_total <- df_skj_ika
    df_yft_total <- df_yft_ika
    
  }
} else {
  # Just use catch from tails if ika data doesn't exist
  df_bet_total <- df_bet
  df_skj_total <- df_skj
  df_yft_total <- df_yft
}
```

```{r prep-unr-data}
# Calculate total effort for the year by landing site using the ereporting trips
tot_trips_ereport = data.frame(landing_site_name = ereporting_trips$landing_site_name,
                    tot_trips = rowSums(ereporting_trips[,-1],na.rm=T)) |>
                    mutate(landing_site_name = str_squish(landing_site_name))
# Reshape trips
best_estimate_trips_long <- best_estimate_trips |>
  janitor::clean_names() |>
  mutate(landing_site_name = str_squish(landing_site_name)) |>
  pivot_longer(cols = 2:13, names_to = "months", values_to = "tot_trips_best")

ereporting_trips_long <- ereporting_trips |>
  janitor::clean_names() |>
  mutate(landing_site_name = str_squish(landing_site_name)) |>
  pivot_longer(cols = 2:13, names_to = "months", values_to = "tot_trips_spc")

# Check if data exists
species = c("skj", "bet", "yft")

unr_list <- list()
raised_list <- list()
for (sp in species){
  df_name <- paste0("df_", sp, "_total")
  if (exists(df_name) && is.data.frame(get(df_name)) && nrow(get(df_name)) > 0) {
    df <- get(df_name)
    # Now calculate monthly average catch per trip
    avg_sp = data.frame(landing_site_name = df$landing_site_name,
                    tot_catch = rowSums(df[,-1],na.rm=T)) |>
                  left_join(tot_trips_ereport) |>
      mutate(cpd = round(tot_catch/tot_trips,0))
    
    unr_sp <- df  |> 
      left_join(avg_sp |> 
                  select(landing_site_name, 
                         cpd))
    
    # Add result to unraised list
    unr_list[[sp]] <- unr_sp # 
    
    # Calculate raised
    unr_sp_long <- unr_sp |>
      janitor::clean_names() |>
      mutate(landing_site_name = str_squish(landing_site_name)) |>
      pivot_longer(cols = 2:13, names_to = "months", values_to = "unr_catch")
  
    raised_df <- unr_sp_long |>
      left_join(ereporting_trips_long, by = c("landing_site_name", "months")) |>
      left_join(best_estimate_trips_long, by = c("landing_site_name", "months")) |>
      mutate(raised = round(case_when(
        is.na(tot_trips_spc) | tot_trips_spc == 0 ~ cpd * tot_trips_best,  # fallback: no sampling data
        .default = (unr_catch / tot_trips_spc) * tot_trips_best            # main: monthly CPUE × best trips
      ), 0))
            # mutate(raised = round(case_when(unr_catch == 0 & tot_trips_spc == 0 ~ cpd * tot_trips_best,
      #                           .default = (unr_catch/tot_trips_spc) * tot_trips_best), 0))

    raised_df_wide <- raised_df |>
      select(landing_site_name, months, raised) |>
      pivot_wider(id_cols = 1, names_from = "months", values_from = "raised")
    
    # Add result to raised list
    raised_list[[sp]] <- raised_df_wide
    
  }
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: data_skj
#| tbl-cap: !expr if ("skj" %in% names(unr_list)) glue::glue("Unraised skipjack catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("skj" %in% names(unr_list)) { 
    flextable(unr_list$skj|> 
                mutate(across(where(is.numeric), as.integer))) |>
    set_header_labels(landing_site_name = "Landing Site") |>
    colformat_num(j = 2:13, digits = 0) |>  
    width(j=1, width=1.2) |> 
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 17, unit = "cm")
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: df_yft_total
#| tbl-cap: !expr if ("yft" %in% names(unr_list)) glue::glue("Unraised yellowfin catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("yft" %in% names(unr_list)) { 
    flextable(unr_list$yft|> 
                mutate(across(where(is.numeric), as.integer))) |>
    set_header_labels(landing_site_name = "Landing Site") |>

    width(j=1, width=1.2) |> 
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 17, unit = "cm")
}

```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: df_bet_total
#| tbl-cap: !expr if ("bet" %in% names(unr_list)) glue::glue("Unraised bigeye catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("bet" %in% names(unr_list)) { 
    flextable(unr_list$bet|> 
                mutate(across(where(is.numeric), as.integer))) |>  
    set_header_labels(landing_site_name = "Landing Site") |>
    colformat_num(j = 2:13, digits = 0) |>  

    width(j=1, width=1.2) |> 
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 17, unit = "cm")
    }

```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: raise-skj
#| tbl-cap: !expr if ("skj" %in% names(raised_list)) glue::glue("Raised skipjack catch (kgs) reported in ereporting tools.") else ""

# Remove rows where no catch was reported - as no average exists...

if ("skj" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$skj 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_skj = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |>
    set_header_labels(landing_site_name = "Landing Site") |>

    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0

if (exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0) {
  cat("Missing skipjack catch with effort are reported below.\n\n")
}
```

```{r missing_skj, include=FALSE}
# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0){
  flextable(missing_unraised_skj |> select(-no_catch)) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

  colformat_num(j = 2:13, digits = 0) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: raise-yft
#| tbl-cap: !expr if ("yft" %in% names(raised_list)) glue::glue("Raised yellowfin catch (kgs) reported in ereporting tools.") else ""
# Remove rows where no catch was reported - as no average exists...

if ("yft" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$yft 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_yft = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |>
    set_header_labels(landing_site_name = "Landing Site") |>

    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0

if (exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0) {
  cat("Missing yellowfin catch with effort are reported below.\n\n")
}
```

```{r missing_yft, include=FALSE}

# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0){
  flextable(missing_unraised_yft |> select(-no_catch)) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

  colformat_num(j = 2:13, digits = 0) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: 'asis'
#| label: raise-bet
#| tbl-cap: !expr if ("bet" %in% names(raised_list)) glue::glue("Raised bigeye catch (kgs) reported in ereporting tools.") else ""

# Remove rows where no catch was reported - as no average exists...

if ("bet" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$bet 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_bet = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```

```{r}
#| echo: false
#| include: false
#| results: asis
#| eval: !expr exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0

if (exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0) {
  cat("Missing bigeye catch with effort are reported below.\n\n")
}
```

```{r missing_bet, include=FALSE}
# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0){
  flextable(missing_unraised_bet|> select(-no_catch)) |> 
    set_header_labels(landing_site_name = "Landing Site") |>

  colformat_num(j = 2:13, digits = 0) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)
}
```


```{r}
#| label: tbl-raised-ace
#| tbl-cap: Raised and unraised annual artisanal catch estimates in mt

ace <- data.frame()

for (sp in species) {
  
  if (sp %in% names(unr_list)) {
    
    unraised <- unr_list[[sp]] |>
      data.frame() |>
      ungroup() |>
      select(-c(1, 14)) |>
      summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) |>
      rowSums()
    
    raised <- raised_list[[sp]] |>
      data.frame() |>
      ungroup() |>
      select(-1) |>
      summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) |>
      rowSums()
    
  } else {
    unraised <- 0
    raised   <- 0
  }
  
  ace <- bind_rows(ace, data.frame(
    sp       = toupper(sp),
    unraised = round(unraised / 1000, 2),
    raised   = round(raised / 1000, 2)
  ))
}

# Produce final output
ace_final = ace |>
  gather(key='Method', value='ACE', -1) |>
  spread(key = sp, value=ACE)

flextable(ace_final) |>  
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)

```

```{r}
#| label: tbl-raised-ace-perc
#| tbl-cap: Percentage of the artisanal tuna ACE associated with each of the key species

ace_final$tot = rowSums(ace_final[,-1])
ace_final <- ace_final |>  mutate(SKJ = round(SKJ/tot*100,2),
                      YFT = round(YFT/tot*100,2),
                      BET = round(BET/tot*100,2)) |>
  mutate(across(BET:YFT, ~replace_na(.,0))) |>
  select(-tot)

flextable(ace_final) |>
  fontsize(size = 8, part = "all") |>
    fit_to_width(max_width = 6.5)

```


