---
date: today
date-format: "DD MMMM YYYY"
format: 
  pdf:
    mainfont: "Times New Roman"
    sansfont: "Times New Roman"
    tbl-cap-location: bottom
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"    
    number-sections: true
    highlight-style: github
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
knitr:
  opts_chunk:
    out-width: "100%"
    dev: "cairo_pdf"
page-layout: full
edit: source
params:
  country_code: 'NR'          
  year: 2025                  
  data_provided: 'Yes'        
  aceman: TRUE
  additional_sections: ""
execute: 
  eval: true
  echo: false
  warning: false
  message: false
---

```{r libs}
library(tidyverse)
library(flextable)
library(scales) 
library(maps)
library(sf)
library(stringr)

source("utils.R")

theme_set(theme_bw(base_family = "Times New Roman"))
flextable::set_flextable_defaults(font.family = "Times New Roman")

# key params
country_cd <- tolower(params$country_code)
report_year = params$year
this_yr_folder = paste0("./data/report_", as.character(report_year), "_", tolower(country_cd), "/") |>
  tolower()

'%nin%' = function(x,y) !x %in% y            

fmt <- function(x, digits = 0) {
  format(round(x, digits), big.mark = ",", nsmall = digits, scientific = FALSE, trim = TRUE)
}

```

\newpage

```{r header}
row1 <- glue::glue("Scientific data was provided to the Commission in accordance with the decision relating to the provision of scientific data to the Commission by 30 April {params$year+1}: [Yes/No]")
row2 <- "If no, please indicate the reason(s) and intended actions:"
row3 = row4 = row5 = ''

tab <- data.frame(lcol = c(row1, row2, row3, row4, row5), 
                  rcol = c(params$data_provided, rep("", 4)) )

ft <- flextable(tab)
# No header row
ft <- delete_rows(ft, i = 1, part = "header")
# Merge cells - there are a ton of merge_* functions - using this one for simplicity
ft <- merge_at(ft, i = 2:5, j = 1:2)
ft <- valign(ft, i = 2:5, valign = 'top') 
# Relative column width
ft <- width(ft, width = c(1, 0.1)) # Set relative width of columns
ft <- width(ft, width = dim(ft)$widths * 6.6 / (flextable_dim(ft)$widths)) 
ft <- border_outer(ft)
ft <- border_inner(ft)

ft
```

```{r}
#| label: prelims

# Define years to use for tables and figures
yrs_long <- (params$year-4):params$year
yr_range <- paste(min(yrs_long),max(yrs_long), sep="-")

# Map for plotting  
newmap = fortify(maps::map(wrap=c(0,360), plot=FALSE, fill=TRUE))  

# Create a vector of the key species to plot for each gear
ps_species <- c("skj", "yft", "bet", "alb", "pbf")
ll_species <- c("skj", "yft", "bet", "alb", "pbf", "blm", "bum", "mls", "swo")
pl_species <- c("skj", "yft", "bet", "alb", "pbf")

# Define species labels for each gear type
ps_species_labels <- c(
  alb = "Albacore",
  bet = "Bigeye tuna",
  skj = "Skipjack tuna",
  yft = "Yellowfin tuna",
  pbf = "Pacific bluefin tuna"
)

ll_species_labels <- c(
  alb = "Albacore",
  bet = "Bigeye tuna",
  skj = "Skipjack tuna",
  yft = "Yellowfin tuna",
  pbf = "Pacific bluefin tuna",
  swo = "Swordfish",
  blm = "Black marlin",
  bum = "Blue marlin",
  mls = "Striped marlin"
)

pl_species_labels <- c(
  alb = "Albacore",
  bet = "Bigeye tuna",
  skj = "Skipjack tuna",
  yft = "Yellowfin tuna",
  pbf = "Pacific bluefin tuna"

)

species_labels_all <-  c(
  alb = "Albacore",
  bet = "Bigeye tuna",
  skj = "Skipjack tuna",
  yft = "Yellowfin tuna",
  pbf = "Pacific bluefin tuna",
  swo = "Swordfish",
  blm = "Black marlin",
  bum = "Blue marlin",
  mls = "Striped marlin",
  ssp = "Shortbill spearfish",
  sfa = "Sailfish",
  bsh = "Blue shark",
  fal = "Silky shark",
  ocs = "Oceanic whitetip shark",
  
  mak = "Mako sharks (mak)",
  sma = "Mako sharks (sma)",
  lma = "Mako sharks (lma)",
  
  thr = "Thresher sharks (thr)",
  alv = "Thresher sharks (alv)",
  pth = "Thresher sharks (pth)",
  bth = "Thresher sharks (bth)",
  
  por = "Porbeagle shark",
  
  spn = "Hammerhead sharks (spn)",
  spk = "Hammerhead sharks (spk)",
  spl = "Hammerhead sharks (spl)",
  spz = "Hammerhead sharks (spz)",
  spq = "Hammerhead sharks (spq)",
  eub = "Hammerhead sharks (eub)",

  rhn = "Whale shark"
  
  )

country_names <- c(
  fj  = "Fiji",
  fm = "Federated States of Micronesia",
  gu = "Guam",
  ki = "Kiribati",
  mh = "Marshall Islands",
  mp = "Northern Mariana Is.",
  nc = "New Caledonia",
  nr = "Naoero",
  nu = "Niue",
  pf = "French Polynesia",
  pg = "Papua New Guinea",
  pn = "Pitcairn",
  pw = "Palau",
  sb = "Solomon Islands",
  tk = "Tokelau",
  to = "Tonga",
  tv = "Tuvalu",
  vu = "Vanuatu",
  wf = "Wallis and Futuna Islands",
  ck = "Cook Islands",
  ws = "Samoa"
)

# get country name
country_name <- country_names[country_cd] 

  
# Give species colours to maintain colour schemes throughout
# --This makes sure species colors are consistent in all plots

sp_colours <- setNames(object = c("seagreen3","red3","black","cyan1","steelblue","orange",
                                  "pink","wheat","purple3","grey40","dodgerblue4","purple4",
                                  "#66CCFF","salmon4","magenta3","gold"),
                         nm = c("Albacore","Bigeye tuna","Black marlin","Blue marlin","Blue shark",
                                "Hammerhead sharks nei","Mako sharks","Oceanic whitetip shark",
                                "Pacific bluefin tuna","Silky shark","Skipjack tuna",
                                "Striped marlin","Swordfish","Thresher sharks nei",
                                "Whale shark","Yellowfin tuna"))
  
```

\newpage

# Abstract 

The major tuna species caught from the foreign fishing vessels in the Vanuatu EEZ in 2024 was dominated by 65% of
albacore, 21 % of yellowfin, 3% of bigeye and lastly 11% for others species of the total catch. In 2024, there was a
noticeable increase in the catch compared to 2023 from 4642.984 metric tons of fish recorded to 6991.41metric tonnes of
fish. This increase in catch can be attributed to a combination of factors, including improved fishing conditions, favorable
environmental factors, and potentially increased fishing effort.
In 2014, the total catch recorded was 6,622.21 metric tons (MT). In 2015, the catch slightly decreased to 6,605.78 MT. This
was followed by an increase in 2016 to 9,918.50 MT, and a further rise in 2017 to 13,232.26 MT. However, in 2018, the
total catch declined to 7,594.43 MT, before increasing again in 2019 to 8,939.64 MT. Between 2016 and 2017, the
significant increase in catch was largely due to fishing vessels returning to operate within the Vanuatu Exclusive Economic
Zone (EEZ). This shift contributed to fluctuations in catch volumes during those years, influenced by factors such as access
to onshore facilities, availability of markets, and favourable fishing patterns.
From 2020 to 2022, catch volumes continued to fluctuate. In 2020, the total catch was 7,996 MT. The lowest catch during
this period occurred in 2021, with only 3,716 MT recorded, mainly due to the severe impacts of the COVID-19 pandemic on
fishing operations. In 2022, catches rebounded slightly to 7,257 MT, a recovery attributed to the normalization of operations
following the easing of COVID-19 restrictions. As economic activities resumed and fishing effort increased, catch volumes
showed improvement.
In the period 2020 – 2024 the annual catch estimates of the Vanuatu longline fleets in the WCPO showed an increase from
catches between 2020 to 2021 with a total estimate of 8,809Mt to 10,205Mt and a decrease from 2021 to 2022 with 2022
catches at 7,838.95Mt. This reduction continued on to 2023 where a record of 6,741.88Mt was recorded, this amount being
the lowest in the last 5 years and which was due to the reduction in vessel effort since where vessel numbers was lowest in
this period. In 2024 however, vessel numbers and effort improved and an increase in catch was seen at 7,425.95MT from
2023 levels.
Purse seiners on the other hand experienced an increase in catch estimates between the years 2020 to 2021 due to the
increase in vessel effort and recording a total of 47,292Mt in 2020 and 48,597Mt in 2021. This number further increased in
2022 to 68,990MT recording the highest in the five years and being attributed to an increase in vessel numbers and effort. In
2023 however, the catch decreased to 53,907Mt and further reduced to 27,682MT in 2024 as vessel effort reduced. This
catch was dominated by skipjack making up 84% of the catch followed by Yellowfin at 14% and lastly bigeye at 0.8% of
the catch in 2024.
Raised 2024 data shows that catches of the main tuna species for Purse seines reduced evidently from 39,859Mt of skipjack
in 2020 to 23,397Mt in 2024, the lowest caught for this species. Longline vessels also experienced a decrease in catches of
Albacore as a main tuna species for this fishery with a catch total of 4,970Mt in 2020 to 2,740Mt in 2024, the lowest
recorded for the 5 years.

Since 2014, locally based vessel operation had ceased as vessels were based entirely in Fiji and the Solomon Islands,
however in 2019, locally based vessels moved back to fish in the Vanuatu EEZ where unloading of fish was experienced
from 6 locally based foreign fishing vessels offloading fish into the Fish processing facility in Port Vila. These operations
were continued to 2024 when the vessels were recalled for maintenance work before proceeding operations in 2025.

# Background

The main commercial tuna and billfish species caught in the Vanuatu EEZ and by the Vanuatu fleet in the WCPFC consists
of albacore (Thunnus alalunga), bigeye (Thunnus obesus), skipjack (Katsuwonus pelamis), yellowfin (Thunnus albacares),
black marlin (Makaira indica), blue marlin (Makaira nigricans), striped marlin (Tetrapturus audax) and swordfish (Xiphias
gladius).

As part of Vanuatu’s obligation to report the WCPFC CMMS’s for key shark species, data has also been compiled, some of
which are now covered in the longline fleet tables, these are blue shark (Prionace glauca), silky shark (Carcharhinus
falciformis), oceanic whitetip shark (Carcharhinus longimanus) and mako shark (Isurus spp.). The main industrial fishing
methods employed in the Vanuatu EEZ has been dominated by the longline gear outside 24 miles, however a few Artisanal
fishers are found fishing within the 12 miles around FAD’s catching mainly Skipjack and Yellowfin. Individual fleets
presented herein cover vessels with high catch and effort data coverage and these are mostly Chinese fleets with a few Fiji
fleets. Other fleets such as Taiwainese fleets have cease operations over the last few years in the Vanuatu EEZ however, in
2021 a few Kiribati vessels were seen fishing in the Vanuatu EEZ. Fishing by these fleets is based through bilateral
agreements and the issuance of Foreign fishing licenses.

The report covers the fishing activities in the Vanuatu EEZ and operations of the Vanuatu flag vessels that were active in the
WCPFC and other broad ocean areas during the period 2017 to 2024. The report mainly focuses on the fleet structures,
annual catch estimates and catch/effort distributions. The report also raises areas where new and further effort is required on
the part of Vanuatu to enhance its role in contributing to the overall conservation and management of highly migratory
stocks in the WCPFC area. Most of the current presented data were obtained from the OFP-SPC DORADO and Tufman2
database and which were originally collected and verified by the Vanuatu Fisheries Data Management Unit (VFDMU).

Vanuatu recognizes that there are critical data ‘gaps’ that need attention and focus on. Therefore, with the limitation of
resources, the department has been working closely with SPC and FFA to collect as much information and data as possible
to fill in these gaps. The delegation of designated ports recognized as PSMA countries for our Flagged Vessels have been
established however are yet to be implemented and these will enable us to monitor landings of fish in foreign ports including
those in Suva, Levuka and Pagopago which are currently the ports mainly being utilized. In 2023 and 2024 a few countries
like Thailand, have begun sharing landing data as part of their PSMA obligations to Vanuatu and these data are useful in
verification of catch data provided.

# Flag State Reporting

Vanuatu is currently a member of WCPFC, IATTC, SPRFMO and has ratified the NPFC. The membership of Vanuatu in
these RFMOs has enabled Vanuatu’s fishing fleet to fish these RFMO’s waters for tuna and other highly migratory fish
species. The Vanuatu fleet consists of 2 purse seiners and 52 long-liners with licenses to fish in the WCPFC area in 2024.
The Vanuatu fleet consists of purse seine and longline vessels fishing between the Pacific and Indian Ocean. Fishing inside
the Exclusive Economic Zones (EEZ) of coastal states had been possible by way of Bilateral Fishing Access (BFA) for both
longlines and purse seiners. Vanuatu operates a vessel registry, the Vanuatu International Shipping Registry (VISR). The
VISR has recorded over 90 vessel registrations since 2014, and currently there is a total of 95 vessels on the Vanuatu
registry and these comprise of 56 Longlines, 14 Refrigerated cargo ship,14 Trawlers, 7 Purse seines, 4 Squid jiggers and 1
oil bunker. It is a requirement through the Vanuatu Fisheries Act that all Vanuatu fishing vessels acquire an International
Authorization to Fish Certificate (IATF) in order to operate in the high seas within the Pacific Ocean.

## Catch and Effort Trends

The annual catch and effort estimates have been estimated for the Vanuatu fleet operating under bilateral arrangements and
the large-scale longline vessels (LSLV) operating in the wider WCPFC Area. The general observation since 2020 was that
there has been a variation in the annual catch and effort estimates for both the purse seine and the longline fleet.
The major tuna species for the Vanuatu longline fleet catch was dominated by albacore then bigeye and lastly yellowfin.
Raised estimates for the longline fleet in 2024 were 2,740.58 MT for albacore, 2,401.69MT for bigeye and 1,266.19MT for
yellowfin respectively and these catch estimates were determined from logsheet data raised using information on actual
vessel Activity (VMS data). During the period 2020-2024, the longline fleet recorded its highest total annual catch estimate
as 10,205.13Mt in 2021 (Table 1(a)). The longline fishery recorded the highest catches for albacore in 2022 being
3,864.15Mt and the lowest in 2024 as 2,740.58Mt. The highest catch for bigeye was in 2021 with 2,650.49Mt and lowest in
2020 with 2,182Mt. Yellowfin catches however, showed an increase from 2020 to 2024 levels with the highest recorded in
2024 with 1,266Mt and the lowest in 2022 and 2023 with 689Mt and 528Mt respectively. Albacore continues to be the
dominant species in the catch for 2024 followed by bigeye and then yellowfin. Effort for the longline fishery has continued
to fluctuate in the recent years from 2020-2024 as can be seen in Table 5. This fluctuation in effort is evident through the
measure of the number of vessels licensed and number of days fished and sets deployed.

The purse seine fleet that operated under bilateral arrangements recorded a decrease in total catch from 2020, 2021 and
2022 levels with catch records as 47,292Mt, 48,597MT and 68,990Mt respectively. This amount reduced in 2023 to
53,907Mt and further reduced to 27,682Mt in 2024 (Table 1(b)). The effort in the total number of sets had also increased
from this period of 2020 to 2022 and reduced and 2024. During this period, the main tuna species in the catch being
Skipjack also showed an increase in catch from 39,859Mt in 2020 to 39,450Mt in 2021 and 57,952Mt in 2022. In 2023 and
2024, the catch for Skipjack reduced to 42,579Mt and 23,397Mt. This is also the same for the other two species Yellowfin
with the catch in 2020 being 6,172Mt and reduced in 2024 to 4,058Mt and for Bigeye also with the highest catch in 2020
being 1,260MT and lowest catch in 2024 being 227Mt.

The purse seine fleets were mainly operating within the 5 degrees North and 5 degrees South and between 150 degrees East
and 175 degrees West. The effort in the purse seine fishery is measured as days fishing and searching, Figures 2) b) shows
the catch distributions of purse seine vessels that operated under the bilateral agreements.
The longline effort is given as 100s of hooks. The longline efforts are distributed between 40 degrees North and 40 degrees
South. This implies that both the southern and northern albacore stocks were targeted. However, there was more effort
experienced in the south i.e., between 10 degrees South and 40 degrees South and this effort has increased from 2020 to
2021 as effort increased in terms of vessels numbers and days fished and decreased in 2022 and this can be seen Figure 2) a)
where catch has increased for both the north and south pacific region targeting Albacore.

The catch and effort data coverage for the Vanuatu fleet are high, but the size data coverages are uncertain as most of these
vessels are landing their catch elsewhere and this would mostly be corroborated by the observers and port samplers in whose
jurisdictions catch may have been landed or transshipped. The inferences for high, medium, and low scores for the
catch/effort, and size data coverage, are provided in Appendix II.
Estimated Annual total catches of non-target, associated and dependent species by the Vanuatu purse seine fleets and longline fleets in 2020-2024 has been sought from the Tufman2 reporting database as shown in Table 3 and 4 and as well as in
Annex 1 where there is a summary table for all CMM’s concerned. However, due to the Covid-19 pandemic observer
placements were very low as repatriation took place, however, post covid and in 2024, observer placement on Vanuatu flag
vessels has immensely improved with Vanuatu meeting its Observer coverage requirement for WCPFC.

Appendix 1 summary table also provides information on the observed species of interest (SSI) collected through observer
reports for the year 2024 by ROP observers on Purse seiners and by Vanuatu observers on the Vanuatu flag longliners.
Observer reports continue to be submitted to SPC for processing for all vessels being observed on by Vanuatu’s Observer
program. 


```{r ps-prep}

# Load all raw data - from t2
data_2953 <- safe_read_and_clean(this_yr_folder, country_cd, r_code = 2953)

# this is a quick fix, it is misleading to have a df called data_3605 when 
# data is being sourced from report 3619...
if(params$aceman){
  data_3605 <- safe_read_and_clean(this_yr_folder, country_cd, r_code = 3619)
  
  if("catch" %in% colnames(data_3605)){
    data_3605$unadjcatch <- data_3605$catch
  } 
  
}else{
  data_3605 <- safe_read_and_clean(this_yr_folder, country_cd, r_code = 3605)
}

# For 3608 with special post-processing
data_3608 <- safe_read_and_clean(
  this_yr_folder, 
  country_cd, 
  r_code = 3608,
  post_process = function(data) {
    data |>
      mutate(
        lon = ifelse(lon < 0, lon + 360, lon),
        lat = ifelse(gear_code == "s", lat + 0.5, lat + 2.5),
        lon = ifelse(gear_code == "s", lon + 0.5, lon + 2.5)
      )
  }
)

if (nrow(data_3605) == 0){
  data_3605_ps <- data.frame()
  data_3605_ll <- data.frame()
  data_3605_pl <- data.frame()
  }else{
    
 # only data until current report year
  data_3605 <- data_3605 |>
    mutate(year = as.numeric(year)) |>
    filter(year <= report_year)
  
  # separate in PS, LL and PL
  data_3605_ps <- data_3605 |>
    filter(gear_code == "s") |>
    filter(sp_code %in% ps_species) |>
    select(sp_code, year, catch)
  
  data_3605_ll <- data_3605 |>
    filter(gear_code == "l") |>
    filter(sp_code %in% ll_species) |>
    select(sp_code, year, catch)
  
  data_3605_pl <- data_3605 |>
    filter(gear_code == "pl") |>
    filter(sp_code %in% ll_species) |>
    select(sp_code, year, catch)
}

# shape for tables
shape_catch_tbl <- function(data) {
  data |>
    pivot_wider(names_from = year, values_from = catch, values_fill = 0) |>
    mutate(across(where(is.numeric), ~replace_na(., 0))) |>  # only needed if catch itself can be NA
    arrange(sp_code) |>
    rename(Species = sp_code) |>
    janitor::adorn_totals("row") |>
    select(Species, sort(tidyselect::peek_vars()))
}

if (nrow(data_3605_ps) > 0) data_3605_ps_tbl <- shape_catch_tbl(data_3605_ps)
if (nrow(data_3605_ll) > 0) data_3605_ll_tbl <- shape_catch_tbl(data_3605_ll)
if (nrow(data_3605_pl) > 0) data_3605_pl_tbl <- shape_catch_tbl(data_3605_pl)

# effort - number of hooks per year for the national fleet
data_3597 <- safe_read_and_clean(this_yr_folder, country_cd, r_code = 3597) 

```

```{r funcs-render}
#| echo: false

# Function to render gear-specific tables in section 3 with header and table
render_gear_tbl_section3 <- function(data, data_tbl, gear_name, gear_code, species_labels, intro_text = NULL)
  {
  
  # Check if data exists
  if (nrow(data) == 0) {
    return(invisible(NULL))
  }
  
  # Print header
  cat(glue::glue("## Annual catch estimates for the national {gear_name} fleet\n\n"))
  
  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }
  
  # Create and return the flextable
  ft <- flextable(data_tbl) |>
    labelizor(labels = species_labels, part = "all") |>
    set_table_properties(width = 0.7, align = "center") |> 
    bold(j = ncol(data_tbl), part = "body") |>
    width(j = 1, width = 2)
  return(ft)
}

# Function to render gear-specific figures in section 3 with header and plot
render_gear_figs_section3 <- function(data, gear_name, gear_code, species_list,
                                      species_labels, intro_text = NULL,
                                      effort_df = data_3597) {

  # Check if data exists
  if (nrow(data) == 0) {
    return(invisible(NULL))
  }

  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }

  # ── Prepare data for plotting ──────────────────────────────────────────────
  data_fig <- data |>
    group_by(sp_code, year) |>
    summarise(catch = sum(catch, na.rm = TRUE), .groups = "drop")

  ref_df <- expand_grid(year = as.numeric(yrs_long), sp_code = species_list)

  plot_data <- left_join(ref_df, data_fig, by = c("year", "sp_code")) |>
    mutate(
      catch   = replace_na(catch, 0),
      Species = factor(recode(sp_code, !!!species_labels),
                       levels = unname(species_labels[species_list]))
    )

  # ── Base plot (shared by all gears) ────────────────────────────────────────
  p <- ggplot(plot_data, aes(x = year, y = catch, fill = Species)) +
    geom_col(position = "stack") +
    scale_fill_manual(values = sp_colours) +
    scale_x_continuous(breaks = seq(min(plot_data$year), max(plot_data$year), 1)) +
    labs(x = NULL, y = "Catch (mt)") +
    theme_bw() +
    theme(
      axis.title   = element_text(size = 16),
      axis.text    = element_text(size = 16),
      axis.text.x  = element_text(angle = 45, hjust = 1),
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text  = element_text(size = 13)
    )

  # ── Longline: overlay effort (hooks) on a secondary axis ──────────────────
  if (gear_code == "l") {
    
    # maybe at some point this will fail in case effort_df doesnt exist or nrow = 0
     effort_annual <- effort_df |>
      filter(flag_id == country_cd) |>
      filter(year %in% plot_data$year) |>          # align with plotted years
      group_by(year) |>
      summarise(effort = sum(hooks_total_logsheet, na.rm = TRUE), .groups = "drop")

    max_catch  <- max(tapply(plot_data$catch, plot_data$year, sum))
    max_effort <- max(effort_annual$effort, 0)

    if (nrow(effort_annual) > 0 && max_effort > 0 && max_catch > 0) {

      scale_factor <- max_catch / max_effort

      p <- p +
        geom_line(data = effort_annual,
                  aes(x = year, y = effort * scale_factor),
                  inherit.aes = FALSE, colour = "darkorange",
                  linewidth = 1.2, linetype = "dashed") +
        geom_point(data = effort_annual,
                   aes(x = year, y = effort * scale_factor),
                   inherit.aes = FALSE, colour = "darkorange", size = 3) +
        scale_y_continuous(
          labels = comma,
          sec.axis = sec_axis(~ . / scale_factor,
                              name = "Effort (hooks)", labels = comma)
        ) +
        theme(
          axis.title.y.right = element_text(colour = "darkorange"),
          axis.text.y.right  = element_text(colour = "darkorange")
        )
    } else {
      message("No usable effort data for gear '", gear_code,
              "' — rendering catch only.")
      p <- p + scale_y_continuous(labels = comma)
    }

  } else {
    p <- p + scale_y_continuous(labels = comma)
  }

  print(p)
  invisible(NULL)
}

# Function to render vessel table for a specific gear
render_vessel_table <- function(ves_data, gear_name, size_categories_fn, intro_text = NULL) {
  
  # Filter for specific gear
  gear_data <- ves_data |> filter(gear_code == gear_name)
  
  # Check if data exists
  if (nrow(gear_data) == 0) {
    return(invisible(NULL))
  }
  
  # Apply the size category function to summarize data
  ves_summary <- gear_data |>
    group_by(year, gear) |> 
    size_categories_fn()
  
  # Check if there's any actual vessel data (not all zeros)
  if (sum(ves_summary[,-c(1:2)]) == 0) {
    return(invisible(NULL))
  }
  
  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }
  
  # Create and return flextable
  ves_summary |> 
    pivot_longer(-c(year, gear), names_to = "Size Category (GRT)", values_to = "Vessels") |>
    pivot_wider(names_from = year, values_from = Vessels, values_fill = 0) |>
    group_by(gear) |> 
    mutate(gear = ifelse(row_number() == 1, gear, "")) |>
    flextable() |>  
    set_table_properties(width = 0.7, align = "center") |> 
    width(j = 1:2, width = 1.2)
}

# Function to render spatial catch distribution map for a specific gear
render_spatial_map <- function(data, gear_code, gear_name, target_species, year_range, intro_text = NULL) {
  
  gear_code <- tolower(gear_code)
  
  # Filter and prepare data
  spatial_data <- data |>  
    filter(
      between(year, year_range[1], year_range[2]), 
      gear_code == !!gear_code, 
      sp_code %in% target_species
    ) |> 
    group_by(lat, lon, year, species) |> 
    summarise(catch = sum(catch), .groups = "drop") |> 
    filter(catch > 0)
  
  # Check if data exists
  if (nrow(spatial_data) == 0) {
    return(invisible(NULL))
  }
  
  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }
  
  # Create the map
  p <- ggplot() +
    geom_map(data = newmap, map = newmap, aes(map_id = region), fill = 'bisque1', col = 'gray') +
    geom_tile(data = spatial_data, aes(lon, lat, fill = catch)) +
    scale_fill_distiller(palette = "Spectral") +
    scale_x_continuous(limits = c(120, 230), expand = c(0, 0), breaks = seq(140, 220, 40)) +
    scale_y_continuous(limits = c(-35, 25), expand = c(0, 0)) +
    facet_grid(year ~ species) +
    xlab("Longitude") + ylab("Latitude") + labs(fill = "Catch (mt)") +
    theme_bw() +
    theme(
      axis.text = element_text(size = 18),
      legend.text = element_text(size = 18),
      legend.title = element_text(size = 18),
      strip.background = element_rect(fill = 'white'),
      strip.text = element_text(size = 18),
      axis.title = element_text(size = 20), 
      title = element_text(size = 20)
    ) +
    coord_sf() +
    ggtitle(glue::glue('Spatial patterns in catch for the {gear_name} fishery'))
  
  print(p)
  return(invisible(NULL))
}

# Function to render SSI (Species of Special Interest) table for a specific gear
render_ssi_table <- function(data, gear_code, gear_name, intro_text = NULL) {
  
  # Filter data for specific gear
  tab_data <- data |>  
    filter(gear == gear_code) |> 
    select(-gear) |>
    arrange(year)
  
  # Check if data exists
  if (nrow(tab_data) == 0) {
    return(invisible(NULL))
  }
  
   # Print header
  cat(glue::glue("### Observed catches of SSIs by the {gear_name} fleet\n\n"))
  
  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }

  # Create and return flextable
  flextable(tab_data) |>
    set_table_properties(layout = "autofit", width = 0.95, align = "center") |> 
    fontsize(size = 9, part = "all") |> 
    colformat_num(j = "year", big.mark = "") |>  
    width(j = c("category", "species"), width = 1.5) |> 
    align(align = "center", part = "all") |> 
    labelizor(
      labels = c(
        year = "Year",
        species = "Species",
        category = "Category",
        number = "Total",
        alive = "Alive",  # Fixed capitalization
        dead = "Dead"
      ), 
      part = "all"
    ) |>
    padding(padding.left = 5, padding.right = 5, part = "all")
}

```

```{r}
#| echo: false

yearly_total <- data_3605 |>
  group_by(year) |>
  summarise(sum = sum(catch), .groups = "drop")

total_5yr <- sum(data_3605$catch)
avg_total <- mean(yearly_total$sum, na.rm = TRUE)

avg_by_sp <- data_3605 |>
  group_by(sp_code, year) |>
  summarise(sum = sum(catch), .groups = "drop") |>
  group_by(sp_code) |>
  summarise(m = mean(sum, na.rm = TRUE), .groups = "drop")

avg_skj <- avg_by_sp$m[avg_by_sp$sp_code == "skj"]
avg_yft <- avg_by_sp$m[avg_by_sp$sp_code == "yft"]
avg_bet <- avg_by_sp$m[avg_by_sp$sp_code == "bet"]
```

`r country_name`'s Flag vessels for the past 5 years (@tbl-tbl1s) had caught a total of `r fmt(total_5yr)` mt of tuna at an average of `r fmt(avg_total)` mt per year. The average breakdown of catches by species are; SKJ at `r fmt(avg_skj)` mt, YFT at `r fmt(avg_yft)` mt, and BET at `r fmt(avg_bet)` mt per year.


```{r}
#| echo: false
#| results: 'asis'
#| label: tbl-tbl1s
#| tbl-cap: !expr if(nrow(data_3605_ps) > 0) glue::glue("Annual catch estimates for the {country_name} purse seine fleet, by primary species and year, in the WCPFC Convention Area. Source - Tufman 2, report 3605.") else ""

render_gear_tbl_section3(
  data = data_3605_ps,
  data_tbl = data_3605_ps_tbl,
  gear_name = "purse seine",
  gear_code = "s",
  species_labels = ps_species_labels,
  intro_text = ""
)
```

```{r}
#| echo: false
#| results: 'asis'
#| label: tbl-tbl1l
#| tbl-cap: !expr if(nrow(data_3605_ll) > 0) glue::glue("Annual catch estimates for the {country_name} longline fleet, by primary species and year, in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605.") else ""

render_gear_tbl_section3(
  data = data_3605_ll,
  data_tbl = data_3605_ll_tbl,
  gear_name = "longline",
  gear_code = "l",
  species_labels = ll_species_labels,
  intro_text = ""
)
```

```{r}
#| echo: false
#| results: 'asis'
#| label: tbl-tbl1pl
#| tbl-cap: !expr if(nrow(data_3605_pl) > 0) glue::glue("Annual catch estimates for the {country_name} pole line fleet, by primary species and year, in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605.") else ""

render_gear_tbl_section3(
  data = data_3605_pl,
  data_tbl = data_3605_pl_tbl,
  gear_name = "pole line",
  gear_code = "pl",
  species_labels = ll_species_labels,
  intro_text = ""
)
```

<!-- \newpage -->

```{r}
#| echo: false
#| results: 'asis'
#| fig-pos: 'H'
#| label: fig-fig1l
#| fig-cap: !expr if(nrow(data_3605_ll) > 0) glue::glue("Historical annual catch estimates (mt) and effort (no. of hooks) for the {country_name} longline fleet, by primary species and year, in the WCPFC Convention Area for {yr_range}. Source - Tufman 2, report 3605 and 3597 (hooks).") else ""
#| fig-width: 12
#| fig-height: 7
#| fig-align: "center"

render_gear_figs_section3(
  data = data_3605_ll,
  gear_name = "longline",
  gear_code = "l",
  species_list = ll_species,
  species_labels = ll_species_labels,
  intro_text = ""
)
```

<!-- \newpage -->

```{r}
#| echo: false
#| results: 'asis'
#| fig-pos: 'H'
#| label: fig-fig1s
#| fig-cap: !expr if(nrow(data_3605_ps) > 0) glue::glue("Historical annual catch estimates (mt) for the {country_name} purse seine fleet, by primary species and year, in the WCPFC Convention Area for {yr_range}. Source - Tufman 2, report 3605.") else ""
#| fig-width: 12
#| fig-height: 7
#| fig-align: "center"

# Modified from original: 

render_gear_figs_section3(
  data = data_3605_ps,
  gear_name = "purse seine",
  gear_code = "s",
  species_list = ps_species,
  species_labels = ps_species_labels,
  intro_text = ""
)
```

\newpage

```{r}
#| echo: false
#| results: 'asis'
#| fig-pos: 'H'
#| label: fig-fig1pl
#| fig-cap: !expr if(nrow(data_3605_pl) > 0) glue::glue("Historical annual catch estimates (mt) for the {country_name} pole line fleet, by primary species and year, in the WCPFC Convention Area for {yr_range}. Source - Tufman 2, report 3605.") else ""
#| fig-width: 12
#| fig-height: 7
#| fig-align: "center"

render_gear_figs_section3(
  data = data_3605_pl,
  gear_name = "pole line",
  gear_code = "pl",
  species_list = pl_species,
  species_labels = pl_species_labels,
  intro_text = ""
)
```

```{r}
#| echo: false
#| results: asis
#| eval: !expr nrow(data_3605) > 0

if (nrow(data_3605) > 0) {
  cat("\\newpage\n\n")
  cat("## Historical annual vessel numbers for the national fleet\n\n")
}
```

@fig-fig2 displays the total number of active vessels in `r country_name`’s national fleet from `r report_year - 4` to `r report_year`.

```{r}
#| label: fig-fig2
#| fig-pos: 'H'
#| eval: !expr nrow(data_3605) > 0
#| fig-cap: !expr if(nrow(data_3605) > 0) glue::glue("Historical annual vessel numbers for the {country_name} fleet, by gear, for the WCPFC Convention Area. Source - Tufman 2, report 3605.") else ""
#| fig-width: 12
#| fig-height: 7
#| fig-align: "center"

if (nrow(data_3605) > 0){
    # Plot of vessel numbers for only Longline, Purse seine and Pole Line, aggregated over size class
    vessel_colours <- c(
    "0 - 50"    = "#A1DAB4",
    "51 - 200"  = "#41B6C4",
    "201 - 500" = "#2C7FB8",
    "500+"      = "#253494"
  )
  
   ves_raw <- 
    data_3605 |>  
    filter(year %in% yrs_long, gear_code %in% c("s", "l", "pl")) |> 
    select(flag, year, gear_code, vessels_sml, vessels_med, vessels_lge, vessels_hge) |> 
    distinct() |>
    mutate(gear = case_when(
      gear_code == "s" ~ "Purse seine",
      gear_code == "l" ~ "Longline",
      gear_code == "pl" ~ "Pole line"
    ))
  
  vessel_plot <- ves_raw |>
    pivot_longer(
      cols = c(vessels_sml, vessels_med, vessels_lge, vessels_hge),
      names_to = "vessel_class",
      values_to = "count"
    ) |>
    mutate(vessel_class = recode(vessel_class,
                                 "vessels_sml" = "0 - 50",
                                 "vessels_med" = "51 - 200",
                                 "vessels_lge" = "201 - 500",
                                 "vessels_hge" = "500+"
    )) |>
    mutate(vessel_class = factor(vessel_class, levels = c("0 - 50", "51 - 200", "201 - 500", "500+")))
  
  p_vessels <- ggplot(vessel_plot, aes(x = year, y = count, fill = vessel_class)) +
    geom_col(position = "stack") +
    scale_fill_manual(values = vessel_colours) +
    scale_x_continuous(breaks = seq(min(vessel_plot$year), max(vessel_plot$year), 1)) +
    scale_y_continuous(breaks = scales::pretty_breaks()) +
    labs(x = NULL, y = "Vessels", fill = "Size category (GRT)") +
    theme_bw() +
    facet_wrap(~gear) +
    theme(
    axis.title   = element_text(size = 16),
    axis.text    = element_text(size = 16),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    strip.text   = element_text(size = 16),
    legend.position = "bottom",
    legend.title = element_text(size = 13),
    legend.text  = element_text(size = 13)
  )
  
  print(p_vessels)
} else {
  ves_raw <- tibble()
}
```


```{r}

#| echo: false

# Prepare vessel data once
if(nrow(data_3605) > 0){
  ves_raw <- 
    data_3605 |>  
    filter(year %in% yrs_long, gear_code %in% c("s", "l", "pl")) |> 
    select(flag, year, gear_code, vessels_sml, vessels_med, vessels_lge, vessels_hge) |> 
    distinct() |>
    mutate(gear = case_when(
      gear_code == "s" ~ "Purse seine",
      gear_code == "l" ~ "Longline",
      gear_code == "pl" ~ "Pole line"
    ))
} else {
  ves_raw <- tibble()
}

# Define size category functions for each gear type
ll_size_fn <- function(data) {
  data |>
  summarise(
    '0 - 50' = sum(vessels_sml),
    '51 - 200' = sum(vessels_med),
    '201 - 500' = sum(vessels_lge),
    '500+' = sum(vessels_hge),
    .groups = "drop"
  )
}

ps_size_fn <- function(data) {
  data |>
  summarise(
    '0 - 50' = sum(vessels_sml),
    '51 - 200' = sum(vessels_med),
    '201 - 500' = sum(vessels_lge),
    '500+' = sum(vessels_hge),
    .groups = "drop"
  )
}

pl_size_fn <- function(data) {
  data |>
  summarise(
    '0 - 50' = sum(vessels_sml),
    '51 - 150' = sum(vessels_med),
    '150+' = sum(vessels_lge + vessels_hge),
    .groups = "drop"
  )
}

```

```{r}
ll_caption <- if(nrow(ves_raw) > 0 && "gear_code" %in% names(ves_raw) && nrow(ves_raw |> filter(gear_code == "l")) > 0) {
  glue::glue("Number of {country_name} longline vessels, by size category, active in the WCPFC Convention Area, over {yr_range}. Source - Tufman 2, report 3605.")
} else NULL
```

```{r}
#| label: tbl-tbl2-longline
#| eval: !expr "!is.null(ll_caption)"
#| tbl-cap: !expr ll_caption
#| echo: false
#| results: 'asis'
render_vessel_table(
  ves_data = ves_raw,
  gear_name = "l",
  size_categories_fn = ll_size_fn,
  intro_text = ""
)
```

```{r}
ps_caption <- if(nrow(ves_raw) > 0 && "gear_code" %in% names(ves_raw) && nrow(ves_raw |> filter(gear_code %in% c("s", "ps"))) > 0) {
  glue::glue("Number of {country_name} purse seine vessels, by size category, active in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605.")
} else NULL
```

```{r}
#| label: tbl-tbl2-ps
#| eval: !expr "!is.null(ps_caption)"
#| tbl-cap: !expr ps_caption
#| echo: false
#| results: 'asis'

render_vessel_table(
  ves_data = ves_raw,
  gear_name = "s",
  size_categories = ps_size_fn,
  intro_text =  ""
)
```

```{r}
#| echo: false
#| results: asis
#| eval: !expr nrow(data_3608) > 0

if (nrow(data_3605) > 0) {
  cat("\\newpage\n\n")
  cat("## Distribution of catches of target species for different national fisheries\n\n")
}
```

In this section we present the spatial distribution of catch (in metric tonnes) by species, 
Albacore, Bigeye tuna, and Yellowfin tuna , from `r country_name`’s fleet across the period `r report_year - 4`-
`r report_year`.

@fig-figmapps highlights key fishing areas in the Western and Central Pacific Ocean, 
primarily concentrated within and just outside `r country_name`’s Exclusive Economic Zone (EEZ).

```{r}
#| label: fig-figmapps
#| fig-cap: !expr glue::glue("Map showing the distribution of target species catch (mt) for the {country_name} purse seine fishery. Source - Tufman 2, report 3608.")
#| fig-pos: "H"
#| echo: false
#| fig-width: 12
#| fig-height: 8

map_ps <- file.path(this_yr_folder, glue::glue("map_ps_{country_cd}_{params$year}.png"))
if (file.exists(map_ps)) 
    knitr::include_graphics(map_ps)
```

```{r}
#| label: fig-figmapl
#| fig-cap: !expr glue::glue("Map showing the distribution of target species catch (mt) for the {country_name} longline fishery. Source - Tufman 2, report 3608.")
#| fig-pos: "H"
#| echo: false
#| fig-width: 12
#| fig-height: 8

map_ll <- file.path(this_yr_folder, glue::glue("map_ll_{country_cd}_{params$year}.png"))
if (file.exists(map_ll)) knitr::include_graphics(map_ll)
```

```{r}
#| label: fig-figmappl
#| eval: !expr "nrow(data_3608) > 0"
#| fig-cap: !expr if(nrow(data_3608 |> filter(gear_code == "pl") > 0)) glue::glue("Map showing the distribution of target species catch (mt) for the {country_name} pole line fishery. Source - Tufman 2, report 3608.") else ""
#| fig-width: 12
#| fig-height: 8
#| fig-align: "center"
#| fig-pos: "H"
#| echo: false
#| message: false
#| warning: false
#| results: 'asis'

render_spatial_map(
  data = data_3608,
  gear_code = "pl",
  gear_name = "pole line",
  target_species = c("skj", "yft", "bet"),
  year_range = c(params$year - 4, params$year),
  intro_text = ""

)
```

```{r}
#| echo: false
#| results: asis
#| eval: !expr nrow(data_2953) > 0 && "gear" %in% names(data_2953) && nrow(data_2953 |> filter(gear == "s")) > 0

if (nrow(data_3605) > 0) {
  cat("\\newpage\n\n")
  cat("## Captures of species of special interest (SSIs)\n\n")
  cat("Note that the tables below are specified to include - seabird, turtle and marine mammals in the instructions but we also include things like some rays and sharks. Note that Alive plus Dead does not necessarily equal Total due to unknowns. ")
}
```

@tbl-tbl3s provides a summary of observed interactions with special interest species by `r country_name`’s purse seine fleet from `r report_year - 4` to `r report_year`. In case of missing years in the table below, this indicates that no interactions with target species were reported during those years.

```{r}
ssi_s_caption <- if(nrow(data_2953) > 0 && "gear" %in% names(data_2953) && nrow(data_2953 |> filter(gear == "s")) > 0) {
  glue::glue("Observed annual estimated catches of species of special interest (seabird, turtle and marine mammals) by the {country_name} purse seine fleet, in the WCPFC Convention Area, for years {yr_range}, to the extent possible. Note that the Total may not reflect the sum of Alive and Dead as some animals are reported with an Unknown status. Source - Tufman 2, report 2953.")
} else NULL
```

```{r}
#| label: tbl-tbl3s
#| eval: !expr "!is.null(ssi_s_caption)"
#| tbl-cap: !expr ssi_s_caption
#| echo: false
#| results: 'asis'
render_ssi_table(
  data = data_2953,
  gear_code = "s",
  gear_name = "purse seine",
  intro_text = ""
)
```

```{r}
ssi_l_caption <- if(nrow(data_2953) > 0 && "gear" %in% names(data_2953) && nrow(data_2953 |> filter(gear == "l")) > 0) {
  glue::glue("Observed annual estimated catches of species of special interest (seabird, turtle and marine mammals) by the {country_name} longline fleet, in the WCPFC Convention Area, for years {yr_range}. Source - Tufman 2, report 2953.")} else NULL

```

```{r}
#| label: tbl-tbl3ll
#| eval: !expr "!is.null(ssi_l_caption)"
#| tbl-cap: !expr ssi_l_caption
#| echo: false
#| results: 'asis'
render_ssi_table(
  data = data_2953,
  gear_code = "l",
  gear_name = "longline",
  intro_text = ""
)
```

```{r}
ssi_pl_caption <- if(nrow(data_2953) > 0 && "gear" %in% names(data_2953) && nrow(data_2953 |> filter(gear == "pl")) > 0) {
  glue::glue("Observed annual estimated catches of species of special interest (seabird, turtle and marine mammals) by the {country_name} pole line fleet, in the WCPFC Convention Area, for years {yr_range}. Source - Tufman 2, report 2953.")} else NULL

```

```{r}
#| label: tbl-tbl3lpl
#| eval: !expr "!is.null(ssi_pl_caption)"
#| tbl-cap: !expr ssi_pl_caption
#| echo: false
#| results: 'asis'
render_ssi_table(
  data = data_2953,
  gear_code = "pl",
  gear_name = "pole line",
  intro_text = ""
)
```

```{r}
if (nrow(data_3605) > 0){
  PS_tab <- 
  data_3605 |>  
  filter(gear_code == "s", 
         year %in% yrs_long,
         !sp_code %in% ps_species) |>  # Exclude only the key species
  group_by(sp_code, year) |> 
  summarise(catch = sum(catch, na.rm = TRUE), .groups = "drop") |> 
  pivot_wider(names_from = year, values_from = catch, values_fill = 0) |> 
  arrange(sp_code)
  
  # Create table of non-target species for longline
  LL_tab <- 
    data_3605 |>  
    filter(gear_code == "l", 
           year %in% yrs_long,
           !sp_code %in% ps_species) |>  # Exclude only the key species
    group_by(sp_code, year) |> 
    summarise(catch = sum(catch, na.rm = TRUE), .groups = "drop") |> 
    pivot_wider(names_from = year, values_from = catch, values_fill = 0) |> 
    arrange(sp_code)

}else{
  PS_tab <- data.frame()
  LL_tab <- data.frame()
}
```

```{r}
#| label: tbl-tbl4s
#| eval: !expr "nrow(data_3605) > 0"
#| tbl-cap: !expr glue::glue("Annual estimated catches of non-target, associated and dependent species, including sharks, by the {country_name} purse seine fleet, in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605.")
#| echo: false
#| results: 'asis'

# Ensure required columns are present even when empty
required_cols <- c("sp_code", as.character((params$year - 4):params$year))
if (nrow(PS_tab) > 0) {
  # Print header
  cat(glue::glue("## Annual catch estimates for non-target, associated and dependent species for the national purse seine fleet\n\n"))
  
 cat(glue::glue(
    "@tbl-tbl4s provides a summary of observed interactions with non-target ",
    "species by {country_name}'s purse seine fleet from {report_year - 4} to {report_year}. In case of missing years in the table below, this indicates that no interactions with non-target species were reported during those years. \n\n"
  ))
  
  flextable(PS_tab |>
              rename(Species = sp_code)) |>  
    set_table_properties(layout = "autofit", width = 0.95, align = "center") |> 
    fontsize(size = 9, part = "all") |> 
    width(j = 1, width = 2) |> 
    colformat_num(big.mark = ",") |> 
    align(align = "center", part = "all") |> 
    labelizor(labels = species_labels_all, part = "all") |>
    padding(padding.left = 5, padding.right = 5, part = "all")
}
```

```{r}
#| label: tbl-tbl4l
#| eval: !expr "nrow(data_3605) > 0"
#| tbl-cap: !expr glue::glue("Annual estimated catches of non-target, associated and dependent species, including sharks, by the {country_name} longline fleet, in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605")
#| echo: false
#| results: 'asis'

# Ensure LL_tab has all expected columns even if empty
required_cols <- c("sp_code", as.character((params$year - 4):params$year))
if (nrow(LL_tab) > 0) {
  # Print header
  cat(glue::glue("## Annual catch estimates for non-target, associated and dependent species for the national longline fleet\n\n"))
  
  flextable(LL_tab |>
            rename(Species = sp_code)) |>    
    set_table_properties(layout = "autofit", width = 0.95, align = "center") |> 
    fontsize(size = 9, part = "all") |> 
    width(j = 1, width = 2) |> 
    colformat_num(big.mark = ",") |> 
    align(align = "center", part = "all") |> 
    labelizor(labels = species_labels_all, part = "all") |>
    padding(padding.left = 5, padding.right = 5, part = "all")
}
```

<!-- # Coastal State Reporting -->

```{r loadcoastal}
#| output: asis
#| echo: false

# parse the comma-separated string into a clean vector
sections <- strsplit(params$additional_sections, ",")[[1]] |>
  trimws() |>
  (\(x) x[x != ""])()   # drop empty strings

if ("artisanal" %in% sections) {
  # Check whether artisanal data exists BEFORE knitting the child
  ika_file        <- paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")
  has_ika         <- file.exists(ika_file)
  has_tails <- length(list.files(this_yr_folder, pattern = "3615")) > 0

  if (has_ika || has_tails) {
    cat("\\newpage\n\n# Coastal State Reporting\n\n")
    cat(knitr::knit_child("template_artisanal_vu.qmd", quiet = TRUE))
    cat("\n\n\\clearpage\n\n")
  } else {
    message("No artisanal data for ", country_cd, " - skipping coastal section")
  }
}
```

# Socio-economic factors

Since 2013 the number of Foreign and locally based foreign license has dropped as
most vessels were moving to the Solomon Islands EEZ and towards the eastern
pacific where fishing was reported to be very good. Vessels that were offloading
their catch in the Vanuatu EEZ through transshipment were also reporting low
catches towards the end of 2013 to early 2014 thus Transshipment in port was not
as regular as before and towards the end of 2014 Transshipment in port has ceased.
Locally based operations however ceased after 2014 and was only ignited back in
2019 under the Sino-Van company which licensed 6 vessels to fish and offload
their catches in Port Vila both for local sales and export to international markets.
However, in 2023, the vessels offload of the 6 locally based foreign fishing vessels
shifted to Fiji as a result of the pontoon being damaged during the twin tropical
cyclone that hit Vanuatu in March 2023 and in 2024 the vessels went to slipway.

For local artisanal fisherman, fishing in FADs have recently become a priority with
the reduction increase in fuel prices due to the Covid pandemic as well as the war
between Ukraine and Russia which affected the global distribution of fuel and in
turn affected fuel prices locally. This has made it an important asset for more
Artisanal fisherman who target FAD’s both to catch skipjack for Baitfish and to
catch yellowfin to sell in the local markets.

The TUFMAN2 database has been fully utilized since July 2016 after the shift from
TUFMAN 1 which it was only used for licensing up until 2019 when the licensing
component of the Tufman system was pushed into the FFA-RIMF system. The
TAILS system has also been trailed out in Vanuatu in 2016 and from 2021 there
has been a successful rollout to all 6 provinces in Vanuatu and present in 26 islands
covering 50 landing sites and 32 Area councils. The extension to all other
remaining 39 Area councils was met in 2022. Fishers including small gleaning
activities, skiffs and motorized canoes are being registered within the TAILS
system with the objective of enhancing the capacity to collect data for coastal, deep
bottom and pelagic fisheries. The RIMF FFA database has been utilized to cater for
the recording of Landing and Unloading data, transshipment data as well as MCS
boarding and inspection information. There is plan to include the entry of catch and
effort data from the non-Tuna RFMOs in the coming year in the RIMF system.

# Onshore Developments

The processing plant (Tuna Fishing Vanuatu Limited) in Port Vila harbor seized
operations in February 2014 due to movement of the fleet to the Solomon Islands.
The Chinese fishing Base known as the Sino-Van company is a joint venture to the
Vanuatu government owned 6 locally based Chinese fishing vessels that were
licensed in late 2019 and currently fishing the Vanuatu EEZ and offload their catch
in Port Vila and export their catch to China, USA, Vietnam, Australia and New
Zealand. Vanuatu has rolled out E reporting on 95% of its Vanuatu flag vessels
from 2020 to 2021 and in 2022 the coverage increases up to 100%. Through the
Vessel Day Scheme, Vanuatu plans to implement the ER system on all licensed
vessels operating within its EEZ in the coming year. Vanuatu has plans to develop
its onshore activities to further enhance and utilize the current fish landing
activities taking place. In 2024, the Department of Fisheries was moved from the
Ministry of Agriculture, Livestock, Fisheries, Forestry and Biosecurity to the newly
established Ministry of Fisheries, Oceans and Maritime Affairs.

# Future Prospects of Fishery

Vanuatu has maintained its position to limit the number of licenses to 70 Foreign
License and 40 Locally Based Foreign license however the license fee has been
increased by 50% of the current fee. Vanuatu has plans to move to vessel day
scheme (VDS). The Vanuatu Fisheries Department in November last year 2023 has
soft launch the longline vessel day scheme (VDS) along with stake holder
consultations regarding the transition to the new system. In 2024, the staff were
trained on how to implement the Vessel Day Scheme (VDS) activities.

# Research and Statistics

## Estimated data coverage

Coverage of logsheets from foreign fleets fishing in the Vanuatu EEZ extends back
as far as the 1970s and has been low and variable among years. There has also been
significant missing data throughout the years thus the difficulty in estimating
coverage rates for some years. However, with the inclusion of VMS data into the
Tufman system that is being used to report catches, there has been tremendous
improvement in estimating logsheet coverage and estimating catch production
respectively.

Logsheet data for Vanuatu flag vessels is currently being submitted through
Electronic reporting from a trial that has been in place since 2019 and which has
extended to 100% of the Vanuatu flag vessels in 2024. ER has proven effective by
reducing the amount of effort on data entry to data checking and quality control.
Data coverage is however, still low on the collection of size frequency data as most
of Vanuatu’s flagged vessels as well as the licensed vessels that fish in the Vanuatu
EEZ have been unloading their catch in foreign ports in the region such as Fiji,
Western Samoa and Solomon Islands and internationally such as Taiwan and
Panama.

Vanuatu is looking into strict measures in terms of estimating catch and effort data,
since most of our licensed vessels are currently offloading all or part of their
catches overseas, either to the factory or on the carrier vessel in port. One of the
major steps for Vanuatu in 2016 was to move to the TUFMAN2 database which
allows for the sharing of logsheet data between countries to which licensed vessels
operated. This sharing has allowed access to view Logsheets from vessels who are
license to operate in the Vanuatu waters with other members sharing the same
interests. And this is currently being fulfilled since 2017 as logsheet data is being
shared by countries which has significantly reduced the amount of work load on
entering data and giving more time to fixing and validating the data.
Most of the current presented data were obtained from the OFP/SPC database, and
were originally collected and supplied by Vanuatu and other member countries.

## Status of Tuna Fisheries Data Collection Systems

### Logsheet Data collection and Verification

There have been vast improvements with the collection of logsheet data since it has
become one of the special licensing conditions; which has forced vessel
owners to keep up with thesubmission of logsheet data. For the licensed vessels
fishing inside VU EEZ the logheet coverage based on VMS Data was between
80%-100% from 2014 to 2023 and in 2024 the coverage is 75.89% and will
improve once all logsheets are submitted and all outstanding data a r e being
entered. Whereas for the Vanuatu flag vessels the coverage for Longline for
2014-2020 is 100% except for 2018 and 2019 the coverage is 98.37% and 99.33%
respectively which is obtained through reconciliation with VMS data sought from
the Vanuatu VMS system. For Purse seine vessels, logsheet coverage from both
2014 and 2024 has been at 100. All Vanuatu fleets in 2024 (longline and purse
seine) are 100% electronic reporting fulfilling a 100% coverage of all fleets
reporting using ER.

Assistance from trainings held by SPC on data verification has assisted in allowing
data verification to be done nationally by officers as a means of improving data
quality and compliance.

### Observer and Port Sampling Programme

The Vanuatu Observer program established in 2008 and now has 54 regional certified PIRFO
observers who observe on Purse Seiners, Long Liners and Fish Carriers that are operating in the
WCPFC area. Since its establishment in 2008, Fisheries Observers have been involved in Longliners
and Purse seiner vessel operations and later covering fish carrier vessels in the effort to collect more
information on carrier vessels at sea.
In 2017, Vanuatu National Observer Programme has managed to put in place its first Emergency
Action Plan (EAP) and Standard operational Procedures. This is a great improvement to the
programe. The programme has also purchased safety gears such as 2-way In Reach Communicating
Device, Personal Locator Device (PLB) and Life vests and has also received few of these equipment
from the Forum Fisheries Agency (FFA. In 2023 and 2024 recruitement of observers took place
within the provinces of Vanuatu which now provides a total of 64 active and trained observers in the
program.

### Unloading and Transshipment

Unloading and Transshipment in Vanuatu port has been by way of locally based foreign vessels
fishing in Vanuatu EEZ. Since 2009 there has been 100% port sampling for all unloading and
transshipment activity in Vanuatu EEZ. Transshipment has been constantly carried out within the
harbor mostly targeting albacore for canning (e.g. Fiji and Solomon) or other species such as sharks
(mainly fins), Marlins, wahoo, Sword fish and other relevant by-catch including low grade yellow fin
and big eye. Transshipment is 100% sampled in measurement and estimated capacity weight of each
fish well; all fish for transshipment are stored frozen in blast freezers. Transshipment often occurs
once a month until 2014 when there were only 4 transshipments in the Vanuatu port. As of then,
transshipment activities has since ceased. For Vanuatu-flagged fishing vessels, transshipment by
longline vessels takes place on the high seas, while 100% of purse seine transshipment occurs at port.
In Vanuatu, unloading activities slowly occurred in 2017 and 2018 where 2 unloading took place as a
way of showcasing and promoting Vanuatu’s capacity of having its vessels resume unloading
activities in Vanuatu. In 2019, vessels resumed landing activities under the joint agreement between
the Vanuatu government and the SinoVan fishing company which experienced a total of 9 unloading
that took place by 6 locally based foreign vessels. Landed catch were exported to China, USA, New
Zealand and Japan with the sales of frozen catch also provided to the local markets. In 2024, all six
locally based vessels went to slipway, and Vanuatu licensed two additional locally based vessels
towards the end of the year. It is expected that more of the company vessels will engage in offloading
their catch in the Port Vila port in the following year developing the local market of Port Vila and
Santo. Vanuatu-flagged longline vessels offload or unload their catch at designated ports such as
Kaohsiung, Busan, American Samoa (when under charter), and Suva, while purse seine vessels
mainly offload in Papua New Guinea. The landed catch is exported to markets in the USA, China, and
Thailand, with some also sold locally.

### Disposal of Catch

Fresh Tuna previously landed in Vanuatu by Locally Based Foreign vessels were exported by air to
Japan as well as USA and New Zealand, while other frozen fish including Albacore are exported to
China and the remaining catch sold in the local markets in Port Vila. The Foreign fleets that have been
licensed to fish in Vanuatu EEZ unload 100% of their catch (both their fresh and frozen) either, in
Pago Pago or Fiji in 2019, while in 2024 the 2 licensed locally based vessels engaged in fishing 100%
offload their catch in Vanuatu port.

## Research Activities

There was no major research activities carried out in 2024.

<!-- # Addendum -->

```{r loadaddendum}
#| output: asis
#| echo: false

# Set values the child can read
applicable_measures <- "all"

if ("addendum" %in% sections) {
  cat("\\newpage\n\n# Addendum\n\n")
  cat(knitr::knit_child("template_addendum.qmd", quiet = TRUE))
}
```
