---
date: today
date-format: "DD MMMM YYYY"
format: 
  pdf:
    mainfont: "Times New Roman"
    sansfont: "Times New Roman"
    tbl-cap-location: bottom
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"    
    number-sections: true
    highlight-style: github
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
knitr:
  opts_chunk:
    out-width: "100%"
    dev: "cairo_pdf"
page-layout: full
edit: source
params:
  country_code: 'NR'          
  year: 2025                  
  data_provided: 'Yes'        
  aceman: TRUE
  additional_sections: ""
execute: 
  eval: true
  echo: false
  warning: false
  message: false
---

```{r libs}
library(tidyverse)
library(flextable)
library(scales) 
library(maps)
library(sf)
library(stringr)

source("utils.R")

theme_set(theme_bw(base_family = "Times New Roman"))
flextable::set_flextable_defaults(font.family = "Times New Roman")

# key params
country_cd <- tolower(params$country_code)
report_year = params$year
this_yr_folder = paste0("./data/report_", as.character(report_year), "_", tolower(country_cd), "/") |>
  tolower()

'%nin%' = function(x,y) !x %in% y            

fmt <- function(x, digits = 0) {
  format(round(x, digits), big.mark = ",", nsmall = digits, scientific = FALSE, trim = TRUE)
}

```

\newpage

```{r header}
row1 <- glue::glue("Scientific data was provided to the Commission in accordance with the decision relating to the provision of scientific data to the Commission by 30 April {params$year+1}: [Yes/No]")
row2 <- "If no, please indicate the reason(s) and intended actions:"
row3 = row4 = row5 = ''

tab <- data.frame(lcol = c(row1, row2, row3, row4, row5), 
                  rcol = c(params$data_provided, rep("", 4)) )

ft <- flextable(tab)
# No header row
ft <- delete_rows(ft, i = 1, part = "header")
# Merge cells - there are a ton of merge_* functions - using this one for simplicity
ft <- merge_at(ft, i = 2:5, j = 1:2)
ft <- valign(ft, i = 2:5, valign = 'top') 
# Relative column width
ft <- width(ft, width = c(1, 0.1)) # Set relative width of columns
ft <- width(ft, width = dim(ft)$widths * 6.6 / (flextable_dim(ft)$widths)) 
ft <- border_outer(ft)
ft <- border_inner(ft)

ft
```

```{r}
#| label: prelims

# Define years to use for tables and figures
yrs_long <- (params$year-4):params$year
yr_range <- paste(min(yrs_long),max(yrs_long), sep="-")

# Map for plotting  
newmap = fortify(maps::map(wrap=c(0,360), plot=FALSE, fill=TRUE))  

# Create a vector of the key species to plot for each gear
ps_species <- c("skj", "yft", "bet", "alb", "pbf")
ll_species <- c("skj", "yft", "bet", "alb", "pbf", "blm", "bum", "mls", "swo")
pl_species <- c("skj", "yft", "bet", "alb", "pbf")

# Define species labels for each gear type
ps_species_labels <- c(
  alb = "Albacore",
  bet = "Bigeye tuna",
  skj = "Skipjack tuna",
  yft = "Yellowfin tuna",
  pbf = "Pacific bluefin tuna"
)

ll_species_labels <- c(
  alb = "Albacore",
  bet = "Bigeye tuna",
  skj = "Skipjack tuna",
  yft = "Yellowfin tuna",
  pbf = "Pacific bluefin tuna",
  swo = "Swordfish",
  blm = "Black marlin",
  bum = "Blue marlin",
  mls = "Striped marlin"
)

pl_species_labels <- c(
  alb = "Albacore",
  bet = "Bigeye tuna",
  skj = "Skipjack tuna",
  yft = "Yellowfin tuna",
  pbf = "Pacific bluefin tuna"

)

species_labels_all <-  c(
  alb = "Albacore",
  bet = "Bigeye tuna",
  skj = "Skipjack tuna",
  yft = "Yellowfin tuna",
  pbf = "Pacific bluefin tuna",
  swo = "Swordfish",
  blm = "Black marlin",
  bum = "Blue marlin",
  mls = "Striped marlin",
  ssp = "Shortbill spearfish",
  sfa = "Sailfish",
  bsh = "Blue shark",
  fal = "Silky shark",
  ocs = "Oceanic whitetip shark",
  
  mak = "Mako sharks (mak)",
  sma = "Mako sharks (sma)",
  lma = "Mako sharks (lma)",
  
  thr = "Thresher sharks (thr)",
  alv = "Thresher sharks (alv)",
  pth = "Thresher sharks (pth)",
  bth = "Thresher sharks (bth)",
  
  por = "Porbeagle shark",
  
  spn = "Hammerhead sharks (spn)",
  spk = "Hammerhead sharks (spk)",
  spl = "Hammerhead sharks (spl)",
  spz = "Hammerhead sharks (spz)",
  spq = "Hammerhead sharks (spq)",
  eub = "Hammerhead sharks (eub)",

  rhn = "Whale shark"
  
  )

country_names <- c(
  fj  = "Fiji",
  fm = "Federated States of Micronesia",
  gu = "Guam",
  ki = "Kiribati",
  mh = "Marshall Islands",
  mp = "Northern Mariana Is.",
  nc = "New Caledonia",
  nr = "Naoero",
  nu = "Niue",
  pf = "French Polynesia",
  pg = "Papua New Guinea",
  pn = "Pitcairn",
  pw = "Palau",
  sb = "Solomon Islands",
  tk = "Tokelau",
  to = "Tonga",
  tv = "Tuvalu",
  vu = "Vanuatu",
  wf = "Wallis and Futuna Islands",
  ck = "Cook Islands",
  ws = "Samoa"
)

# get country name
country_name <- country_names[country_cd] 

  
# Give species colours to maintain colour schemes throughout
# --This makes sure species colors are consistent in all plots

sp_colours <- setNames(object = c("seagreen3","red3","black","cyan1","steelblue","orange",
                                  "pink","wheat","purple3","grey40","dodgerblue4","purple4",
                                  "#66CCFF","salmon4","magenta3","gold"),
                         nm = c("Albacore","Bigeye tuna","Black marlin","Blue marlin","Blue shark",
                                "Hammerhead sharks nei","Mako sharks","Oceanic whitetip shark",
                                "Pacific bluefin tuna","Silky shark","Skipjack tuna",
                                "Striped marlin","Swordfish","Thresher sharks nei",
                                "Whale shark","Yellowfin tuna"))
  
```

\newpage

# Abstract 

This report provides a comprehensive overview of Papua New Guinea’s (PNG) tuna fisheries for 2025, detailing trends in catch, effort, and management within its Exclusive Economic Zone (EEZ) and the wider Western and Central Pacific Fisheries Commission (WCPFC) Convention Area. The purse-seine fishery remains the dominant sector, contributing over 99% of total production. In 2025, the national purse-seine fleet recorded a total catch of 329,075 metric tonnes (mt), a 4.09% increase from 2024. This growth was driven by favourable La Niña conditions, with skipjack tuna (215,043 mt) and yellowfin tuna (112,652 mt) as the primary species.

In contrast, the domestic longline fishery experienced a significant contraction, with no PNG-flagged vessels operational in 2025 following the reflagging of the chartered fleet in 2024. However, foreign bilateral longline operations continued under regulated access agreements, maintaining an average annual catch of approximately 4,078 mt over the last five years. The report also highlights the closure of the shark longline and handline fisheries, reflecting a strategic shift toward sustainable tuna management and the prohibition of certain shark species landings.

Fisheries management in PNG is governed by the National Tuna Fishery Management and Development Plan (NTFMDP), revised in 2025 to align with regional Conservation and Management Measures (CMMs) and the Vessel Day Scheme (VDS). Beyond biological sustainability, the PNG government is prioritizing socio-economic optimization through the East New Britain Initiative (ENBi), focusing on downstream processing, employment, and increased export earnings. While tuna exports peaked in 2024 at 280,000 mt ($600 million), 2025 figures indicate a stabilization of the market, emphasizing the need for continued monitoring of both production volumes and global market values.

# Background

Tuna resources within Papua New Guinea’s (PNG) Exclusive Economic Zone (EEZ) are exploited predominantly through two industrial fishing methods: purse seine and longline. The purse-seine fishery is the dominant harvesting sector, consistently contributing approximately 99% of total national tuna production by volume.

Purse-seine operations commenced in PNG waters in the early 1980s and have since expanded substantially in both spatial extent and fishing effort. This expansion reflects increased participation by domestic fleets and foreign bilateral arrangements, coupled with enhanced fishing efficiency associated with Fish Aggregating Device (FAD) use and technological advancements. Peak aggregate purse-seine catches (inclusive of domestic and FB fleets) were recorded in 2010, 2022, and 2024, each exceeding 600,000 metric tonnes, indicating periods of elevated stock availability and/or fishing effort concentration within PNG waters.

In contrast, the longline fishery has a longer historical presence, initially dominated by distant water fishing nation (DWFN) fleets operating under access agreements. A significant policy shift in 1996 resulted in the closure of the longline fishery to foreign participation, restricting operations exclusively to nationally flagged or domestically based entities. This policy was intended to promote domestic industry development and value retention within PNG.

In 2015, the longline sector was partially liberalized under a controlled re-entry framework through the implementation of a trial albacore fishery. This initiative aimed to evaluate catch-per-unit-effort (CPUE) performance and economic viability of albacore targeting within PNG waters. Under this arrangement, five foreign-flagged ultra-low temperature (ULT) longline vessels were authorized to operate under domestic charter agreements, marking a cautious reintroduction of foreign participation into the longline fishery.
Subsequent expansion occurred in 2016, with 20 foreign longline vessels operating under bilateral access agreements. This number increased progressively, reaching 61 licensed foreign longline vessels by 2023. In parallel, PNG’s domestically based fleet comprised 10 chartered longline vessels actively operating within the EEZ during the same period. However, a restructuring of vessel registration status in 2024 resulted in the reflagging of these chartered vessels to foreign registries, leaving only a single PNG-flagged longline vessel operational in that year. Subsequently, in 2025 saw no national tuna longline flagged vessels fishing in PNG EEZ. 

PNG’s tuna fisheries governance framework reflects a deliberate policy balance between domestic industry development and the strategic allocation of access to DWFNs. This is operationalized through a licensing regime that incentivizes onshore value addition, including mandatory catch offloading for in-country processing. Incentive mechanisms include preferential licensing fees, eligibility for reflagging under PNG registry, authorized access to Archipelagic Waters (AW), and participation under the Federated States of Micronesia Arrangement (FSMA), enabling qualifying vessels to operate within other Parties to the Nauru Agreement (PNA) jurisdictions.

The fishery is managed under the National Tuna Fishery Management and Development Plan (NTFMDP), which provides the overarching legal and policy framework governing all tuna fishing activities within PNG’s jurisdiction. The NTFMDP incorporates a comprehensive suite of management measures, including effort controls (e.g., Vessel Days Scheme alignment), catch monitoring, gear restrictions, FAD management regulations, and monitoring, control, and surveillance (MCS) mechanisms aimed at mitigating Illegal, Unreported, and Unregulated (IUU) fishing.

Management of the purse-seine fishery is further reinforced through integration with regional and sub-regional frameworks, including Conservation and Management Measures (CMMs) under the Western and Central Pacific Fisheries Commission (WCPFC), PNA Implementing Arrangements, and the Forum Fisheries Agency (FFA) Harmonized Minimum Terms and Conditions (HMTCs). These instruments are operationalized domestically through the NTFMDP, ensuring coherence between national policy objectives and PNG’s regional and international fisheries obligations. The NTFMDP was most recently revised in 2025 to reflect evolving management priorities, regional commitments, and industry development objectives.

# Flag State Reporting

This section reports activities by the national fleet inside the waters of the Western and Central Pacific Fisheries Commission Convention Area (WCPFC CA) including PNG’s Exclusive Economic Zone (EEZ). The national fleet comprises of longline and purse seine vessels under PNG flag and those foreign flag vessels under domestic charter arrangements.

## National Fleet: Tuna Longline

PNG did not have any domestic tuna longline vessels fishing in its EEZ nor the wider the WCPFC CA in 2025. This reflected a sharp decline in domestic vessel numbers, from 10 vessels in 2023 to one vessel in 2024, and ultimately no longline fishing operations took place for this category in 2025.

## National Fleet: Purse Seine

Catches from the Papua New Guinea (PNG) national purse-seine fleet are overwhelmingly dominated by skipjack tuna (Katsuwonus pelamis), which consistently constitutes the largest proportion of total landings equating to approximately 65% of total tuna catches caught. This is followed by yellowfin (Thunnus albacares) and, to a lesser extent, bigeye tuna (Thunnus obesus). Although skipjack remains the principal target species, both yellowfin and bigeye represent important secondary species, contributing substantially to the overall economic value and species composition of the fishery.

In 2025, total catches from the PNG domestic purse-seine fleet increased by 4.09% relative to the 2024 fishing season, reaching a cumulative production of 329,075 metric tonnes (mt). This incremental growth reflects a continuation of the positive catch trajectory observed in recent years and is largely attributable to favourable oceanographic conditions associated with a La Niña phase in the Western Pacific Ocean. Such conditions are known to enhance skipjack availability and aggregation within PNG waters.

Utilizing an overall logsheet effort of 3,775 fish days, the 2025 logsheet data indicates that skipjack accounted for 215,043 mt, followed by yellowfin at 112,652 mt, and bigeye at 1,298 mt (Table 1). These figures reaffirm the strong dominance of skipjack within the purse-seine fishery, while also highlighting the sustained contribution of yellowfin and the relatively minor, yet commercially relevant, presence of bigeye tuna.

```{r ps-prep}

# Load all raw data - from t2
data_2953 <- safe_read_and_clean(this_yr_folder, country_cd, r_code = 2953)

# this is a quick fix, it is misleading to have a df called data_3605 when 
# data is being sourced from report 3619...
if(params$aceman){
  data_3605 <- safe_read_and_clean(this_yr_folder, country_cd, r_code = 3619)
  
  if("catch" %in% colnames(data_3605)){
    data_3605$unadjcatch <- data_3605$catch
  } 
  
}else{
  data_3605 <- safe_read_and_clean(this_yr_folder, country_cd, r_code = 3605)
}

# For 3608 with special post-processing
data_3608 <- safe_read_and_clean(
  this_yr_folder, 
  country_cd, 
  r_code = 3608,
  post_process = function(data) {
    data |>
      mutate(
        lon = ifelse(lon < 0, lon + 360, lon),
        lat = ifelse(gear_code == "s", lat + 0.5, lat + 2.5),
        lon = ifelse(gear_code == "s", lon + 0.5, lon + 2.5)
      )
  }
)

if (nrow(data_3605) == 0){
  data_3605_ps <- data.frame()
  data_3605_ll <- data.frame()
  data_3605_pl <- data.frame()
  }else{
    
 # only data until current report year
  data_3605 <- data_3605 |>
    mutate(year = as.numeric(year)) |>
    filter(year <= report_year)
  
  # separate in PS, LL and PL
  data_3605_ps <- data_3605 |>
    filter(gear_code == "s") |>
    filter(sp_code %in% ps_species) |>
    select(sp_code, year, catch)
  
  data_3605_ll <- data_3605 |>
    filter(gear_code == "l") |>
    filter(sp_code %in% ll_species) |>
    select(sp_code, year, catch)
  
  data_3605_pl <- data_3605 |>
    filter(gear_code == "pl") |>
    filter(sp_code %in% ll_species) |>
    select(sp_code, year, catch)
}

# shape for tables
shape_catch_tbl <- function(data) {
  data |>
    pivot_wider(names_from = year, values_from = catch, values_fill = 0) |>
    mutate(across(where(is.numeric), ~replace_na(., 0))) |>  # only needed if catch itself can be NA
    arrange(sp_code) |>
    rename(Species = sp_code) |>
    janitor::adorn_totals("row") |>
    select(Species, sort(tidyselect::peek_vars()))
}

if (nrow(data_3605_ps) > 0) data_3605_ps_tbl <- shape_catch_tbl(data_3605_ps)
if (nrow(data_3605_ll) > 0) data_3605_ll_tbl <- shape_catch_tbl(data_3605_ll)
if (nrow(data_3605_pl) > 0) data_3605_pl_tbl <- shape_catch_tbl(data_3605_pl)

# effort - number of hooks per year for the national fleet
data_3597 <- safe_read_and_clean(this_yr_folder, country_cd, r_code = 3597) 

```

```{r funcs-render}
#| echo: false

# Function to render gear-specific tables in section 3 with header and table
render_gear_tbl_section3 <- function(data, data_tbl, gear_name, gear_code, species_labels, intro_text = NULL)
  {
  
  # Check if data exists
  if (nrow(data) == 0) {
    return(invisible(NULL))
  }
  
  # Print header
  cat(glue::glue("## Annual catch estimates for the national {gear_name} fleet\n\n"))
  
  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }
  
  # Create and return the flextable
  ft <- flextable(data_tbl) |>
    labelizor(labels = species_labels, part = "all") |>
    set_table_properties(width = 0.7, align = "center") |> 
    bold(j = ncol(data_tbl), part = "body") |>
    width(j = 1, width = 2)
  return(ft)
}

# Function to render gear-specific figures in section 3 with header and plot
render_gear_figs_section3 <- function(data, gear_name, gear_code, species_list,
                                      species_labels, intro_text = NULL,
                                      effort_df = data_3597) {

  # Check if data exists
  if (nrow(data) == 0) {
    return(invisible(NULL))
  }

  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }

  # ── Prepare data for plotting ──────────────────────────────────────────────
  data_fig <- data |>
    group_by(sp_code, year) |>
    summarise(catch = sum(catch, na.rm = TRUE), .groups = "drop")

  ref_df <- expand_grid(year = as.numeric(yrs_long), sp_code = species_list)

  plot_data <- left_join(ref_df, data_fig, by = c("year", "sp_code")) |>
    mutate(
      catch   = replace_na(catch, 0),
      Species = factor(recode(sp_code, !!!species_labels),
                       levels = unname(species_labels[species_list]))
    )

  # ── Base plot (shared by all gears) ────────────────────────────────────────
  p <- ggplot(plot_data, aes(x = year, y = catch, fill = Species)) +
    geom_col(position = "stack") +
    scale_fill_manual(values = sp_colours) +
    scale_x_continuous(breaks = seq(min(plot_data$year), max(plot_data$year), 1)) +
    labs(x = NULL, y = "Catch (mt)") +
    theme_bw() +
    theme(
      axis.title   = element_text(size = 16),
      axis.text    = element_text(size = 16),
      axis.text.x  = element_text(angle = 45, hjust = 1),
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text  = element_text(size = 13)
    )

  # ── Longline: overlay effort (hooks) on a secondary axis ──────────────────
  if (gear_code == "l") {
    
    # maybe at some point this will fail in case effort_df doesnt exist or nrow = 0
     effort_annual <- effort_df |>
      filter(flag_id == country_cd) |>
      filter(year %in% plot_data$year) |>          # align with plotted years
      group_by(year) |>
      summarise(effort = sum(hooks_total_logsheet, na.rm = TRUE), .groups = "drop")

    max_catch  <- max(tapply(plot_data$catch, plot_data$year, sum))
    max_effort <- max(effort_annual$effort, 0)

    if (nrow(effort_annual) > 0 && max_effort > 0 && max_catch > 0) {

      scale_factor <- max_catch / max_effort

      p <- p +
        geom_line(data = effort_annual,
                  aes(x = year, y = effort * scale_factor),
                  inherit.aes = FALSE, colour = "darkorange",
                  linewidth = 1.2, linetype = "dashed") +
        geom_point(data = effort_annual,
                   aes(x = year, y = effort * scale_factor),
                   inherit.aes = FALSE, colour = "darkorange", size = 3) +
        scale_y_continuous(
          labels = comma,
          sec.axis = sec_axis(~ . / scale_factor,
                              name = "Effort (hooks)", labels = comma)
        ) +
        theme(
          axis.title.y.right = element_text(colour = "darkorange"),
          axis.text.y.right  = element_text(colour = "darkorange")
        )
    } else {
      message("No usable effort data for gear '", gear_code,
              "' — rendering catch only.")
      p <- p + scale_y_continuous(labels = comma)
    }

  } else {
    p <- p + scale_y_continuous(labels = comma)
  }

  print(p)
  invisible(NULL)
}

# Function to render vessel table for a specific gear
render_vessel_table <- function(ves_data, gear_name, size_categories_fn, intro_text = NULL) {
  
  # Filter for specific gear
  gear_data <- ves_data |> filter(gear_code == gear_name)
  
  # Check if data exists
  if (nrow(gear_data) == 0) {
    return(invisible(NULL))
  }
  
  # Apply the size category function to summarize data
  ves_summary <- gear_data |>
    group_by(year, gear) |> 
    size_categories_fn()
  
  # Check if there's any actual vessel data (not all zeros)
  if (sum(ves_summary[,-c(1:2)]) == 0) {
    return(invisible(NULL))
  }
  
  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }
  
  # Create and return flextable
  ves_summary |> 
    pivot_longer(-c(year, gear), names_to = "Size Category (GRT)", values_to = "Vessels") |>
    pivot_wider(names_from = year, values_from = Vessels, values_fill = 0) |>
    group_by(gear) |> 
    mutate(gear = ifelse(row_number() == 1, gear, "")) |>
    flextable() |>  
    set_table_properties(width = 0.7, align = "center") |> 
    width(j = 1:2, width = 1.2)
}

# Function to render spatial catch distribution map for a specific gear
render_spatial_map <- function(data, gear_code, gear_name, target_species, year_range, intro_text = NULL) {
  
  gear_code <- tolower(gear_code)
  
  # Filter and prepare data
  spatial_data <- data |>  
    filter(
      between(year, year_range[1], year_range[2]), 
      gear_code == !!gear_code, 
      sp_code %in% target_species
    ) |> 
    group_by(lat, lon, year, species) |> 
    summarise(catch = sum(catch), .groups = "drop") |> 
    filter(catch > 0)
  
  # Check if data exists
  if (nrow(spatial_data) == 0) {
    return(invisible(NULL))
  }
  
  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }
  
  # Create the map
  p <- ggplot() +
    geom_map(data = newmap, map = newmap, aes(map_id = region), fill = 'bisque1', col = 'gray') +
    geom_tile(data = spatial_data, aes(lon, lat, fill = catch)) +
    scale_fill_distiller(palette = "Spectral") +
    scale_x_continuous(limits = c(120, 230), expand = c(0, 0), breaks = seq(140, 220, 40)) +
    scale_y_continuous(limits = c(-35, 25), expand = c(0, 0)) +
    facet_grid(year ~ species) +
    xlab("Longitude") + ylab("Latitude") + labs(fill = "Catch (mt)") +
    theme_bw() +
    theme(
      axis.text = element_text(size = 18),
      legend.text = element_text(size = 18),
      legend.title = element_text(size = 18),
      strip.background = element_rect(fill = 'white'),
      strip.text = element_text(size = 18),
      axis.title = element_text(size = 20), 
      title = element_text(size = 20)
    ) +
    coord_sf() +
    ggtitle(glue::glue('Spatial patterns in catch for the {gear_name} fishery'))
  
  print(p)
  return(invisible(NULL))
}

# Function to render SSI (Species of Special Interest) table for a specific gear
render_ssi_table <- function(data, gear_code, gear_name, intro_text = NULL) {
  
  # Filter data for specific gear
  tab_data <- data |>  
    filter(gear == gear_code) |> 
    select(-gear) |>
    arrange(year)
  
  # Check if data exists
  if (nrow(tab_data) == 0) {
    return(invisible(NULL))
  }
  
   # Print header
  cat(glue::glue("### Observed catches of SSIs by the {gear_name} fleet\n\n"))
  
  # Print intro text if provided (supports markdown)
  if (!is.null(intro_text)) {
    cat(glue::glue(intro_text), "\n\n")
  }

  # Create and return flextable
  flextable(tab_data) |>
    set_table_properties(layout = "autofit", width = 0.95, align = "center") |> 
    fontsize(size = 9, part = "all") |> 
    colformat_num(j = "year", big.mark = "") |>  
    width(j = c("category", "species"), width = 1.5) |> 
    align(align = "center", part = "all") |> 
    labelizor(
      labels = c(
        year = "Year",
        species = "Species",
        category = "Category",
        number = "Total",
        alive = "Alive",  # Fixed capitalization
        dead = "Dead"
      ), 
      part = "all"
    ) |>
    padding(padding.left = 5, padding.right = 5, part = "all")
}

```

```{r}
#| echo: false

yearly_total <- data_3605 |>
  group_by(year) |>
  summarise(sum = sum(catch), .groups = "drop")

total_5yr <- sum(data_3605$catch)
avg_total <- mean(yearly_total$sum, na.rm = TRUE)

avg_by_sp <- data_3605 |>
  group_by(sp_code, year) |>
  summarise(sum = sum(catch), .groups = "drop") |>
  group_by(sp_code) |>
  summarise(m = mean(sum, na.rm = TRUE), .groups = "drop")

avg_skj <- avg_by_sp$m[avg_by_sp$sp_code == "skj"]
avg_yft <- avg_by_sp$m[avg_by_sp$sp_code == "yft"]
avg_bet <- avg_by_sp$m[avg_by_sp$sp_code == "bet"]
```

`r country_name`'s Flag vessels for the past 5 years (@tbl-tbl1s) had caught a total of `r fmt(total_5yr)` mt of tuna at an average of `r fmt(avg_total)` mt per year. The average breakdown of catches by species are; SKJ at `r fmt(avg_skj)` mt, YFT at `r fmt(avg_yft)` mt, and BET at `r fmt(avg_bet)` mt per year.


```{r}
#| echo: false
#| results: 'asis'
#| label: tbl-tbl1s
#| tbl-cap: !expr if(nrow(data_3605_ps) > 0) glue::glue("Annual catch estimates for the {country_name} purse seine fleet, by primary species and year, in the WCPFC Convention Area. Source - Tufman 2, report 3605.") else ""

render_gear_tbl_section3(
  data = data_3605_ps,
  data_tbl = data_3605_ps_tbl,
  gear_name = "purse seine",
  gear_code = "s",
  species_labels = ps_species_labels,
  intro_text = ""
)
```

```{r}
#| echo: false
#| results: 'asis'
#| label: tbl-tbl1l
#| tbl-cap: !expr if(nrow(data_3605_ll) > 0) glue::glue("Annual catch estimates for the {country_name} longline fleet, by primary species and year, in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605.") else ""

render_gear_tbl_section3(
  data = data_3605_ll,
  data_tbl = data_3605_ll_tbl,
  gear_name = "longline",
  gear_code = "l",
  species_labels = ll_species_labels,
  intro_text = ""
)
```

```{r}
#| echo: false
#| results: 'asis'
#| label: tbl-tbl1pl
#| tbl-cap: !expr if(nrow(data_3605_pl) > 0) glue::glue("Annual catch estimates for the {country_name} pole line fleet, by primary species and year, in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605.") else ""

render_gear_tbl_section3(
  data = data_3605_pl,
  data_tbl = data_3605_pl_tbl,
  gear_name = "pole line",
  gear_code = "pl",
  species_labels = ll_species_labels,
  intro_text = ""
)
```

<!-- \newpage -->

```{r}
#| echo: false
#| results: 'asis'
#| fig-pos: 'H'
#| label: fig-fig1l
#| fig-cap: !expr if(nrow(data_3605_ll) > 0) glue::glue("Historical annual catch estimates (mt) and effort (no. of hooks) for the {country_name} longline fleet, by primary species and year, in the WCPFC Convention Area for {yr_range}. Source - Tufman 2, report 3605 and 3597 (hooks).") else ""
#| fig-width: 12
#| fig-height: 7
#| fig-align: "center"

render_gear_figs_section3(
  data = data_3605_ll,
  gear_name = "longline",
  gear_code = "l",
  species_list = ll_species,
  species_labels = ll_species_labels,
  intro_text = ""
)
```

<!-- \newpage -->

```{r}
#| echo: false
#| results: 'asis'
#| fig-pos: 'H'
#| label: fig-fig1s
#| fig-cap: !expr if(nrow(data_3605_ps) > 0) glue::glue("Historical annual catch estimates (mt) for the {country_name} purse seine fleet, by primary species and year, in the WCPFC Convention Area for {yr_range}. Source - Tufman 2, report 3605.") else ""
#| fig-width: 12
#| fig-height: 7
#| fig-align: "center"

# Modified from original: 

render_gear_figs_section3(
  data = data_3605_ps,
  gear_name = "purse seine",
  gear_code = "s",
  species_list = ps_species,
  species_labels = ps_species_labels,
  intro_text = ""
)
```

\newpage

```{r}
#| echo: false
#| results: 'asis'
#| fig-pos: 'H'
#| label: fig-fig1pl
#| fig-cap: !expr if(nrow(data_3605_pl) > 0) glue::glue("Historical annual catch estimates (mt) for the {country_name} pole line fleet, by primary species and year, in the WCPFC Convention Area for {yr_range}. Source - Tufman 2, report 3605.") else ""
#| fig-width: 12
#| fig-height: 7
#| fig-align: "center"

render_gear_figs_section3(
  data = data_3605_pl,
  gear_name = "pole line",
  gear_code = "pl",
  species_list = pl_species,
  species_labels = pl_species_labels,
  intro_text = ""
)
```

```{r}
#| echo: false
#| results: asis
#| eval: !expr nrow(data_3605) > 0

if (nrow(data_3605) > 0) {
  cat("\\newpage\n\n")
  cat("## Historical annual vessel numbers for the national fleet\n\n")
}
```

@fig-fig2 displays the total number of active vessels in `r country_name`’s national fleet from `r report_year - 4` to `r report_year`.

```{r}
#| label: fig-fig2
#| fig-pos: 'H'
#| eval: !expr nrow(data_3605) > 0
#| fig-cap: !expr if(nrow(data_3605) > 0) glue::glue("Historical annual vessel numbers for the {country_name} fleet, by gear, for the WCPFC Convention Area. Source - Tufman 2, report 3605.") else ""
#| fig-width: 12
#| fig-height: 7
#| fig-align: "center"

if (nrow(data_3605) > 0){
    # Plot of vessel numbers for only Longline, Purse seine and Pole Line, aggregated over size class
    vessel_colours <- c(
    "0 - 50"    = "#A1DAB4",
    "51 - 200"  = "#41B6C4",
    "201 - 500" = "#2C7FB8",
    "500+"      = "#253494"
  )
  
   ves_raw <- 
    data_3605 |>  
    filter(year %in% yrs_long, gear_code %in% c("s", "l", "pl")) |> 
    select(flag, year, gear_code, vessels_sml, vessels_med, vessels_lge, vessels_hge) |> 
    distinct() |>
    mutate(gear = case_when(
      gear_code == "s" ~ "Purse seine",
      gear_code == "l" ~ "Longline",
      gear_code == "pl" ~ "Pole line"
    ))
  
  vessel_plot <- ves_raw |>
    pivot_longer(
      cols = c(vessels_sml, vessels_med, vessels_lge, vessels_hge),
      names_to = "vessel_class",
      values_to = "count"
    ) |>
    mutate(vessel_class = recode(vessel_class,
                                 "vessels_sml" = "0 - 50",
                                 "vessels_med" = "51 - 200",
                                 "vessels_lge" = "201 - 500",
                                 "vessels_hge" = "500+"
    )) |>
    mutate(vessel_class = factor(vessel_class, levels = c("0 - 50", "51 - 200", "201 - 500", "500+")))
  
  p_vessels <- ggplot(vessel_plot, aes(x = year, y = count, fill = vessel_class)) +
    geom_col(position = "stack") +
    scale_fill_manual(values = vessel_colours) +
    scale_x_continuous(breaks = seq(min(vessel_plot$year), max(vessel_plot$year), 1)) +
    scale_y_continuous(breaks = scales::pretty_breaks()) +
    labs(x = NULL, y = "Vessels", fill = "Size category (GRT)") +
    theme_bw() +
    facet_wrap(~gear) +
    theme(
    axis.title   = element_text(size = 16),
    axis.text    = element_text(size = 16),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    strip.text   = element_text(size = 16),
    legend.position = "bottom",
    legend.title = element_text(size = 13),
    legend.text  = element_text(size = 13)
  )
  
  print(p_vessels)
} else {
  ves_raw <- tibble()
}
```


```{r}

#| echo: false

# Prepare vessel data once
if(nrow(data_3605) > 0){
  ves_raw <- 
    data_3605 |>  
    filter(year %in% yrs_long, gear_code %in% c("s", "l", "pl")) |> 
    select(flag, year, gear_code, vessels_sml, vessels_med, vessels_lge, vessels_hge) |> 
    distinct() |>
    mutate(gear = case_when(
      gear_code == "s" ~ "Purse seine",
      gear_code == "l" ~ "Longline",
      gear_code == "pl" ~ "Pole line"
    ))
} else {
  ves_raw <- tibble()
}

# Define size category functions for each gear type
ll_size_fn <- function(data) {
  data |>
  summarise(
    '0 - 50' = sum(vessels_sml),
    '51 - 200' = sum(vessels_med),
    '201 - 500' = sum(vessels_lge),
    '500+' = sum(vessels_hge),
    .groups = "drop"
  )
}

ps_size_fn <- function(data) {
  data |>
  summarise(
    '0 - 50' = sum(vessels_sml),
    '51 - 200' = sum(vessels_med),
    '201 - 500' = sum(vessels_lge),
    '500+' = sum(vessels_hge),
    .groups = "drop"
  )
}

pl_size_fn <- function(data) {
  data |>
  summarise(
    '0 - 50' = sum(vessels_sml),
    '51 - 150' = sum(vessels_med),
    '150+' = sum(vessels_lge + vessels_hge),
    .groups = "drop"
  )
}

```

```{r}
ll_caption <- if(nrow(ves_raw) > 0 && "gear_code" %in% names(ves_raw) && nrow(ves_raw |> filter(gear_code == "l")) > 0) {
  glue::glue("Number of {country_name} longline vessels, by size category, active in the WCPFC Convention Area, over {yr_range}. Source - Tufman 2, report 3605.")
} else NULL
```

```{r}
#| label: tbl-tbl2-longline
#| eval: !expr "!is.null(ll_caption)"
#| tbl-cap: !expr ll_caption
#| echo: false
#| results: 'asis'
render_vessel_table(
  ves_data = ves_raw,
  gear_name = "l",
  size_categories_fn = ll_size_fn,
  intro_text = ""
)
```

```{r}
ps_caption <- if(nrow(ves_raw) > 0 && "gear_code" %in% names(ves_raw) && nrow(ves_raw |> filter(gear_code %in% c("s", "ps"))) > 0) {
  glue::glue("Number of {country_name} purse seine vessels, by size category, active in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605.")
} else NULL
```

```{r}
#| label: tbl-tbl2-ps
#| eval: !expr "!is.null(ps_caption)"
#| tbl-cap: !expr ps_caption
#| echo: false
#| results: 'asis'

render_vessel_table(
  ves_data = ves_raw,
  gear_name = "s",
  size_categories = ps_size_fn,
  intro_text =  ""
)
```

```{r}
#| echo: false
#| results: asis
#| eval: !expr nrow(data_3608) > 0

if (nrow(data_3605) > 0) {
  cat("\\newpage\n\n")
  cat("## Distribution of catches of target species for different national fisheries\n\n")
}
```

In this section we present the spatial distribution of catch (in metric tonnes) by species, 
Albacore, Bigeye tuna, and Yellowfin tuna , from `r country_name`’s fleet across the period `r report_year - 4`-
`r report_year`, primarily concentrated within and just outside `r country_name`’s Exclusive Economic Zone (EEZ).

```{r}
#| label: fig-figmapps
#| fig-cap: !expr glue::glue("Map showing the distribution of target species catch (mt) for the {country_name} purse seine fishery. Source - Tufman 2, report 3608.")
#| fig-pos: "H"
#| echo: false
#| fig-width: 12
#| fig-height: 8

map_ps <- file.path(this_yr_folder, glue::glue("map_ps_{country_cd}_{params$year}.png"))
if (file.exists(map_ps)) 
    knitr::include_graphics(map_ps)
```

```{r}
#| label: fig-figmapl
#| fig-cap: !expr glue::glue("Map showing the distribution of target species catch (mt) for the {country_name} longline fishery. Source - Tufman 2, report 3608.")
#| fig-pos: "H"
#| echo: false
#| fig-width: 12
#| fig-height: 8

map_ll <- file.path(this_yr_folder, glue::glue("map_ll_{country_cd}_{params$year}.png"))
if (file.exists(map_ll)) knitr::include_graphics(map_ll)
```

```{r}
#| label: fig-figmappl
#| eval: !expr "nrow(data_3608) > 0"
#| fig-cap: !expr if(nrow(data_3608 |> filter(gear_code == "pl") > 0)) glue::glue("Map showing the distribution of target species catch (mt) for the {country_name} pole line fishery. Source - Tufman 2, report 3608.") else ""
#| fig-width: 12
#| fig-height: 8
#| fig-align: "center"
#| fig-pos: "H"
#| echo: false
#| message: false
#| warning: false
#| results: 'asis'

render_spatial_map(
  data = data_3608,
  gear_code = "pl",
  gear_name = "pole line",
  target_species = c("skj", "yft", "bet"),
  year_range = c(params$year - 4, params$year),
  intro_text = ""

)
```

```{r}
#| echo: false
#| results: asis
#| eval: !expr nrow(data_2953) > 0 && "gear" %in% names(data_2953) && nrow(data_2953 |> filter(gear == "s")) > 0

if (nrow(data_3605) > 0) {
  cat("\\newpage\n\n")
  cat("## Captures of species of special interest (SSIs)\n\n")
  cat("Note that the tables below are specified to include - seabird, turtle and marine mammals in the instructions but we also include things like some rays and sharks. Note that Alive plus Dead does not necessarily equal Total due to unknowns. ")
}
```

@tbl-tbl3s provides a summary of observed interactions with special interest species by `r country_name`’s purse seine fleet from `r report_year - 4` to `r report_year`. In case of missing years in the table below, this indicates that no interactions with target species were reported during those years.

```{r}
ssi_s_caption <- if(nrow(data_2953) > 0 && "gear" %in% names(data_2953) && nrow(data_2953 |> filter(gear == "s")) > 0) {
  glue::glue("Observed annual estimated catches of species of special interest (seabird, turtle and marine mammals) by the {country_name} purse seine fleet, in the WCPFC Convention Area, for years {yr_range}, to the extent possible. Note that the Total may not reflect the sum of Alive and Dead as some animals are reported with an Unknown status. Source - Tufman 2, report 2953.")
} else NULL
```

```{r}
#| label: tbl-tbl3s
#| eval: !expr "!is.null(ssi_s_caption)"
#| tbl-cap: !expr ssi_s_caption
#| echo: false
#| results: 'asis'
render_ssi_table(
  data = data_2953,
  gear_code = "s",
  gear_name = "purse seine",
  intro_text = ""
)
```

```{r}
ssi_l_caption <- if(nrow(data_2953) > 0 && "gear" %in% names(data_2953) && nrow(data_2953 |> filter(gear == "l")) > 0) {
  glue::glue("Observed annual estimated catches of species of special interest (seabird, turtle and marine mammals) by the {country_name} longline fleet, in the WCPFC Convention Area, for years {yr_range}. Source - Tufman 2, report 2953.")} else NULL

```

```{r}
#| label: tbl-tbl3ll
#| eval: !expr "!is.null(ssi_l_caption)"
#| tbl-cap: !expr ssi_l_caption
#| echo: false
#| results: 'asis'
render_ssi_table(
  data = data_2953,
  gear_code = "l",
  gear_name = "longline",
  intro_text = ""
)
```

```{r}
ssi_pl_caption <- if(nrow(data_2953) > 0 && "gear" %in% names(data_2953) && nrow(data_2953 |> filter(gear == "pl")) > 0) {
  glue::glue("Observed annual estimated catches of species of special interest (seabird, turtle and marine mammals) by the {country_name} pole line fleet, in the WCPFC Convention Area, for years {yr_range}. Source - Tufman 2, report 2953.")} else NULL

```

```{r}
#| label: tbl-tbl3lpl
#| eval: !expr "!is.null(ssi_pl_caption)"
#| tbl-cap: !expr ssi_pl_caption
#| echo: false
#| results: 'asis'
render_ssi_table(
  data = data_2953,
  gear_code = "pl",
  gear_name = "pole line",
  intro_text = ""
)
```

```{r}
if (nrow(data_3605) > 0){
  PS_tab <- 
  data_3605 |>  
  filter(gear_code == "s", 
         year %in% yrs_long,
         !sp_code %in% ps_species) |>  # Exclude only the key species
  group_by(sp_code, year) |> 
  summarise(catch = sum(catch, na.rm = TRUE), .groups = "drop") |> 
  pivot_wider(names_from = year, values_from = catch, values_fill = 0) |> 
  arrange(sp_code)
  
  # Create table of non-target species for longline
  LL_tab <- 
    data_3605 |>  
    filter(gear_code == "l", 
           year %in% yrs_long,
           !sp_code %in% ps_species) |>  # Exclude only the key species
    group_by(sp_code, year) |> 
    summarise(catch = sum(catch, na.rm = TRUE), .groups = "drop") |> 
    pivot_wider(names_from = year, values_from = catch, values_fill = 0) |> 
    arrange(sp_code)

}else{
  PS_tab <- data.frame()
  LL_tab <- data.frame()
}
```

```{r}
#| label: tbl-tbl4s
#| eval: !expr "nrow(data_3605) > 0"
#| tbl-cap: !expr glue::glue("Annual estimated catches of non-target, associated and dependent species, including sharks, by the {country_name} purse seine fleet, in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605.")
#| echo: false
#| results: 'asis'

# Ensure required columns are present even when empty
required_cols <- c("sp_code", as.character((params$year - 4):params$year))
if (nrow(PS_tab) > 0) {
  # Print header
  cat(glue::glue("## Annual catch estimates for non-target, associated and dependent species for the national purse seine fleet\n\n"))
  
 cat(glue::glue(
    "@tbl-tbl4s provides a summary of observed interactions with non-target ",
    "species by {country_name}'s purse seine fleet from {report_year - 4} to {report_year}. In case of missing years in the table below, this indicates that no interactions with non-target species were reported during those years. \n\n"
  ))
  
  flextable(PS_tab |>
              rename(Species = sp_code)) |>  
    set_table_properties(layout = "autofit", width = 0.95, align = "center") |> 
    fontsize(size = 9, part = "all") |> 
    width(j = 1, width = 2) |> 
    colformat_num(big.mark = ",") |> 
    align(align = "center", part = "all") |> 
    labelizor(labels = species_labels_all, part = "all") |>
    padding(padding.left = 5, padding.right = 5, part = "all")
}
```

```{r}
#| label: tbl-tbl4l
#| eval: !expr "nrow(data_3605) > 0"
#| tbl-cap: !expr glue::glue("Annual estimated catches of non-target, associated and dependent species, including sharks, by the {country_name} longline fleet, in the WCPFC Convention Area over {yr_range}. Source - Tufman 2, report 3605")
#| echo: false
#| results: 'asis'

# Ensure LL_tab has all expected columns even if empty
required_cols <- c("sp_code", as.character((params$year - 4):params$year))
if (nrow(LL_tab) > 0) {
  # Print header
  cat(glue::glue("## Annual catch estimates for non-target, associated and dependent species for the national longline fleet\n\n"))
  
  flextable(LL_tab |>
            rename(Species = sp_code)) |>    
    set_table_properties(layout = "autofit", width = 0.95, align = "center") |> 
    fontsize(size = 9, part = "all") |> 
    width(j = 1, width = 2) |> 
    colformat_num(big.mark = ",") |> 
    align(align = "center", part = "all") |> 
    labelizor(labels = species_labels_all, part = "all") |>
    padding(padding.left = 5, padding.right = 5, part = "all")
}
```

<!-- # Coastal State Reporting -->

```{r loadcoastal}
#| output: asis
#| echo: false

# parse the comma-separated string into a clean vector
sections <- strsplit(params$additional_sections, ",")[[1]] |>
  trimws() |>
  (\(x) x[x != ""])()   # drop empty strings

if ("artisanal" %in% sections) {
  # Check whether artisanal data exists BEFORE knitting the child
  ika_file        <- paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")
  has_ika         <- file.exists(ika_file)
  has_tails <- length(list.files(this_yr_folder, pattern = "3615")) > 0

  if (has_ika || has_tails) {
    cat("\\newpage\n\n# Coastal State Reporting\n\n")
    cat(knitr::knit_child("template_artisanal_pg.qmd", quiet = TRUE))
    cat("\n\n\\clearpage\n\n")
  } else {
    message("No artisanal data for ", country_cd, " - skipping coastal section")
  }
}
```

# Socio-economic factors

Papua New Guinea is focused on building its domestic tuna industry to an extent where the socio-economic benefits generated from the domestic tuna industry can offset those that are currently obtained from the bilateral access arrangements. 
Some of the government’s focus areas include increase downstream processing of tuna products and value adding, increase employment, increase export earnings, improve enabling infrastructure and improve industry operating environment. The government is also looking at expanding and enhancing International Trade and market access for fishery products as well as supporting, maintaining and encouraging genuine Foreign Direct Investment, and optimization of national participation in the tuna fishery and processing sector. 
The government of PNG is now working in collaboration with members within the Forum Fisheries Agency (FFA) under the East New Britain Initiative (ENBi) in implementing FFA Leaders directives to optimize socio-economic returns from their shared tuna resources.

# Exports

The quantity of processed tuna products exported from the PNG domestic industry has been steady between 100,000 mt and 200 000 mt from 2020 onward (@fig-revenue). Peak export quantity was recorded in 2024, reaching slightly above 250,000 tonnes with corresponding export revenues of approximately 690 million USD. The export quantities and values have been largely driven by market demands and availability of suitable raw materials, specifically, market preferred tuna sizes for processing. This increase in production was attributed to high catches within PNG’s national waters, primarily due to the La Nina southern oscillation.

![This chart illustrates the relationship between the total volume of tuna exports (measured in tonnes, represented by blue bars) and the corresponding export value (measured in millions of USD, represented by the red line) over a six-year period. After a period of relative stability between 2021 and 2023, both quantity and revenue saw a significant surge in 2024, peaking at approximately 280,000 tonnes and nearly $600 million, respectively. Projections for 2025 indicate a return to levels more consistent with the 2020 baseline, suggesting a stabilization of the market following the 2024 peak.](data/report_2025_pg/additional_files/coastal_revenue.png){#fig-revenue}

Export quantity and value from 2020 to 2025 show substantial variability when expressed in percentage terms (@fig-export-value). Export volumes declined by approximately 19% from 2020 to 2021 and a further 5% from 2021 to 2022. This downward trend reversed in 2023, with a moderate increase of about 11%, followed by a sharp surge of roughly 58% in 2024, marking the most significant expansion in the series. However, this growth was not sustained, as export quantity fell again by approximately 25% in 2025.
Export value exhibited a similar but less pronounced pattern. There was an initial decrease of about 9% between 2020 and 2021, followed by a slight increase of around 1% in 2022 and a further 2–3% rise in 2023. A substantial jump of approximately 41% occurred in 2024, aligning with the surge in export volume. In 2025, export value declined by roughly 29%, reflecting the contraction in export quantity and indicating sensitivity of total revenue to production changes.

![Trends in export value per tonne (USD/t) from 2020 to 2025, showing an increase in unit value between 2020 and 2022, followed by a steady decline through 2025. The peak in 2022 indicates the highest market value per unit, while the subsequent decrease suggests weakening prices or a shift toward lower-value exports despite fluctuations in export volume.](data/report_2025_pg/additional_files/coastal_trends_export.png){#fig-export-value}


Analysis of value per tonne reveals contrasting dynamics. Unit values increased consistently from 2020 to 2022, with gains of approximately 12% and 7% in successive years, indicating improving market prices or higher-value exports. This trend reversed in 2023 with a decline of about 8%, followed by further decreases of roughly 11% in 2024 and 6% in 2025. The decline in unit value during periods of increased production suggests potential market saturation or a shift toward lower-value products.

Overall, the percentage trends highlight a non-linear relationship between export quantity and value. While both indicators experienced strong growth in 2024, declining unit values suggest diminishing returns at higher production levels. These patterns point to the influence of market demand, supply dynamics, and external economic factors, emphasizing the need to monitor both volume and price indicators when assessing export performance and informing fisheries management strategies.

<!-- # Addendum -->

```{r loadaddendum}
#| output: asis
#| echo: false

# Set values the child can read
applicable_measures <- "all"

if ("addendum" %in% sections) {
  cat("\\newpage\n\n# Addendum\n\n")
  cat(knitr::knit_child("template_addendum.qmd", quiet = TRUE))
}
```
